---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
rm(list = ls())
#make figure 1
```

```{r}
library(stats)
library(tidyverse)
library(eha)
library(gridExtra)
library(patchwork)
library(viridis)
```

```{r}
#Create failure pdfs for 3 different shape parameters
times = c(0:500)
fail2.0a = dweibull(times,shape=2,scale=100)
fail1.0b = dweibull(times,shape=1,scale=50)
fail3.0b = dweibull(times,shape=3,scale=50)
fail2.0b = dweibull(times,shape=2,scale=50)
fail2.0c = dweibull(times,shape=2,scale=12.5)


examples = data.frame(Time=times,fail2.0a = fail2.0a, fail2.0b = fail2.0b, fail2.0c = fail2.0c, fail1.0b = fail1.0b,fail3.0b = fail3.0b)
#create cumulative survival for 3 different shape parameters
cum1.0b = 1-pweibull(times,shape=1,scale=50)
cum3.0b = 1-pweibull(times,shape=3,scale=50)
cum2.0a = 1-pweibull(times,shape=2,scale=100)
cum2.0b = 1-pweibull(times,shape=2,scale=50)
cum2.0c = 1-pweibull(times,shape=2,scale=12.5)
examples = cbind(examples, data.frame(cum2.0a = cum2.0a, cum2.0b=cum2.0b, cum2.0c=cum2.0c,cum3.0b = cum3.0b,cum1.0b = cum1.0b))

favg2.0a=which.max(cum2.0a<=0.5)
favg2.0b=which.max(cum2.0b<=0.5)
favg2.0c=which.max(cum2.0c<=0.5)
favg3.0b=which.max(cum3.0b<=0.5)
favg1.0b=which.max(cum1.0b<=0.5)
#create hazard function plot for 3 different shape parameters
haz2.0a = hweibull(times,shape=2,scale=100)
haz1.0b = hweibull(times,shape=1,scale=50)
haz2.0b = hweibull(times,shape=2,scale=50)
haz3.0b = hweibull(times,shape=3,scale=50)
haz2.0c = hweibull(times,shape=2,scale=12.5)
examples = cbind(examples, data.frame(haz2.0a = haz2.0a, haz2.0b = haz2.0b,haz1.0b = haz1.0b, haz3.0b = haz3.0b,haz2.0c=haz2.0c))
```



```{r}
common_scale_color <- scale_color_manual(name = "Scale", values = viridis(3),breaks=c("12.5","50","100"))
common_scale_linetype <- scale_linetype_manual(name="Shape", values = c("1"="dotted","2"="solid","3"="dashed"))
common_scale_shape <- scale_shape_manual(name="Shape", values = c("1"=15,"2"=16,"3"=17))

#cumulative survival plot
cum.plot = ggplot(data=examples)+
  geom_line(aes(Time,cum2.0a,color="100",linetype="2"),size=1.25)+
  geom_line(aes(Time,cum2.0b,color="50",linetype="2"),size=1.25)+
  geom_line(aes(Time,cum2.0c,color="12.5",linetype="2"),size=1.25)+
  geom_line(aes(Time,cum1.0b,color="50",linetype="1"),size=1.25)+
  geom_line(aes(Time,cum3.0b,color="50",linetype="3"),size=1.25)+
    geom_point(aes(x = Time, y = cum2.0a, shape = "1", color = "100"), alpha = 0) +  # Invisible points
  geom_point(aes(x = Time, y = cum2.0b, shape = "2", color = "50"), alpha = 0) +  # Invisible points
  geom_point(aes(x = Time, y = cum2.0c, shape = "3", color = "12.5"), alpha = 0)+
  labs(x="Time", y="S(t)")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), legend.position = "none")+
  coord_cartesian(xlim = c(0, 110))+
  common_scale_color+
  common_scale_linetype+
  common_scale_shape

#hazard function plot
haz.plot = ggplot(data=examples[c(1:110),])+
  geom_line(aes(Time,haz2.0a,color="100",linetype="2"),size=1.25)+
  geom_line(aes(Time,haz2.0b,color="50",linetype="2"),size=1.25)+
  geom_line(aes(Time,haz2.0c,color="12.5",linetype="2"),size=1.25)+
  geom_line(aes(Time,haz1.0b,color="50",linetype="1"),size=1.25)+
  geom_line(aes(Time,haz3.0b,color="50",linetype="3"),size=1.25)+
    geom_point(aes(x = Time, y = haz2.0a, shape = "1", color = "100"), alpha = 0) +  # Invisible points
  geom_point(aes(x = Time, y = haz2.0b, shape = "2", color = "50"), alpha = 0) +  # Invisible points
  geom_point(aes(x = Time, y = haz2.0c, shape = "3", color = "12.5"), alpha = 0)+
  labs(x="Time", y="h(t)")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), legend.position = "none")+
  coord_cartesian(xlim = c(0, 110))+
  common_scale_color+
  common_scale_linetype+
  common_scale_shape
#failtime density plot

#custom legend shape sizes
large_points <- function(data, params, size) {
  # Multiply by some number
  data$size <- data$size * 2
  draw_key_point(data = data, params = params, size = size)
}


fail.plot = ggplot(data=examples)+
  geom_line(aes(Time,fail2.0a,color="100",linetype="2"),size=1.25)+
  geom_line(aes(Time,fail2.0b,color="50",linetype="2"),size=1.25)+
  geom_line(aes(Time,fail2.0c,color="12.5",linetype="2"),size=1.25)+
  geom_line(aes(Time,fail1.0b,color="50",linetype="1"),size=1.25)+
  geom_line(aes(Time,fail3.0b,color="50",linetype="3"),size=1.25)+
  geom_point(aes(x=favg2.0a,y=-0.005,color="100",shape="2"),size=2,key_glyph = large_points)+
  geom_point(aes(x=favg2.0b,y=-0.005,color="50",shape="2"),size=2,key_glyph = large_points)+
  geom_point(aes(x=favg2.0c,y=-0.005,color="12.5",shape="2"),size=2,key_glyph = large_points)+
  geom_point(aes(x=favg1.0b,y=-0.005,color="50",shape="1"),size=2,key_glyph = large_points)+
  geom_point(aes(x=favg3.0b,y=-0.005,color="50",shape="3"),size=2,key_glyph = large_points)+
  labs(x="", y="f(t)")+
  common_scale_color+
  common_scale_linetype+
  common_scale_shape+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(),legend.key.width = unit(2,"cm"))+
  coord_cartesian(xlim = c(0, 110))

```


```{r}
combined_plot <- haz.plot / fail.plot / cum.plot + 
  plot_layout(guides = 'collect') &
  plot_annotation(tag_levels = "A")
combined_plot
```

```{r}
rm(list = ls())
#make figure 2
```

```{r}
library(flexsurv)
library(tidyverse)
library(gridExtra)
library(patchwork)
library(RColorBrewer)
```


```{r}
#Jorgenson et al. data
jorg = read.csv("Jorgensen_Static.csv")
jorg$event= rep(1,nrow(jorg))
colnames(jorg)[colnames(jorg)=="KDTime"] = "time"
colnames(jorg)[colnames(jorg)=="Temp"] = "RearTemp"
#see what species have enough data points to be usable
table(jorg$Species)
#see how many temperatures are covered by each species
spc.temp.tab = table(jorg$RearTemp,jorg$Species)
spc.temp.tab
#most of the species have enough data to be somewhat useful. But, some species/temp combinations do not. So I will remove any trial without at least 10 replicates. For trials with less data, flexsurv models typically can't converge. But if I instead excluded only trials for which the model wouldn't converge, I would be biasing the dataset.
trials.df = as.data.frame(spc.temp.tab>=10)
```

```{r}
sci.names = c("D. buzzatii","D. equinoxialis","D. immigrans","D. melanogaster","D. mercatorum", "D. mojavensis", "D. montana", "D. rufa", "D. subobscura", "D. suzukii", "D. virilis")
```

```{r}
shape.list = list()
shape.var.list = list()
scale.list = list()
scale.var.list = list()
var.log.tkd.list = list()
tkd.list = list()

for (i in 1:length(colnames(trials.df))){
  species = colnames(trials.df)[i]
#narrow down to this particular species
species.data = jorg[jorg$Species==species,]
temp.list = rownames(trials.df)[trials.df[,species]]
#narrow down to only include trials with enough data
species.data = species.data[species.data$RearTemp %in% temp.list,]
#extract weibull model parameters
k.species = c()
k.species.var = c()
#IMPORTANT: this is log(lambda), because flexsurvreg automatically puts scale on a log scale
lambda.species = c()
lambda.species.var = c()

var.log.tkd.species = c()
tkd.species = c()

#make Weibull fits for every temperature for this species. Append its shape and scale parameters to the vectors above
for (j in unique(species.data$RearTemp)){
mod.j = flexsurvreg(Surv(time,event) ~ 1, data = species.data[(species.data$RearTemp==j),], dist=
                               "weibull")
k.species.var = c(k.species.var, mod.j$cov[[1,1]])
lambda.species.var = c(lambda.species.var, mod.j$cov[[2,2]])
k.species = c(k.species,mod.j$coefficients[[1]])
lambda.species = c(lambda.species,mod.j$coefficients[[2]])

var.log.tkd.species = c(var.log.tkd.species,var(log10(species.data[(species.data$RearTemp==j),]$time)))
tkd.species = c(tkd.species, median(species.data[(species.data$RearTemp==j),]$time))

}
shape.list[[length(shape.list)+1]] = k.species
scale.var.list[[length(scale.var.list)+1]] = k.species.var
scale.list[[length(scale.list)+1]] = lambda.species
shape.var.list[[length(shape.var.list)+1]] = lambda.species.var
var.log.tkd.list[[length(var.log.tkd.list)+1]] = var.log.tkd.species
tkd.list[[length(tkd.list)+1]] = tkd.species
}
```



```{r}
#make dataframe for species combined line plot
wb.pars.df = data.frame(species=rep(NA,0),temp=rep(NA,0),shape=rep(NA,0),scale=rep(NA,0),var.log.tKD=rep(NA,0))

for (i in 1:length(sci.names)){
  species = colnames(trials.df)[i]
  temp.list = as.numeric(rownames(trials.df)[trials.df[,species]])
  shapes=shape.list[[i]]
  scales=scale.list[[i]]
  n.trials=length(shapes)
  spec.df=data.frame(species=rep(species,n.trials),temp=temp.list,shape=shapes,scale=scales,var.log.tKD=var.log.tkd.list[[i]],median.tKD=tkd.list[[i]])
  wb.pars.df=rbind(wb.pars.df,spec.df)
  
}
```

```{r}
#get modeled value of sCTmax (as defined by Jorgensen et al 2021) for each species
#these will be used to color code and order the species in the figure
sCTmax.df = data.frame(species=rep(NA,length(sci.names)),sCTmax=rep(NA,length(sci.names)))
for (i in 1:length(sci.names)){
  species = unique(wb.pars.df$species)[i]
  temp.list = as.numeric(rownames(trials.df)[trials.df[,i]])
  tdt.mod = lm(log10(median.tKD)~temp,data=wb.pars.df[wb.pars.df$species==species,])
  lm.slope = tdt.mod$coefficients[2]
  lm.int = tdt.mod$coefficients[1]
  sCT_max = (1-lm.int)/lm.slope
  sCTmax.df[i,]=c(species,sCT_max)
  
}

#enter scientific names
sci.names.key=data.frame(sci.names=sci.names,abbrev=c("Buz","Equ","Imm","Mel","Mer","Moj","Mon","Ruf","Sub","Suz","Vir"))
new.names.vec=rep(NA,nrow(wb.pars.df))
for (i in 1:nrow(wb.pars.df)){
new.names.vec[i]=sci.names.key$sci.names[sci.names.key$abbrev==wb.pars.df$species[i]]
  
}
wb.pars.df$species=new.names.vec

wb.pars.df$species = factor(wb.pars.df$species, levels=unique(wb.pars.df$species)[order(sCTmax.df$sCTmax, decreasing=TRUE)])
```



```{r}
#create tdt plot
tdt.plot = ggplot(data=wb.pars.df)+
  geom_line(aes(y=log10(median.tKD),x=temp,color=species),linewidth=1)+
  geom_hline(yintercept=1,linetype="dotted")+
  xlab("")+
  ylab(expression(log[10](t[f])))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species")




shape.plot = ggplot(data=wb.pars.df)+
  geom_line(aes(y=shape,x=temp,color=species),linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab("Shape")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species")

scale.plot = ggplot(data=wb.pars.df)+
  geom_line(aes(y=scale,x=temp,color=species),linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(log[10]("Scale")))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species")

varlogtkd.plot = ggplot(data=wb.pars.df)+
  geom_line(aes(y=var.log.tKD,x=temp,color=species),linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(var(log[10](t[f]))))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species") 

combined_plot <- wrap_plots(tdt.plot, scale.plot, varlogtkd.plot, shape.plot)+
  plot_layout(guides = 'collect')&
  theme(legend.position="bottom",
        legend.key.size = unit(0.3, "cm"),
    legend.key = element_rect(color = NA, fill = NA),
    legend.text = element_text(face = "italic")) & 
  plot_annotation(tag_levels = "A")
  
  
  combined_plot
```

#supplemental plot recreating fig 2 but with each species separately
Fig S3

```{r}

#iterate through each species
combined.plots=list()

for (i in 1: length(unique(wb.pars.df$species))){
  species_i = unique(wb.pars.df$species)[i]
  color_i = brewer.pal(n=11,name="RdYlBu")[which.max(levels(wb.pars.df$species)==species_i)]
  print(color_i)
  data_i = wb.pars.df[wb.pars.df$species==species_i,]
  
tdt.plot = ggplot(data=data_i)+
  geom_point(aes(y=log10(median.tKD),x=temp),color=color_i)+
  geom_hline(yintercept=1,linetype="dotted")+
  geom_smooth(aes(y=log10(median.tKD),x=temp),color=color_i,linewidth=1,method="lm",se=F)+
  xlab("")+
  ylab(expression(log[10](t[f])))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")


shape.mod_i = lm(shape~temp,data=data_i) 
shape.p.value_i = summary(shape.mod_i)$coefficients[2,4]

shape.plot = ggplot(data=data_i)+
  geom_point(aes(y=shape,x=temp),color=color_i)+
  geom_smooth(aes(y=shape,x=temp),color=color_i,linewidth=1,method="lm",se=F)+
  xlab("Temperature (\u00B0C)")+
  ylab("Shape")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(),legend.position="bottom")+
  annotate("text", 
           x = Inf, y = Inf, 
           label = paste0("p = ", signif(shape.p.value_i, 3)),
           hjust = 1.1, vjust = 1.5,
           size = 4) 

scale.plot = ggplot(data=data_i)+
  geom_point(aes(y=scale,x=temp),color=color_i)+
  geom_smooth(aes(y=scale,x=temp),color=color_i,linewidth=1,method="lm",se=F)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(log[10]("Scale")))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")

varlogtkd.plot = ggplot(data=data_i)+
  geom_line(aes(y=var.log.tKD,x=temp),color=color_i,linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(var(log[10](t[f]))))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),legend.position="bottom") 

combined_plot <- wrap_plots(tdt.plot, scale.plot, varlogtkd.plot, shape.plot)+
  plot_layout(guides = 'collect')&
  theme(legend.position="none",
        legend.key.size = unit(0.3, "cm"),
    legend.key = element_rect(color = NA, fill = NA),
    legend.text = element_text(face = "italic")) & 
  plot_annotation(tag_levels = "A")&
  ggtitle(species_i)
  
  
  print(combined_plot)
  
  # Ensure the output folder exists
if (!dir.exists("suppfig3")) dir.create("suppfig3")

# Build file name safely
filename <- paste0("suppfig3/", gsub(" ", "_", species_i), ".jpeg")

# Save the plot
ggsave(filename, plot = combined_plot, width = 10, height = 8, units = "in", dpi = 300)
combined.plots[[i]] = combined_plot
}
```

```{r}
fig.s3.a = wrap_plots(combined.plots[[1]],combined.plots[[2]],combined.plots[[3]],combined.plots[[4]],combined.plots[[5]],combined.plots[[6]])
fig.s3.a
fig.s3.b = wrap_plots(combined.plots[[7]],combined.plots[[8]],combined.plots[[9]],combined.plots[[10]],combined.plots[[11]])
fig.s3.b
```



#supplemental figure S1 residual plots
```{r}
#create a list to hold the residual plots for each species
plot.list = list()
#iterate through all species, create a residual plot for each, and append to the list
for (i in 1:length(unique(wb.pars.df$species))){
  species_i = unique(wb.pars.df$species)[i]
  data_i = wb.pars.df[wb.pars.df$species==species_i,]
  tdt.mod_i = lm(log10(median.tKD)~temp,data=data_i)
  tdt.slope_i = tdt.mod_i$coefficients[[2]]
  tdt.int_i = tdt.mod_i$coefficients[[1]]
  resids_i = log10(data_i$median.tKD)-(tdt.slope_i*data_i$temp+tdt.int_i)
  resid_df <- data.frame(
  temp = data_i$temp,
  resids = resids_i
)

plot_i <- ggplot(resid_df, aes(x = temp, y = resids)) +
  geom_point() +
  ggtitle(species_i) +
  ylab(expression("Residual "*log[10]*" "*t[f])) +
  xlab("Temperature (\u00B0C)")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"))

  plot.list[[i]]=plot_i
}
```

```{r}
combined_resid_plot <- wrap_plots(plot.list[[1]],plot.list[[2]],plot.list[[3]],plot.list[[4]],plot.list[[5]],plot.list[[6]],plot.list[[7]],plot.list[[8]],plot.list[[9]],plot.list[[10]],plot.list[[11]])



combined_resid_plot
```

```{r}
rm(list = ls())
#make figure 3
```

```{r}
library(stats)
library(tidyverse)
library(eha)
library(gridExtra)
library(flexsurv)
library(survival)
library(ggsurvfit)
library(RColorBrewer)
library(survminer)
library(ggplotify)
library(patchwork)
library(viridis)
```
First, I'll run the functions to produce Rezende and Weibull models. Rezende code obtained from Rezende et al 2020, slightly modified to account for different sampling rates.


```{r}
#Steps to this model:
#1) First fit weibull curves to each tdt curve (each temperature) for the data supplied
#2) Then fit a linear model to both weibull shape and scale vs temp
#3) Next, we deal with the fluctuating temp data:
#     start cumulative survival (S(t)) at 1 (S(t_0)=1)
#     for each timestep t_i:
#         Find the weibull shape and scale parameters corresponding to the temperature T_i at timestep t_i. These parameters give a cumulative survival function of time (time spent under the constant temperature used to make the Weibull TDT curve, which we will call t*. Not the fluctuating temperature t_i). This distribution has CDF F_w(t*), PDF f_w(t*), cumulative survival S_w(t*)=1-F_w(t*)
#         Find chance of surviving from time t_i-1 to t_i: 
#             i) First, we find: S_w^-1(S(t_i-1))=t*_i
#             ii) Calculate failure in that time interval f.rate_i=S(t_i)/S(t_i-1)
#             iii) Calculate cumulative mortality between t_i-1 and t_i as (t_i-(t_i-1))*f.rate_i
#             iiii)Let S_i=(S_i-1)-[(t_i-(t_i-1))*f_w(t*_i)]
#Note that this loop assumes the temperature is constant at T_i for the time interval from t_i-1 to t_i. You could alternatively take the max or average of the temperatures at these time points. 
#replace with shape="constant" if you want the model to assume constant shape
weibull.mod = function(TDT.dat,temp.dat,shape="linear"){
  #step 1
shape.vec = c()
shape.vec.var = c()
scale.vec = c()
scale.vec.var = c()
for (j in unique(TDT.dat$Tref)){
mod.j = flexsurvreg(Surv(time,event) ~ 1, data = TDT.dat[(TDT.dat$Tref==j),], dist=
                               "weibull")
shape.vec.var = c(shape.vec.var, mod.j$cov[[1,1]])
scale.vec.var = c(scale.vec.var, exp(mod.j$cov[[2,2]]))
shape.vec = c(shape.vec,mod.j$coefficients[[1]])
#exponentiate because the function flexsurvreg automatically puts scale on a log scale
scale.vec = c(scale.vec,exp(mod.j$coefficients[[2]]))
}

#step 2
shape.lm=lm(shape.vec~unique(TDT.dat$Tref))
shape.slope=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[2]]
shape.int=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[1]]
#constant shape lm
cons.shape.lm=lm(shape.vec~1)
cons.shape.int=cons.shape.lm$coefficients[[1]]
#note the log transformation
scale.lm=lm(log10(scale.vec)~unique(TDT.dat$Tref))
scale.slope=scale.lm$coefficients[[2]]
scale.int=scale.lm$coefficients[[1]]

shape.pars=c(shape.slope,shape.int,cons.shape.int)
logscale.pars=c(scale.slope,scale.int)

#predict under fluctuating temperatures with shape as a linear function of temperature
if (shape=="linear"){
#step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  shape_i = shape.slope*mean(T_i,T_prev)+shape.int
  mean.list=c(mean.list,scale_i*gamma(1+1/shape_i))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  t_star_i=scale_i*(-log(cum.surv[i-1]))^(1/shape_i)
  #weibull pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pweibull(t_star_i,shape_i,scale_i))-(1-pweibull(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)]))
}

if (shape=="constant"){
 #step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  #this is the only line that is different
  shape_i = cons.shape.int
  mean.list=c(mean.list,scale_i*gamma(1+1/shape_i))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  t_star_i=scale_i*(-log(cum.surv[i-1]))^(1/shape_i)
  #weibull pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pweibull(t_star_i,shape_i,scale_i))-(1-pweibull(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)])) 
  
  
  
}

output=list(med.tdt,expected.tdt,cum.surv.post,shape.pars,logscale.pars)
names(output)=c("med.tdt","expected.tdt","cum.surv","shape.pars","logscale.pars")
return(output)
}
```

#3 Rezende et al model

```{r}
#First function produces static "tolerance landscape" from a vector of temperatures and a vector of thermal death times


	tolerance.landscape <- function(ta,time){
	
			data <- data.frame(ta,time)
			data <- data[order(data$ta,data$time),]

			# Step 1: Calculate CTmax and z from TDT curve
			ta <- as.numeric(levels(as.factor(data$ta)))			
			model <- lm(log10(data$time) ~ data$ta); summary(model)
			ctmax <- -coef(model)[1]/coef(model)[2]
			z <- -1/coef(model)[2]

			# Step 2: Calculate average log10 time and Ta (mean x and y for interpolation purposes)
			time.mn <- mean(log10(data$time))
			ta.mn <- mean(data$ta)
	
			# Step 3: Interpolating survival probabilities to make them comparable across treatments 
			time.interpol <- matrix(,1001,length(ta))
			for(i in 1:length(ta)){	
			time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(0,100,length.out = length(time))
			time.interpol[,i] <- approx(p,time,n = 1001)$y}			

			# Step 4: Overlap all survival curves into a single one by shifting each curve to mean x and y employing z
			# Step 5: Build expected survival curve with median survival time for each survival probability
			shift <- (10^((ta - ta.mn)/z))
			time.interpol.shift <- t(t(time.interpol)*shift)[-1,]
			surv.pred <- 10^apply(log10(time.interpol.shift),1,median) 	
	
			# Step 6: Expand predicted survival curves to measured Ta (matrix m arranged from lower to higher ta)
			# Step 7: Obtain predicted values comparable to each empirical measurement
			m <- surv.pred*matrix ((10^((ta.mn - rep(ta, each = 1000))/z)), nrow = 1000)
			out <-0
			for(i in 1:length(ta)){
				time <- c(0,data$time[data$ta==ta[i]]); p <- seq(0,100,length.out = length(time))
				out <- c(out,approx(seq(0,100,length.out = 1000),m[,i],xout=p[-1])$y)}
				data$time.pred <- out[-1]
				colnames(m) <- paste("time.at",ta,sep=".")
				m <- cbind(surv.prob=seq(1,0.001,-0.001),m)

			par(mfrow=c(1,2),mar=c(4.5,4,1,1),cex.axis=1.1)
			plot(-10,-10,las=1,xlab="Time (min)",ylab="Survival (%)",col="white",xaxs="i",yaxs="i",xlim=c(0,max(data$time)*1.05),ylim=c(0,105))
				for(i in 1:length(ta)){
				time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(100,0,length.out = length(time))
				points(time,p,pch=21,bg="black",cex=0.5)
				time <- c(0,sort(data$time.pred[data$ta==ta[i]]))
				points(m[,i+1],100*m[,1],type="l",lty=2)}
				segments(max(data$time)*0.7,90,max(data$time)*0.8,90,lty=2)
				text(max(data$time)*0.82,90,"fitted",adj=c(0,0.5))
			plot(log10(data$time.pred),log10(data$time),pch=21,bg="black",cex=0.5,lwd=0.7,las=1,xlab="Fitted Log10 time",ylab="Measured Log10 time")
			abline(0,1,lty=2)
			rsq <- round(summary(lm(log10(data$time) ~ log10(data$time.pred)))$r.square,3)
			text(min(log10(data$time.pred)),max(log10(data$time)),substitute("r"^2*" = "*rsq),adj=c(0,1))
			list(ctmax = as.numeric(ctmax), z = as.numeric(z), ta.mn = ta.mn,  S = data.frame(surv=seq(0.999,0,-0.001),time=surv.pred),
			time.obs.pred=cbind(data$time,data$time.pred), rsq = rsq)}	
```

```{r}
#ta is vector of fluctuating temperature conditions
#times is a vector of the corresponding times
	dynamic.landscape <- function(ta,times,tolerance.landscape){
	  #These are the predicted survival times that correspond to set cumulative survival values for the average temperature
			surv <- tolerance.landscape$S[,2]
			#mean temperature from TDT curves
			ta.mn <- tolerance.landscape$ta.mn
			#thermal tolerance coefficient
			z <- tolerance.landscape$z
			#the change in cumulative survival for a given temperature
			shift <- 10^((ta.mn - ta)/z)	
			time.rel <- 0
			#this will become a vector of cumulative survival percentages
			alive <- 100
			for(i in 1:(length(ta)-1)){	
			  #this is the size of the timestep in question
			  tstep=times[i+1]-times[i]
				if(alive[length(na.omit(alive))] > 0) {		
				  #approx fits a linear regression of relative time 
					alive <- try(c(alive,approx(c(0,shift[i]*surv),seq(100,0,length.out = length(c(0,surv))),xout = time.rel[i] + tstep)$y),silent=TRUE)
					#This essentially uses the inverse CDF of the shifted survival curve to approximate the comparable static temperature time using the next temperature
					time.rel <- try(c(time.rel,approx(seq(100,0,length.out = length(c(0,surv))),c(0,shift[i + 1]*surv),xout = alive[i + 1])$y),silent=TRUE)}
				else{
					alive <- 0}}				
			out <- data.frame(cbind(ta=ta[1:(length(alive)-1)],time=times[1:(length(alive)-1)],alive=alive[1:(length(alive)-1)]))
			par(mar=c(4,4,1,1),mfrow=c(1,2))
			plot(1:length(ta),ta,type="l",xlim=c(0,length(ta)),ylim=c(min(ta),max(ta)),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Temperature (ºC)")			
			plot(out$time,out$alive,type="l",xlim=c(0,length(ta)),ylim=c(0,100),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Survival (%)")
			list(time = out$time,ta = out$ta, alive = out$alive)}
```



```{r}
#Jorgenson et al. data
jorg = read.csv("Jorgensen_Static.csv")
jorg$event= rep(1,nrow(jorg))
colnames(jorg)[colnames(jorg)=="KDTime"] = "time"
colnames(jorg)[colnames(jorg)=="Temp"] = "RearTemp"
#see what species have enough data points to be usable
table(jorg$Species)
#see how many temperatures are covered by each species
spc.temp.tab = table(jorg$RearTemp,jorg$Species)
spc.temp.tab
#most of the species have enough data to be somewhat useful. But, some species/temp combinations do not. So I will remove any trial without at least 10 replicates. For trials with less data, models typically can't converge. But if I instead excluded only trials for which the model wouldn't converge, I would be biasing the dataset.
trials.df = as.data.frame(spc.temp.tab>=10)
```

```{r}
sci.names = c("D. buzzatii","D. equinoxialis","D. immigrans","D. melanogaster","D. mercatorum", "D. mojavensis", "D. montana", "D. rufa", "D. subobscura", "D. suzukii", "D. virilis")
```


```{r}
surv.fits.list=list()
rz.fits.list=list()
#weibull with changing shape
wb.fits.list=list()
#for every species
for (i in 1:length(colnames(trials.df))){
  species = colnames(trials.df)[i]
#narrow down to this particular species
species.data = jorg[jorg$Species==species,]
temp.list = rownames(trials.df)[trials.df[,species]]
#narrow down to only include trials with enough data
species.data = species.data[species.data$RearTemp %in% temp.list,]
#km survival fit to the data
surv.fit = survfit(Surv(time, event) ~ as.numeric(RearTemp), data = species.data)
fit.summary = summary(surv.fit)
times = fit.summary$time
survs=fit.summary$surv
strats=fit.summary$strata
temps=sapply(strats, function(x) sub("^as\\.numeric\\(RearTemp\\)=", "", x))
#make cumulative survival curves out of raw data
surv.fits.list[[i]]=data.frame(times=times,cum.surv=survs,temp=as.numeric(temps))
names(surv.fits.list)[i]=species

#Rezende fits
tol.land.spec = tolerance.landscape(species.data$RearTemp,species.data$time)

ntemp=length(unique(species.data$RearTemp))
             
rz.cum.surv=rep(tol.land.spec$S[,1],ntemp)
rz.surv.times=rep(tol.land.spec$S[,2],ntemp)
rz.temp=c()
rz.times=c()
#run models from Rezende et al 2020
for (j in 1:ntemp){
  n.per.temp=length(tol.land.spec$S[,1])
  rz.temp[c((1+(j-1)*n.per.temp):(j*n.per.temp))] = rep(unique(species.data$RearTemp)[j],n.per.temp)
}

ta.mn = tol.land.spec$ta.mn
z = tol.land.spec$z


rz.times = rz.surv.times*10^((ta.mn - rz.temp)/z)

rz.fits.list[[i]] = data.frame(times=rz.times,cum.surv=rz.cum.surv,temp=as.numeric(rz.temp))
names(rz.fits.list)[i]=species


#failure density by temp
#failure rates by temp
rz.fits.list[[i]]$fail.rates = rep(0,nrow(rz.fits.list[[i]]))
for (j in 1:ntemp){
  temp_j=unique(rz.fits.list[[i]]$temp)[j]
  cum.surv_j = rz.fits.list[[i]]$cum.surv[rz.fits.list[[i]]$temp==temp_j]
  times_j = rz.fits.list[[i]]$times[rz.fits.list[[i]]$temp==temp_j]
density_j = c(-(cum.surv_j[2:length(cum.surv_j)]-cum.surv_j[1:(length(cum.surv_j)-1)]))
delta_t_j= c(times_j[2:length(times_j)]-times_j[1:(length(times_j)-1)])
rz.fits.list[[i]]$fail.rates[rz.fits.list[[i]]$temp==temp_j]=c(NA,density_j/delta_t_j)
}

#Weibull fits
species.data$Tref=species.data$RearTemp
#temp dat doesn't matter: it is fluctuating temp conditions, but we only care about the parameters used to make them. I've included filler here.
wb.spec = weibull.mod(temp.dat=data.frame(Time_min=rep(1,10),Temperature=rep(35,10)),TDT.dat=species.data)
#cut off times as end of last rezende or survival fit, just for graphical clarity
t.max = max(c(rz.times,times))
wb.cum.surv=rep(tol.land.spec$S[,1],ntemp)
wb.surv.times=rep(tol.land.spec$S[,2],ntemp)
wb.temp=c()
wb.times=c()

for (j in 1:ntemp){
  time.seq = seq(0,t.max,0.1)
  n.times=length(time.seq)
  wb.times[(1+(j-1)*n.times):(j*n.times)] = time.seq
  
  wb.temp[(1+(j-1)*n.times):(j*n.times)] = rep(unique(species.data$RearTemp)[j],n.times)
}

#fit weibull distribution parameters as a linear function of temperature
shapes.vec = wb.spec$shape.pars[1]*wb.temp+wb.spec$shape.pars[2]
scales.vec = 10^(wb.spec$logscale.pars[1]*wb.temp+wb.spec$logscale.pars[2])

for (j in 1:nrow(species.data)){
  #failtime for individual j
  x_j = species.data$time[j]
  #temp of individual j
  temp_j = species.data$Tref[j]
  lin.shape_j = wb.spec$shape.pars[1]*temp_j+wb.spec$shape.pars[2]
  scale_j = 10^(wb.spec$logscale.pars[1]*temp_j+wb.spec$logscale.pars[2])
}

#cumulative survival
wb.cum.surv = 1-pweibull(wb.times,shape=shapes.vec,scale=scales.vec)

wb.fits.list[[i]] = data.frame(times=wb.times,cum.surv=wb.cum.surv,temp=as.numeric(wb.temp),shapes=shapes.vec,scales=scales.vec)
names(wb.fits.list)[i]=species


#failure rates by temp
wb.fits.list[[i]]$density = rep(0,nrow(wb.fits.list[[i]]))
for (j in 1:ntemp){
  temp_j=unique(wb.fits.list[[i]]$temp)[j]
  cum.surv_j = wb.fits.list[[i]]$cum.surv[wb.fits.list[[i]]$temp==temp_j]
  times_j = wb.fits.list[[i]]$times[wb.fits.list[[i]]$temp==temp_j]
density_j = c(-(cum.surv_j[2:length(cum.surv_j)]-cum.surv_j[1:(length(cum.surv_j)-1)]))
delta_t_j= c(times_j[2:length(times_j)]-times_j[1:(length(times_j)-1)])
wb.fits.list[[i]]$fail.rates[wb.fits.list[[i]]$temp==temp_j]=c(NA,density_j/delta_t_j)
}


}
```


Note that the functions to produce Rezende model fits are in the files "Thermal_landscape_functions.R" 



#density plots
Need to calculate density. But this is difficult given that some of our curves are discrete and some are continuous. To make matters simpler and put all on an even playing field, I'll estimate densities by simulation. I'll generate 10000 unif(0,1) random variables, and treat these as cumulative survival values. Then, I'll map them to associated failure time values, giving me simulated random variables with the density of the model in question.
```{r}
#this function takes a dataframe with one column "times", a second column "cum.surv", and a third column "temp". These should be ordered in increasing time/decreasing cumulative survival. n is the number of random data points generated.
generate.density = function(cdf,n){
  
  temps.list = unique(cdf$temp)
  fail.ests = data.frame(
    tf = rep(NA,0),
    temp = rep(NA,0),
    rv = rep(NA,0))
    rvs = runif(n)
  
for (i in 1:length(temps.list)){
  temp_i = temps.list[i]
  cdf_i = cdf[cdf$temp==temp_i,]
#vector of n unif(0,1) random variables
#vector that will contain all of the simulated failure times
fail.ests.i = rep(NA,n)
for (j in 1:n){
  #the first time at which cumulative survival is less than or equal to our simulated random variable.
  #rounding our time up like this is practical, because if data is collected every dt minutes, and failure occurs just after time t, it won't be detected until time t+dt.
  fail.ests.i[j] = cdf_i$times[cdf_i$cum.surv<=rvs[j]][1]
}
#set an arbitrarily large value to the simulated points that are outside of our graph window. This way they will still be accounted for in the density metric, but won't show up on the graph.
fail.ests.i[is.na(fail.ests.i)==TRUE]=100000
fail.ests=rbind(fail.ests,data.frame(tf=fail.ests.i,temp=rep(temp_i,n),rv=rvs))
}
return(fail.ests)
}
```

```{r}
temp.list=seq(33,41,0.5)
common_color_scale=scale_color_manual(name = "Temperature", values = plasma(length(temp.list)),breaks=as.character(temp.list),limits=as.character(temp.list))

common.linetype.scale=scale_linetype_manual(values=c("Rezende"="dotted","Weibull"="dashed","Data"="solid"))

data.a=surv.fits.list$Mel[surv.fits.list$Mel$temp%in%c("36","41"),]
rz.a = rz.fits.list$Mel[rz.fits.list$Mel$temp%in%c("36","41"),]
wb.a = wb.fits.list$Mel[wb.fits.list$Mel$temp%in%c("36","41"),]

data.dens.a=generate.density(
  cdf = data.a[,c("times","cum.surv","temp")],n=10000)
rz.dens.a = generate.density(
  cdf = rz.a[,c("times","cum.surv","temp")],n=10000)
wb.dens.a = generate.density(
  cdf = wb.a[,c("times","cum.surv","temp")],n=10000)

densplot.a = ggplot(data=NULL)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Rezende"),data=rz.dens.a,size=1)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Weibull"),data=wb.dens.a,size=1)+
  geom_rug(aes(x=log10(tf),group=as.factor(temp),color=temp),data=data.dens.a,sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp, linetype="Data"),data=data.dens.a)+
  ggtitle(expression("B) " * italic("D. melanogaster")))+
  xlim(0,3)+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))

data.b=surv.fits.list$Sub[surv.fits.list$Sub$temp%in%c("33","40"),]
rz.b = rz.fits.list$Sub[rz.fits.list$Sub$temp%in%c("33","40"),]
wb.b = wb.fits.list$Sub[wb.fits.list$Sub$temp%in%c("33","40"),]
data.dens.b=generate.density(
  cdf = data.b[,c("times","cum.surv","temp")],n=10000)
rz.dens.b = generate.density(
  cdf = rz.b[,c("times","cum.surv","temp")],n=10000)
wb.dens.b = generate.density(
  cdf = wb.b[,c("times","cum.surv","temp")],n=10000)

densplot.b = ggplot(data=NULL)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Rezende"),data=rz.dens.b,size=1)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Weibull"),data=wb.dens.b,size=1)+
  geom_rug(aes(x=log10(tf),group=as.factor(temp),color=temp),data=data.dens.b,sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp, linetype="Data"),data=data.dens.b)+
  ggtitle(expression("A) " * italic("D. subobscura")))+
  xlim(0,3)+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))

data.c=surv.fits.list$Moj[surv.fits.list$Moj$temp%in%c("39.5","43.5"),]
rz.c = rz.fits.list$Moj[rz.fits.list$Moj$temp%in%c("39.5","43.5"),]
wb.c = wb.fits.list$Moj[wb.fits.list$Moj$temp%in%c("39.5","43.5"),]
data.dens.c=generate.density(
  cdf = data.c[,c("times","cum.surv","temp")],n=10000)
rz.dens.c = generate.density(
  cdf = rz.c[,c("times","cum.surv","temp")],n=10000)
wb.dens.c = generate.density(
  cdf = wb.c[,c("times","cum.surv","temp")],n=10000)

densplot.c = ggplot(data=NULL)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Rezende"),data=rz.dens.c,size=1)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Weibull"),data=wb.dens.c,size=1)+
  geom_rug(aes(x=log10(tf),group=as.factor(temp),color=temp),data=data.dens.c,sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp, linetype="Data"),data=data.dens.c)+
  ggtitle(expression("D) " * italic("D. mojavensis")))+
  xlim(0,3)+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))

data.d=surv.fits.list$Vir[surv.fits.list$Vir$temp%in%c("37.5","41.5"),]
rz.d = rz.fits.list$Vir[rz.fits.list$Vir$temp%in%c("37.5","41.5"),]
wb.d = wb.fits.list$Vir[wb.fits.list$Vir$temp%in%c("37.5","41.5"),]

data.dens.d=generate.density(
  cdf = data.d[,c("times","cum.surv","temp")],n=10000)
rz.dens.d = generate.density(
  cdf = rz.d[,c("times","cum.surv","temp")],n=10000)
wb.dens.d = generate.density(
  cdf = wb.d[,c("times","cum.surv","temp")],n=10000)

densplot.d = ggplot(data=NULL)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Rezende"),data=rz.dens.d,size=1)+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp,linetype="Weibull"),data=wb.dens.d,size=1)+
  geom_rug(aes(x=log10(tf),group=as.factor(temp),color=temp),data=data.dens.d,sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp),color=temp, linetype="Data"),data=data.dens.d)+
  ggtitle(expression("C) " * italic("D. virilis")))+
  xlim(0,3)+
  common.linetype.scale+
  theme(axis.title = element_blank())

combined.dens.plots = wrap_plots(densplot.b,densplot.a,densplot.d,densplot.c)+
  plot_layout(guides = 'collect') & 
  scale_colour_gradientn(colours = plasma(length(temp.list)),
                         limits = range(data.a$temp, data.b$temp, data.c$temp, data.d$temp))&
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line())&
  ylab("Failure Density")&
  xlab(expression(log[10]*"(Minutes)"))&
  labs(color=expression(""*~degree*C*""),linetype="Fit")

```


```{r}
main_grid <- wrap_plots(densplot.b, densplot.a, densplot.d, densplot.c, ncol = 2) +
  plot_layout(guides = "collect") &
  scale_colour_gradientn(
    colours = plasma(length(temp.list)),
    limits = range(data.a$temp, data.b$temp, data.c$temp, data.d$temp)
  ) &
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line()
  ) &
  labs(color = expression("" * ~degree * C * ""), linetype = "Fit")

# Create axis label plots
x_lab_plot <- ggplot() + 
  theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = expression(log[10]*"(Minutes)"), size = 5, angle = 0)

y_lab_plot <- ggplot() + 
  theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = "Failure Density", size = 5, angle = 90)

# Assemble everything in a layout
final_plot <- (
  y_lab_plot + main_grid + plot_spacer() +
  plot_spacer() + x_lab_plot + plot_spacer()
) +
  plot_layout(
    ncol = 3,
    nrow = 2,
    widths = c(0.05, 0.9, 0.05),
    heights = c(0.9, 0.1)
  )

print(final_plot)
```


```{r}
#Make Fig. 4
rm(list = ls())
```

```{r}
library(tidyverse)
library(patchwork)
library(viridis)
library(flexsurv)
```

```{r}
#load functions for making Weibull and Rezende models
```

```{r}
#Steps to this model:
#1) First fit weibull curves to each tdt curve (each temperature) for the data supplied
#2) Then fit a linear model to both weibull shape and scale vs temp
#3) Next, we deal with the fluctuating temp data:
#     start cumulative survival (S(t)) at 1 (S(t_0)=1)
#     for each timestep t_i:
#         Find the weibull shape and scale parameters corresponding to the temperature T_i at timestep t_i. These parameters give a cumulative survival function of time (time spent under the constant temperature used to make the Weibull TDT curve, which we will call t*. Not the fluctuating temperature t_i). This distribution has CDF F_w(t*), PDF f_w(t*), cumulative survival S_w(t*)=1-F_w(t*)
#         Find chance of surviving from time t_i-1 to t_i: 
#             i) First, we find: S_w^-1(S(t_i-1))=t*_i
#             ii) Calculate failure in that time interval f.rate_i=S(t_i)/S(t_i-1)
#             iii) Calculate cumulative mortality between t_i-1 and t_i as (t_i-(t_i-1))*f.rate_i
#             iiii)Let S_i=(S_i-1)-[(t_i-(t_i-1))*f_w(t*_i)]
#Note that this loop assumes the temperature is constant at T_i for the time interval from t_i-1 to t_i. You could alternatively take the max or average of the temperatures at these time points. 
#replace with shape="constant" if you want the model to assume constant shape
weibull.mod = function(TDT.dat,temp.dat,shape="linear"){
  #step 1
shape.vec = c()
shape.vec.var = c()
scale.vec = c()
scale.vec.var = c()
for (j in unique(TDT.dat$Tref)){
mod.j = flexsurvreg(Surv(time,event) ~ 1, data = TDT.dat[(TDT.dat$Tref==j),], dist=
                               "weibull")
shape.vec.var = c(shape.vec.var, mod.j$cov[[1,1]])
scale.vec.var = c(scale.vec.var, exp(mod.j$cov[[2,2]]))
shape.vec = c(shape.vec,mod.j$coefficients[[1]])
#exponentiate because the function flexsurvreg automatically puts scale on a log scale
scale.vec = c(scale.vec,exp(mod.j$coefficients[[2]]))
}

#step 2
shape.lm=lm(shape.vec~unique(TDT.dat$Tref))
shape.slope=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[2]]
shape.int=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[1]]
#constant shape lm
cons.shape.lm=lm(shape.vec~1)
cons.shape.int=cons.shape.lm$coefficients[[1]]
#note the log transformation
scale.lm=lm(log10(scale.vec)~unique(TDT.dat$Tref))
scale.slope=scale.lm$coefficients[[2]]
scale.int=scale.lm$coefficients[[1]]

shape.pars=c(shape.slope,shape.int,cons.shape.int)
logscale.pars=c(scale.slope,scale.int)

#predict under fluctuating temperatures with shape as a linear function of temperature
if (shape=="linear"){
#step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  shape_i = shape.slope*mean(T_i,T_prev)+shape.int
  mean.list=c(mean.list,scale_i*gamma(1+1/shape_i))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  t_star_i=scale_i*(-log(cum.surv[i-1]))^(1/shape_i)
  #weibull pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pweibull(t_star_i,shape_i,scale_i))-(1-pweibull(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)]))
}

if (shape=="constant"){
 #step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  #this is the only line that is different
  shape_i = cons.shape.int
  mean.list=c(mean.list,scale_i*gamma(1+1/shape_i))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  t_star_i=scale_i*(-log(cum.surv[i-1]))^(1/shape_i)
  #weibull pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pweibull(t_star_i,shape_i,scale_i))-(1-pweibull(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)])) 
  
  
  
}

output=list(med.tdt,expected.tdt,cum.surv.post,shape.pars,logscale.pars)
names(output)=c("med.tdt","expected.tdt","cum.surv","shape.pars","logscale.pars")
return(output)
}
```

#3 Rezende et al model

```{r}
#First function produces static "tolerance landscape" from a vector of temperatures and a vector of thermal death times


	tolerance.landscape <- function(ta,time){
	
			data <- data.frame(ta,time)
			data <- data[order(data$ta,data$time),]

			# Step 1: Calculate CTmax and z from TDT curve
			ta <- as.numeric(levels(as.factor(data$ta)))			
			model <- lm(log10(data$time) ~ data$ta); summary(model)
			ctmax <- -coef(model)[1]/coef(model)[2]
			z <- -1/coef(model)[2]

			# Step 2: Calculate average log10 time and Ta (mean x and y for interpolation purposes)
			time.mn <- mean(log10(data$time))
			ta.mn <- mean(data$ta)
	
			# Step 3: Interpolating survival probabilities to make them comparable across treatments 
			time.interpol <- matrix(,1001,length(ta))
			for(i in 1:length(ta)){	
			time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(0,100,length.out = length(time))
			time.interpol[,i] <- approx(p,time,n = 1001)$y}			

			# Step 4: Overlap all survival curves into a single one by shifting each curve to mean x and y employing z
			# Step 5: Build expected survival curve with median survival time for each survival probability
			shift <- (10^((ta - ta.mn)/z))
			time.interpol.shift <- t(t(time.interpol)*shift)[-1,]
			surv.pred <- 10^apply(log10(time.interpol.shift),1,median) 	
	
			# Step 6: Expand predicted survival curves to measured Ta (matrix m arranged from lower to higher ta)
			# Step 7: Obtain predicted values comparable to each empirical measurement
			m <- surv.pred*matrix ((10^((ta.mn - rep(ta, each = 1000))/z)), nrow = 1000)
			out <-0
			for(i in 1:length(ta)){
				time <- c(0,data$time[data$ta==ta[i]]); p <- seq(0,100,length.out = length(time))
				out <- c(out,approx(seq(0,100,length.out = 1000),m[,i],xout=p[-1])$y)}
				data$time.pred <- out[-1]
				colnames(m) <- paste("time.at",ta,sep=".")
				m <- cbind(surv.prob=seq(1,0.001,-0.001),m)

			par(mfrow=c(1,2),mar=c(4.5,4,1,1),cex.axis=1.1)
			plot(-10,-10,las=1,xlab="Time (min)",ylab="Survival (%)",col="white",xaxs="i",yaxs="i",xlim=c(0,max(data$time)*1.05),ylim=c(0,105))
				for(i in 1:length(ta)){
				time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(100,0,length.out = length(time))
				points(time,p,pch=21,bg="black",cex=0.5)
				time <- c(0,sort(data$time.pred[data$ta==ta[i]]))
				points(m[,i+1],100*m[,1],type="l",lty=2)}
				segments(max(data$time)*0.7,90,max(data$time)*0.8,90,lty=2)
				text(max(data$time)*0.82,90,"fitted",adj=c(0,0.5))
			plot(log10(data$time.pred),log10(data$time),pch=21,bg="black",cex=0.5,lwd=0.7,las=1,xlab="Fitted Log10 time",ylab="Measured Log10 time")
			abline(0,1,lty=2)
			rsq <- round(summary(lm(log10(data$time) ~ log10(data$time.pred)))$r.square,3)
			text(min(log10(data$time.pred)),max(log10(data$time)),substitute("r"^2*" = "*rsq),adj=c(0,1))
			list(ctmax = as.numeric(ctmax), z = as.numeric(z), ta.mn = ta.mn,  S = data.frame(surv=seq(0.999,0,-0.001),time=surv.pred),
			time.obs.pred=cbind(data$time,data$time.pred), rsq = rsq)}	
```

```{r}
#ta is vector of fluctuating temperature conditions
#times is a vector of the corresponding times
	dynamic.landscape <- function(ta,times,tolerance.landscape){
	  #These are the predicted survival times that correspond to set cumulative survival values for the average temperature
			surv <- tolerance.landscape$S[,2]
			#mean temperature from TDT curves
			ta.mn <- tolerance.landscape$ta.mn
			#thermal tolerance coefficient
			z <- tolerance.landscape$z
			#the change in cumulative survival for a given temperature
			shift <- 10^((ta.mn - ta)/z)	
			time.rel <- 0
			#this will become a vector of cumulative survival percentages
			alive <- 100
			for(i in 1:(length(ta)-1)){	
			  #this is the size of the timestep in question
			  tstep=times[i+1]-times[i]
				if(alive[length(na.omit(alive))] > 0) {		
				  #approx fits a linear regression of relative time 
					alive <- try(c(alive,approx(c(0,shift[i]*surv),seq(100,0,length.out = length(c(0,surv))),xout = time.rel[i] + tstep)$y),silent=TRUE)
					#This essentially uses the inverse CDF of the shifted survival curve to approximate the comparable static temperature time using the next temperature
					time.rel <- try(c(time.rel,approx(seq(100,0,length.out = length(c(0,surv))),c(0,shift[i + 1]*surv),xout = alive[i + 1])$y),silent=TRUE)}
				else{
					alive <- 0}}				
			out <- data.frame(cbind(ta=ta[1:(length(alive)-1)],time=times[1:(length(alive)-1)],alive=alive[1:(length(alive)-1)]))
			par(mar=c(4,4,1,1),mfrow=c(1,2))
			plot(1:length(ta),ta,type="l",xlim=c(0,length(ta)),ylim=c(min(ta),max(ta)),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Temperature (ºC)")			
			plot(out$time,out$alive,type="l",xlim=c(0,length(ta)),ylim=c(0,100),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Survival (%)")
			list(time = out$time,ta = out$ta, alive = out$alive)}
```

```{r}
#for every time interval (t_i-1,t_i):
#1) Find max temperature over that time interval max(T_i,T_i-1)
#2) Find size of time interval (t_i-(t_i-1))
#3) Find damage accumulation rate: dd/dt=1/tKD(T)=1/(10^(Beta*T+alpha))
#4) Multiply rate by time interval to calculate accumulated damage for that time interval
#5) Add to total damage, record total damage at time t_i
#we start i-1 at the start time for the given trial group, and assume d=0 prior to this point. So, for trial group 1, i=2, but for trial group 3, which starts at about 117 mins, we start indexing at i=70 (calculations above)




jorg.mod = function(TDT.dat,temp.dat,KD.dat){
  #extract tdt coefficients
  tdt.lm = lm(log10.tcoma.~as.numeric(Tref),data=TDT.dat)
  tdt.slope=tdt.lm$coefficients[2]
  tdt.int=tdt.lm$coefficients[1]
  
  
  #injury ACC rate vector
acc.rates=rep(0,nrow(temp.dat))
#trial group 1
#initialize cumulative damage vector
cd.vec = rep(0,nrow(temp.dat))
#initialize timestep damage vector
td.vec = rep(0,nrow(temp.dat))
#initialize times vector
t.vec = temp.dat$Time_min
for (i in 2:nrow(temp.dat)){
  t_i = temp.dat$Time_min[i]
  t_prev = temp.dat$Time_min[i-1]
  T_i = temp.dat$Temperature[i]
  T_prev = temp.dat$Temperature[i-1]
  #1
  max.temp=max(T_i,T_prev)
  #2
  size.int=t_i-t_prev
  #3
  est.tKD = 10^(tdt.slope*max.temp+tdt.int)
  dd_dt = 1/est.tKD
  acc.rates[i]=dd_dt
  #4
  d.int = dd_dt*size.int
  td.vec[i-1]=d.int
  #5
  cd.vec[i]=cd.vec[i-1]+d.int
}
# 
# #now generate estimated damage at knockdown
# #initialize vector
dKD = rep(NA,nrow(KD.dat))

for (i in 1:nrow(KD.dat)){
   start.time = KD.dat$start[i]
   tKD = KD.dat$t_coma[i]
   dKD[i]=100*sum(td.vec[t.vec>=start.time & t.vec<=tKD])
}

est.tKD = rep(NA,nrow(KD.dat))
for (i in 1:nrow(KD.dat)){
  start.index.group = which.max(t.vec>=KD.dat$start[i])
  cd.group = cd.vec-cd.vec[start.index.group]
  est.tKD.index = which.max(cd.group>=1)
  est.tKD[i]=t.vec[est.tKD.index]
}
  
damage.df = data.frame(minutes=t.vec,cum.damage=cd.vec,interval.damage=td.vec)  
KD.dat$est.dKD = dKD  
KD.dat$est.tKD = est.tKD
#estimated and actual kdtimes starting at start time
KD.dat$est.tKD.adj = est.tKD-KD.dat$start
KD.dat$tcoma.adj=KD.dat$t_coma-KD.dat$start
ret.list=list(damage.df, KD.dat)
names(ret.list)=c("damage.df","KD.dat")


return(ret.list)
}
```

```{r}
#constant temperature knockdown data
#used to train models
niko.tdt = read.csv("Niko_temps_csvs/Nikolaj_TDT_raw.csv")
#remove rows without Tref
niko.tdt=niko.tdt[is.na(niko.tdt$Tref)==F,]
#add columns for flexsurv formatting
niko.tdt$time=niko.tdt$tcoma
niko.tdt$event=rep(1,nrow(niko.tdt))
head(niko.tdt)
females.tdt = niko.tdt[niko.tdt$sex=="f",]
males.tdt = niko.tdt[niko.tdt$sex=="m",]
#remove 2 points with no tcoma (not censored, just missing)
males.tdt=males.tdt[is.na(males.tdt$tcoma)==FALSE,]
```

```{r}
#import trial temp condition data
niko.temps = list(
  read.csv("Niko_temps_csvs/Niko_temps_exp1.csv"),
  read.csv("Niko_temps_csvs/Niko_temps_exp2.csv"),
  read.csv("Niko_temps_csvs/Niko_temps_exp3.csv"))
names(niko.temps)=c("exp1","exp2","exp3")
```

```{r}
#knockdown data associated with these trial conditions
#used to test models
niko.fluc.tKD = read.csv("Niko_temps_csvs/Nikolaj_fluc_tKD_formatted.csv")
head(niko.fluc.tKD)
```

```{r}
#fluctuating knockdown data
#females exp 1
exp1.f.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==1),]
exp2.f.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2),]
exp3.f.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==3),]
```

Jorgensen analysis
```{r}
#females
jorg.exp1.f = jorg.mod(TDT.dat=females.tdt,temp.dat=niko.temps[[1]],KD.dat=exp1.f.tKD.dat)
jorg.exp2.f = jorg.mod(TDT.dat=females.tdt,temp.dat=niko.temps[[2]],KD.dat=exp2.f.tKD.dat)
jorg.exp3.f = jorg.mod(TDT.dat=females.tdt,temp.dat=niko.temps[[3]],KD.dat=exp3.f.tKD.dat)
jorg.f.comb=rbind(jorg.exp1.f$KD.dat,jorg.exp2.f$KD.dat,jorg.exp3.f$KD.dat)
```

```{r}
#fluctuating knockdown data
#males exp 1
exp1.m.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==1),]
exp2.m.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2),]
exp3.m.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==3),]
```

```{r}
#males
jorg.exp1.m = jorg.mod(TDT.dat=males.tdt,temp.dat=niko.temps[[1]],KD.dat=exp1.m.tKD.dat)
jorg.exp2.m = jorg.mod(TDT.dat=males.tdt,temp.dat=niko.temps[[2]],KD.dat=exp2.m.tKD.dat)
jorg.exp3.m = jorg.mod(TDT.dat=males.tdt,temp.dat=niko.temps[[3]],KD.dat=exp3.m.tKD.dat)
jorg.m.comb=rbind(jorg.exp1.m$KD.dat,jorg.exp2.m$KD.dat,jorg.exp3.m$KD.dat)
plot(est.dKD~tcoma.adj,data=jorg.m.comb)
```

The Weibull and Rezende model functions just make predictions based on temp conditions: they don't adjust for group start times. So, we need to define the temperature conditions for each group.
```{r}
strt.times = exp1.f.tKD.dat %>% group_by(grp) %>% summarise(mean=mean(start))

#fluctuating temperature dataframe starting at each groups start time
temps.e1.g1 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[1,2]):nrow(niko.temps[[1]]),]
#subtract start time to rescale
temps.e1.g1$Time_min=temps.e1.g1$Time_min - temps.e1.g1$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[1,2])]

temps.e1.g2 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[[2,2]]):nrow(niko.temps[[1]]),]

temps.e1.g2$Time_min=temps.e1.g2$Time_min - temps.e1.g2$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[2,2])]

temps.e1.g3 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[[3,2]]):nrow(niko.temps[[1]]),]

temps.e1.g3$Time_min=temps.e1.g3$Time_min - temps.e1.g3$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[3,2])]

temps.e1.g4 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[[4,2]]):nrow(niko.temps[[1]]),]

temps.e1.g4$Time_min=temps.e1.g4$Time_min - temps.e1.g4$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[4,2])]
```

#exp 2
```{r}
strt.times = exp2.f.tKD.dat %>% group_by(grp) %>% summarise(mean=mean(start))

#fluctuating temperature dataframe starting at each groups start time
temps.e2.g1 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[1,2]):nrow(niko.temps[[2]]),]
#subtract start time to rescale
temps.e2.g1$Time_min=temps.e2.g1$Time_min - temps.e2.g1$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[1,2])]

temps.e2.g2 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[2,2]]):nrow(niko.temps[[2]]),]

temps.e2.g2$Time_min=temps.e2.g2$Time_min - temps.e2.g2$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[2,2])]

temps.e2.g3 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[3,2]]):nrow(niko.temps[[2]]),]

temps.e2.g3$Time_min=temps.e2.g3$Time_min - temps.e2.g3$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[3,2])]

temps.e2.g4 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[4,2]]):nrow(niko.temps[[2]]),]

temps.e2.g4$Time_min=temps.e2.g4$Time_min - temps.e2.g4$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[4,2])]

temps.e2.g5 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[5,2]]):nrow(niko.temps[[2]]),]

temps.e2.g5$Time_min=temps.e2.g5$Time_min - temps.e2.g5$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[5,2])]
```

```{r}
#exp 3
strt.times = exp3.f.tKD.dat %>% group_by(grp) %>% summarise(mean=mean(start))

#fluctuating temperature dataframe starting at each groups start time
temps.e3.g1 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[1,2]):nrow(niko.temps[[3]]),]
#subtract start time to rescale
temps.e3.g1$Time_min=temps.e3.g1$Time_min - temps.e3.g1$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[1,2])]

temps.e3.g2 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[[2,2]]):nrow(niko.temps[[3]]),]

temps.e3.g2$Time_min=temps.e3.g2$Time_min - temps.e3.g2$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[2,2])]

temps.e3.g3 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[[3,2]]):nrow(niko.temps[[3]]),]

temps.e3.g3$Time_min=temps.e3.g3$Time_min - temps.e3.g3$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[3,2])]

temps.e3.g4 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[[4,2]]):nrow(niko.temps[[3]]),]

temps.e3.g4$Time_min=temps.e3.g4$Time_min - temps.e3.g4$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[4,2])]

```

Weibull Analysis (linear shape)
```{r}
#exp 1
#females
wb.e1.g1.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g1)
wb.e1.g2.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g2)
wb.e1.g3.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g3)
wb.e1.g4.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g4)
c(wb.e1.g1.f$med.tdt,wb.e1.g2.f$med.tdt,wb.e1.g3.f$med.tdt,wb.e1.g4.f$med.tdt)
c(wb.e1.g1.f$expected.tdt,wb.e1.g2.f$expected.tdt,wb.e1.g3.f$expected.tdt,wb.e1.g4.f$expected.tdt)
#males
wb.e1.g1.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g1)
wb.e1.g2.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g2)
wb.e1.g3.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g3)
wb.e1.g4.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g4)
```

```{r}
#exp 2
wb.e2.g1.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g1)
wb.e2.g2.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g2)
wb.e2.g3.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g3)
wb.e2.g4.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g4)
wb.e2.g5.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g5)
c(wb.e2.g1.f$med.tdt,wb.e2.g2.f$med.tdt,wb.e2.g3.f$med.tdt,wb.e2.g4.f$med.tdt,wb.e2.g5.f$med.tdt)
c(wb.e2.g1.f$expected.tdt,wb.e2.g2.f$expected.tdt,wb.e2.g3.f$expected.tdt,wb.e2.g4.f$expected.tdt,wb.e2.g5.f$med.tdt)
#males
wb.e2.g1.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g1)
wb.e2.g2.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g2)
wb.e2.g3.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g3)
wb.e2.g4.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g4)
wb.e2.g5.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g5)
```

```{r}
#exp 3
wb.e3.g1.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g1)
wb.e3.g2.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g2)
wb.e3.g3.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g3)
wb.e3.g4.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g4)
c(wb.e3.g1.f$med.tdt,wb.e3.g2.f$med.tdt,wb.e3.g3.f$med.tdt,wb.e3.g4.f$med.tdt)
c(wb.e3.g1.f$expected.tdt,wb.e3.g2.f$expected.tdt,wb.e3.g3.f$expected.tdt,wb.e3.g4.f$expected.tdt)
#males
wb.e3.g1.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g1)
wb.e3.g2.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g2)
wb.e3.g3.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g3)
wb.e3.g4.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g4)
```

Weibull Analysis (constant shape)

```{r}
#exp 1
#females
wb.cons.e1.g1.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g1,shape="constant")
wb.cons.e1.g2.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g2,shape="constant")
wb.cons.e1.g3.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g3,shape="constant")
wb.cons.e1.g4.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g4,shape="constant")
c(wb.cons.e1.g1.f$med.tdt,wb.cons.e1.g2.f$med.tdt,wb.cons.e1.g3.f$med.tdt,wb.cons.e1.g4.f$med.tdt)
c(wb.cons.e1.g1.f$expected.tdt,wb.cons.e1.g2.f$expected.tdt,wb.cons.e1.g3.f$expected.tdt,wb.cons.e1.g4.f$expected.tdt)
#males
wb.cons.e1.g1.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g1,shape="constant")
wb.cons.e1.g2.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g2,shape="constant")
wb.cons.e1.g3.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g3,shape="constant")
wb.cons.e1.g4.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g4,shape="constant")
```

```{r}
#exp 2
wb.cons.e2.g1.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g1,shape="constant")
wb.cons.e2.g2.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g2,shape="constant")
wb.cons.e2.g3.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g3,shape="constant")
wb.cons.e2.g4.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g4,shape="constant")
wb.cons.e2.g5.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g5,shape="constant")
#males
wb.cons.e2.g1.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g1,shape="constant")
wb.cons.e2.g2.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g2,shape="constant")
wb.cons.e2.g3.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g3,shape="constant")
wb.cons.e2.g4.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g4,shape="constant")
wb.cons.e2.g5.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g5,shape="constant")
```

```{r}
#exp 3
wb.cons.e3.g1.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g1,shape="constant")
wb.cons.e3.g2.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g2,shape="constant")
wb.cons.e3.g3.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g3,shape="constant")
wb.cons.e3.g4.f = weibull.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g4,shape="constant")
#males
wb.cons.e3.g1.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g1,shape="constant")
wb.cons.e3.g2.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g2,shape="constant")
wb.cons.e3.g3.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g3,shape="constant")
wb.cons.e3.g4.m = weibull.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g4,shape="constant")
```


Rezende model output
Note: these functions are slightly modified to account for temperatures not collected once per minute
```{r}
#females
tol.land.females=tolerance.landscape(females.tdt$Tref,females.tdt$tcoma)
#exp 1
dy.land.e1.g1.f = dynamic.landscape(temps.e1.g1$Temperature,temps.e1.g1$Time_min,tol.land.females)
dy.land.e1.g2.f = dynamic.landscape(temps.e1.g2$Temperature,temps.e1.g2$Time_min,tol.land.females)
dy.land.e1.g3.f = dynamic.landscape(temps.e1.g3$Temperature,temps.e1.g3$Time_min,tol.land.females)
dy.land.e1.g4.f = dynamic.landscape(temps.e1.g4$Temperature,temps.e1.g4$Time_min,tol.land.females)
#median fail time
rz.med.e1.g1.f = dy.land.e1.g1.f$time[which.max(dy.land.e1.g1.f$alive<50)]
rz.med.e1.g2.f = dy.land.e1.g2.f$time[which.max(dy.land.e1.g2.f$alive<50)]
rz.med.e1.g3.f = dy.land.e1.g3.f$time[which.max(dy.land.e1.g3.f$alive<50)]
rz.med.e1.g4.f = dy.land.e1.g4.f$time[which.max(dy.land.e1.g4.f$alive<50)]

#exp 2
dy.land.e2.g1.f = dynamic.landscape(temps.e2.g1$Temperature,temps.e2.g1$Time_min,tol.land.females)
dy.land.e2.g2.f = dynamic.landscape(temps.e2.g2$Temperature,temps.e2.g2$Time_min,tol.land.females)
dy.land.e2.g3.f = dynamic.landscape(temps.e2.g3$Temperature,temps.e2.g3$Time_min,tol.land.females)
dy.land.e2.g4.f = dynamic.landscape(temps.e2.g4$Temperature,temps.e2.g4$Time_min,tol.land.females)
dy.land.e2.g5.f = dynamic.landscape(temps.e2.g5$Temperature,temps.e2.g5$Time_min,tol.land.females)
#median fail time
rz.med.e2.g1.f = dy.land.e2.g1.f$time[which.max(dy.land.e2.g1.f$alive<50)]
rz.med.e2.g2.f = dy.land.e2.g2.f$time[which.max(dy.land.e2.g2.f$alive<50)]
rz.med.e2.g3.f = dy.land.e2.g3.f$time[which.max(dy.land.e2.g3.f$alive<50)]
rz.med.e2.g4.f = dy.land.e2.g4.f$time[which.max(dy.land.e2.g4.f$alive<50)]
rz.med.e2.g5.f = dy.land.e2.g5.f$time[which.max(dy.land.e2.g5.f$alive<50)]

#exp 3
dy.land.e3.g1.f = dynamic.landscape(temps.e3.g1$Temperature,temps.e3.g1$Time_min,tol.land.females)
dy.land.e3.g2.f = dynamic.landscape(temps.e3.g2$Temperature,temps.e3.g2$Time_min,tol.land.females)
dy.land.e3.g3.f = dynamic.landscape(temps.e3.g3$Temperature,temps.e3.g3$Time_min,tol.land.females)
dy.land.e3.g4.f = dynamic.landscape(temps.e3.g4$Temperature,temps.e3.g4$Time_min,tol.land.females)
#median fail time
rz.med.e3.g1.f = dy.land.e3.g1.f$time[which.max(dy.land.e3.g1.f$alive<50)]
rz.med.e3.g2.f = dy.land.e3.g2.f$time[which.max(dy.land.e3.g2.f$alive<50)]
rz.med.e3.g3.f = dy.land.e3.g3.f$time[which.max(dy.land.e3.g3.f$alive<50)]
rz.med.e3.g4.f = dy.land.e3.g4.f$time[which.max(dy.land.e3.g4.f$alive<50)]

#males
tol.land.males=tolerance.landscape(males.tdt$Tref,males.tdt$tcoma)
#exp 1
dy.land.e1.g1.m = dynamic.landscape(temps.e1.g1$Temperature,temps.e1.g1$Time_min,tol.land.males)
dy.land.e1.g2.m = dynamic.landscape(temps.e1.g2$Temperature,temps.e1.g2$Time_min,tol.land.males)
dy.land.e1.g3.m = dynamic.landscape(temps.e1.g3$Temperature,temps.e1.g3$Time_min,tol.land.males)
dy.land.e1.g4.m = dynamic.landscape(temps.e1.g4$Temperature,temps.e1.g4$Time_min,tol.land.males)
#median fail time
rz.med.e1.g1.m = dy.land.e1.g1.m$time[which.max(dy.land.e1.g1.m$alive<50)]
rz.med.e1.g2.m = dy.land.e1.g2.m$time[which.max(dy.land.e1.g2.m$alive<50)]
rz.med.e1.g3.m = dy.land.e1.g3.m$time[which.max(dy.land.e1.g3.m$alive<50)]
rz.med.e1.g4.m = dy.land.e1.g4.m$time[which.max(dy.land.e1.g4.m$alive<50)]

#exp 2
dy.land.e2.g1.m = dynamic.landscape(temps.e2.g1$Temperature,temps.e2.g1$Time_min,tol.land.males)
dy.land.e2.g2.m = dynamic.landscape(temps.e2.g2$Temperature,temps.e2.g2$Time_min,tol.land.males)
dy.land.e2.g3.m = dynamic.landscape(temps.e2.g3$Temperature,temps.e2.g3$Time_min,tol.land.males)
dy.land.e2.g4.m = dynamic.landscape(temps.e2.g4$Temperature,temps.e2.g4$Time_min,tol.land.males)
dy.land.e2.g5.m = dynamic.landscape(temps.e2.g5$Temperature,temps.e2.g5$Time_min,tol.land.males)
#median fail time
rz.med.e2.g1.m = dy.land.e2.g1.m$time[which.max(dy.land.e2.g1.m$alive<50)]
rz.med.e2.g2.m = dy.land.e2.g2.m$time[which.max(dy.land.e2.g2.m$alive<50)]
rz.med.e2.g3.m = dy.land.e2.g3.m$time[which.max(dy.land.e2.g3.m$alive<50)]
rz.med.e2.g4.m = dy.land.e2.g4.m$time[which.max(dy.land.e2.g4.m$alive<50)]
rz.med.e2.g5.m = dy.land.e2.g5.m$time[which.max(dy.land.e2.g5.m$alive<50)]

#exp 3
dy.land.e3.g1.m = dynamic.landscape(temps.e3.g1$Temperature,temps.e3.g1$Time_min,tol.land.males)
dy.land.e3.g2.m = dynamic.landscape(temps.e3.g2$Temperature,temps.e3.g2$Time_min,tol.land.males)
dy.land.e3.g3.m = dynamic.landscape(temps.e3.g3$Temperature,temps.e3.g3$Time_min,tol.land.males)
dy.land.e3.g4.m = dynamic.landscape(temps.e3.g4$Temperature,temps.e3.g4$Time_min,tol.land.males)
#median fail time
rz.med.e3.g1.m = dy.land.e3.g1.m$time[which.max(dy.land.e3.g1.m$alive<50)]
rz.med.e3.g2.m = dy.land.e3.g2.m$time[which.max(dy.land.e3.g2.m$alive<50)]
rz.med.e3.g3.m = dy.land.e3.g3.m$time[which.max(dy.land.e3.g3.m$alive<50)]
rz.med.e3.g4.m = dy.land.e3.g4.m$time[which.max(dy.land.e3.g4.m$alive<50)]

```


```{r}
#let's compare model outputs from the three models
#females
#exp 1
exp1.f.summ = jorg.exp1.f$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
#linear shape
exp1.f.summ$wb.med.tkd=c(wb.e1.g1.f$med.tdt,wb.e1.g2.f$med.tdt,wb.e1.g3.f$med.tdt,wb.e1.g4.f$med.tdt)
exp1.f.summ$wb.cons.med.tkd=c(wb.cons.e1.g1.f$med.tdt,wb.cons.e1.g2.f$med.tdt,wb.cons.e1.g3.f$med.tdt,wb.cons.e1.g4.f$med.tdt)
#linear shape
exp1.f.summ$wb.mean.tkd=c(wb.e1.g1.f$expected.tdt,wb.e1.g2.f$expected.tdt,wb.e1.g3.f$expected.tdt,wb.e1.g4.f$expected.tdt)
#constant shape
exp1.f.summ$wb.cons.mean.tkd=c(wb.cons.e1.g1.f$expected.tdt,wb.cons.e1.g2.f$expected.tdt,wb.cons.e1.g3.f$expected.tdt,wb.cons.e1.g4.f$expected.tdt)
exp1.f.summ$rz.med.tkd=c(rz.med.e1.g1.f,rz.med.e1.g2.f,rz.med.e1.g3.f,rz.med.e1.g4.f)
exp1.f.summ$exp=1

#exp 2
exp2.f.summ = jorg.exp2.f$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
#linear shape
exp2.f.summ$wb.med.tkd=c(wb.e2.g1.f$med.tdt,wb.e2.g2.f$med.tdt,wb.e2.g3.f$med.tdt,wb.e2.g4.f$med.tdt,wb.e2.g5.f$med.tdt)
exp2.f.summ$wb.mean.tkd=c(wb.e2.g1.f$expected.tdt,wb.e2.g2.f$expected.tdt,wb.e2.g3.f$expected.tdt,wb.e2.g4.f$expected.tdt,wb.e2.g5.f$med.tdt)
#constant shape
exp2.f.summ$wb.cons.med.tkd=c(wb.cons.e2.g1.f$med.tdt,wb.cons.e2.g2.f$med.tdt,wb.cons.e2.g3.f$med.tdt,wb.cons.e2.g4.f$med.tdt,wb.cons.e2.g5.f$med.tdt)
exp2.f.summ$wb.cons.mean.tkd=c(wb.cons.e2.g1.f$expected.tdt,wb.cons.e2.g2.f$expected.tdt,wb.cons.e2.g3.f$expected.tdt,wb.cons.e2.g4.f$expected.tdt,wb.cons.e2.g5.f$med.tdt)
exp2.f.summ$rz.med.tkd=c(rz.med.e2.g1.f,rz.med.e2.g2.f,rz.med.e2.g3.f,rz.med.e2.g4.f,rz.med.e2.g5.f)
exp2.f.summ$exp=2

#exp 3
exp3.f.summ = jorg.exp3.f$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
#linear shape
exp3.f.summ$wb.med.tkd=c(wb.e3.g1.f$med.tdt,wb.e3.g2.f$med.tdt,wb.e3.g3.f$med.tdt,wb.e3.g4.f$med.tdt)
exp3.f.summ$wb.mean.tkd=c(wb.e3.g1.f$expected.tdt,wb.e3.g2.f$expected.tdt,wb.e3.g3.f$expected.tdt,wb.e3.g4.f$expected.tdt)
#constant shape
exp3.f.summ$wb.cons.med.tkd=c(wb.cons.e3.g1.f$med.tdt,wb.cons.e3.g2.f$med.tdt,wb.cons.e3.g3.f$med.tdt,wb.cons.e3.g4.f$med.tdt)
exp3.f.summ$wb.cons.mean.tkd=c(wb.cons.e3.g1.f$expected.tdt,wb.cons.e3.g2.f$expected.tdt,wb.cons.e3.g3.f$expected.tdt,wb.cons.e3.g4.f$expected.tdt)
exp3.f.summ$rz.med.tkd=c(rz.med.e3.g1.f,rz.med.e3.g2.f,rz.med.e3.g3.f,rz.med.e3.g4.f)
exp3.f.summ$exp=3

three.mod.results.f = rbind(exp1.f.summ,exp2.f.summ,exp3.f.summ)
three.mod.results.f$sex="f"
#note that later I will calculate error based on individual points, not medians. That will be better, but this will give us an idea for now.
three.mod.results.f$jorg.error=three.mod.results.f$est.tKD-three.mod.results.f$true.tKD.med
three.mod.results.f$wb.error=three.mod.results.f$wb.med.tkd-three.mod.results.f$true.tKD.med
three.mod.results.f$rz.error=three.mod.results.f$rz.med.tkd-three.mod.results.f$true.tKD.med
sum(abs(three.mod.results.f$jorg.error))
sum(abs(three.mod.results.f$wb.error))
sum(abs(three.mod.results.f$rz.error))
sum((three.mod.results.f$jorg.error)^2)
sum((three.mod.results.f$wb.error)^2)
sum((three.mod.results.f$rz.error)^2)
```

```{r}
#males
#exp 1
exp1.m.summ = jorg.exp1.m$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
#linear shape
exp1.m.summ$wb.med.tkd=c(wb.e1.g1.m$med.tdt,wb.e1.g2.m$med.tdt,wb.e1.g3.m$med.tdt,wb.e1.g4.m$med.tdt)
exp1.m.summ$wb.mean.tkd=c(wb.e1.g1.m$expected.tdt,wb.e1.g2.m$expected.tdt,wb.e1.g3.m$expected.tdt,wb.e1.g4.m$expected.tdt)
#constant shape
exp1.m.summ$wb.cons.med.tkd=c(wb.cons.e1.g1.m$med.tdt,wb.cons.e1.g2.m$med.tdt,wb.cons.e1.g3.m$med.tdt,wb.cons.e1.g4.m$med.tdt)
exp1.m.summ$wb.cons.mean.tkd=c(wb.cons.e1.g1.m$expected.tdt,wb.cons.e1.g2.m$expected.tdt,wb.cons.e1.g3.m$expected.tdt,wb.cons.e1.g4.m$expected.tdt)

exp1.m.summ$rz.med.tkd=c(rz.med.e1.g1.m,rz.med.e1.g2.m,rz.med.e1.g3.m,rz.med.e1.g4.m)
exp1.m.summ$exp=1

#exp 2
exp2.m.summ = jorg.exp2.m$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
#linear shape
exp2.m.summ$wb.med.tkd=c(wb.e2.g1.m$med.tdt,wb.e2.g2.m$med.tdt,wb.e2.g3.m$med.tdt,wb.e2.g4.m$med.tdt,wb.e2.g5.m$med.tdt)
exp2.m.summ$wb.mean.tkd=c(wb.e2.g1.m$expected.tdt,wb.e2.g2.m$expected.tdt,wb.e2.g3.m$expected.tdt,wb.e2.g4.m$expected.tdt,wb.e2.g5.m$med.tdt)
#constant shape
exp2.m.summ$wb.cons.med.tkd=c(wb.cons.e2.g1.m$med.tdt,wb.cons.e2.g2.m$med.tdt,wb.cons.e2.g3.m$med.tdt,wb.cons.e2.g4.m$med.tdt,wb.cons.e2.g5.m$med.tdt)
exp2.m.summ$wb.cons.mean.tkd=c(wb.cons.e2.g1.m$expected.tdt,wb.cons.e2.g2.m$expected.tdt,wb.cons.e2.g3.m$expected.tdt,wb.cons.e2.g4.m$expected.tdt,wb.cons.e2.g5.m$med.tdt)
exp2.m.summ$rz.med.tkd=c(rz.med.e2.g1.m,rz.med.e2.g2.m,rz.med.e2.g3.m,rz.med.e2.g4.m,rz.med.e2.g5.m)
exp2.m.summ$exp=2

#exp 3
exp3.m.summ = jorg.exp3.m$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
#linear shape
exp3.m.summ$wb.med.tkd=c(wb.e3.g1.m$med.tdt,wb.e3.g2.m$med.tdt,wb.e3.g3.m$med.tdt,wb.e3.g4.m$med.tdt)
exp3.m.summ$wb.mean.tkd=c(wb.e3.g1.m$expected.tdt,wb.e3.g2.m$expected.tdt,wb.e3.g3.m$expected.tdt,wb.e3.g4.m$expected.tdt)
#constant shape
exp3.m.summ$wb.cons.med.tkd=c(wb.cons.e3.g1.m$med.tdt,wb.cons.e3.g2.m$med.tdt,wb.cons.e3.g3.m$med.tdt,wb.cons.e3.g4.m$med.tdt)
exp3.m.summ$wb.cons.mean.tkd=c(wb.cons.e3.g1.m$expected.tdt,wb.cons.e3.g2.m$expected.tdt,wb.cons.e3.g3.m$expected.tdt,wb.cons.e3.g4.m$expected.tdt)

exp3.m.summ$rz.med.tkd=c(rz.med.e3.g1.m,rz.med.e3.g2.m,rz.med.e3.g3.m,rz.med.e3.g4.m)
exp3.m.summ$exp=3

three.mod.results.m = rbind(exp1.m.summ,exp2.m.summ,exp3.m.summ)
three.mod.results.m$sex="m"
#note that later I will calculate error based on individual points, not medians. That will be better, but this will give us an idea for now.
three.mod.results.m$jorg.error=three.mod.results.m$est.tKD-three.mod.results.m$true.tKD.med
three.mod.results.m$wb.error=three.mod.results.m$wb.med.tkd-three.mod.results.m$true.tKD.med
three.mod.results.m$rz.error=three.mod.results.m$rz.med.tkd-three.mod.results.m$true.tKD.med
sum(abs(three.mod.results.m$jorg.error))
sum(abs(three.mod.results.m$wb.error))
sum(abs(three.mod.results.m$rz.error))
sum((three.mod.results.m$jorg.error)^2)
sum((three.mod.results.m$wb.error)^2)
sum((three.mod.results.m$rz.error)^2)
```

```{r}
#combine data
three.mod.results = rbind(three.mod.results.f[,colnames(three.mod.results.f)%in%colnames(three.mod.results.m)],three.mod.results.m)
```


#make plot (Fig 4)
```{r}
fig4.males = ggplot(data = three.mod.results.m) +
  geom_point(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Weibull"), size = 3) +
  geom_point(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), size = 3) +
  geom_point(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Weibull"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), method = "lm", se = FALSE) +
  labs(
    title = "B) Males",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("#440154FF", "#22A884FF", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
  
```

```{r}
fig4.females = ggplot(data = three.mod.results.f) +
  geom_point(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Weibull"), size = 3) +
  geom_point(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), size = 3) +
  geom_point(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Weibull"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), method = "lm", se = FALSE) +
  labs(
    title = "A) Females",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("#440154FF", "#22A884FF", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
  
```

```{r}
# Remove individual axis titles and legends from each plot
fig4.females2 <- fig4.females + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")

fig4.males2 <- fig4.males + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")

# Create shared axis label plots as empty ggplots with just text
x_lab_plot <- ggplot() + 
  theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = expression("Observed Median t"[f] * " (minutes)"), size = 6)

y_lab_plot <- ggplot() + 
  theme_void() +
  annotate("text", x = 0.5, y = 0.5, label = expression("Predicted Median t"[f] * " (minutes)"), size = 6, angle = 90)

#Combine the two plots with patchwork
main_plots <- wrap_plots(fig4.females2, fig4.males2, ncol = 2) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")  

#Assemble everything into one patchwork layout with shared axis labels

final_plot <- (
  y_lab_plot + main_plots + plot_spacer() +    # 1st row: y label, plots, spacer
  plot_spacer() + x_lab_plot + plot_spacer()   # 2nd row: spacer, x label, spacer
) +
  plot_layout(
    ncol = 3,
    nrow = 2,
    widths = c(0.06, 0.88, 0.06),
    heights = c(0.9, 0.1)
  )

print(final_plot)
```


#Make supplemental figure (S2) comparing Constant and Varying Shape Weibull Models
```{r}
ggplot(data = three.mod.results.f) +
  geom_point(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Weibull"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(y = log10(wb.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Weibull"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Weibull"), method = "lm", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(wb.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Weibull"), method = "lm", se = FALSE) +
  labs(
    title = "Females",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("gray", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
```

```{r}
ggplot(data = three.mod.results.m) +
  geom_point(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Weibull"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(y = log10(wb.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Weibull"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(wb.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Weibull"), method = "lm", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(wb.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Weibull"), method = "lm", se = FALSE) +
  labs(
    title = "Males",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("gray", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
```

