---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
rm(list = ls())
#make figure 1
```

```{r}
library(flexsurv)
library(tidyverse)
library(patchwork)
library(viridis)
library(survival)
library(survminer)
```


```{r}
set.seed(13)
#generate temperatures

temps = seq(32,40,0.1)
times = seq(0,500,0.1)
exp.log.fail = -0.2*temps+9
lin.shapes = -0.144*temps+7
cons.shape.low = lin.shapes[temps==32]
cons.shape.high = lin.shapes[temps==40]
#use the equation var(log(x))=pi^2/3*shape^2 for x~llogis
#note these are ln not log10
var.low=pi^2/(3*cons.shape.low^2)
var.high=pi^2/(3*cons.shape.high^2)
var.med=mean(c(var.low,var.high))
cons.shape.med = sqrt(pi^2/(3*var.med))
scales.lin = 10^(exp.log.fail)
scales.cons = 10^(exp.log.fail)
#calculate var(log(tf)) for each temp for both lin increasing and constant shapes
lin.vars = c()
cons.vars.low = c()
cons.vars.med = c()
cons.vars.high = c()
#exp.vars = c()
for (i in 1: length(temps)){
  temp_i = temps[i]
  preds.lin_i = rllogis(1000000, shape=lin.shapes[temps==temp_i],scale=scales.lin[temps==temp_i])
  preds.cons.low_i = rllogis(1000000, shape=cons.shape.low,scale=scales.cons[temps==temp_i])
  preds.cons.med_i = rllogis(1000000, shape=cons.shape.med,scale=scales.cons[temps==temp_i])
  preds.cons.high_i = rllogis(1000000, shape=cons.shape.high,scale=scales.cons[temps==temp_i])
  lin.vars = c(lin.vars,var(log10(preds.lin_i)))
  cons.vars.low = c(cons.vars.low,var(log10(preds.cons.low_i)))
  cons.vars.med = c(cons.vars.med,var(log10(preds.cons.med_i)))
  cons.vars.high = c(cons.vars.high,var(log10(preds.cons.high_i)))
}
```

```{r}
logx_vals <- seq(-1, 4, length.out = 300)   


grid <- expand.grid(
  temp = temps,
  logx = logx_vals
)


grid$X <- 10^grid$logx


grid$shape <- lin.shapes[match(grid$temp, temps)]
grid$scale <- scales.lin[match(grid$temp, temps)]

# Log-logistic PDF

grid$density <- with(grid,
  (shape/scale) * ( (X/scale)^(shape - 1) ) /
           (1 + (X/scale)^shape)^2
)

# Heatmap plot
p <- ggplot(data=NULL) +
  geom_tile(data=grid, aes(x = temp, y = logx, fill = density)) +
  scale_fill_viridis_c(option = "mako",trans="log") +
  labs(
    x = "Temperature",
    y = "log10(Failure time)",
    fill = "Density"
  ) +
  geom_line(aes(x=temps,y=exp.log.fail),linewidth=1.25,color="gray",show.legend=FALSE)+
  ylab(expression(log[10](t[f]~(min))))+
  xlab("Temperature (C)")+
  theme_classic()+
  theme(legend.text=element_blank())

#add quantiles
q_lower = qllogis(0.025, shape = lin.shapes, scale = scales.lin)
q_upper = qllogis(0.975, shape = lin.shapes, scale = scales.lin)
logq_lower = log10(q_lower)
logq_upper = log10(q_upper)
quantiles_df <- data.frame(
  temp = temps,
  lower = log10(qllogis(0.025, shape = lin.shapes, scale = scales.lin)),
  upper = log10(qllogis(0.975, shape = lin.shapes, scale = scales.lin))
)

p=p+
  geom_line(aes(x = quantiles_df$temp, y = quantiles_df$lower), color = "white", linetype = "dotted", linewidth = 1.5) +
  geom_line(aes(x = quantiles_df$temp, y = quantiles_df$upper), color = "white", linetype = "dotted", linewidth = 1.5)+
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))+
  ylim(c(-0.5,3.5))

```

```{r}
colors <- viridis_pal(option = "mako")(7)
set.seed(11)
#generate points from linear and constant shape cases for p1
cons.shape.points = rep(NA,length(temps))
lin.shape.points = rep(NA,length(temps))
for (i in 1:length(temps)){
  temp_i = temps[i]
  lin.shape.points[i] = rllogis(1,shape=lin.shapes[i],scale=scales.lin[i])
}
  
p2 = ggplot(data=NULL)+
  geom_line(aes(x=temps,y=cons.vars.low,color="Low Var"),lty="solid",linewidth=1.25,show.legend=FALSE)+
  geom_line(aes(x=temps,y=cons.vars.med,color="Medium Var"),lty="solid",linewidth=1.25,show.legend=FALSE)+
  geom_line(aes(x=temps,y=cons.vars.high,color="High Var"),lty="solid",linewidth=1.25,show.legend=FALSE)+
    geom_line(aes(x=temps,y=lin.vars,color="Increasing Var"),lty="dashed",linewidth=1.25,show.legend=FALSE)+
  ylab(expression(var(log[10](t[f]~(min)))))+
  xlab("Temperature (C)")+
    theme_classic()+
  scale_color_manual(
    name = NULL,  
    breaks = c("Increasing Var","Low Var","Medium Var","High Var"),
    values = c(
      "Increasing Var" = colors[1],  
      "Low Var" = colors[4],
      "Medium Var" = colors[5],
      "High Var" = colors[6]))+
  xlim(c(32,40))+
  ylim(c(0.05,0.45))+
  # scale_color_viridis_d(option = "mako", name = NULL)+ 
  theme(legend.position = "none")


pC = ggplot(data=NULL)+
  geom_density(aes(x=rllogis(100000, shape=cons.shape.low,scale=scales.cons[temps==32]),lty="Low Var",color="Low Var"),linewidth=1.5)+
  geom_density(aes(x=rllogis(100000, shape=cons.shape.med,scale=scales.cons[temps==32]),lty="Medium Var",color="Medium Var"),linewidth=1.5)+
  geom_density(aes(x=rllogis(100000, shape=cons.shape.high,scale=scales.cons[temps==32]),lty="High Var",color="High Var"),linewidth=1.5)+
  geom_density(aes(x=rllogis(100000, shape=lin.shapes[temps==32],scale=scales.lin[temps==32]),lty="Increasing Var",color="Increasing Var"),linewidth=1.5)+
  xlab(expression(t[f]~"(minutes)"))+
  ylab("Density")+
    theme_classic()+
  guides(lty="none")+
  scale_color_manual(
    name = NULL, 
    breaks = c("Increasing Var","Low Var","Medium Var","High Var"),
    values = c(
      "Increasing Var" = colors[1],  
      "Low Var" = colors[4],
      "Medium Var" = colors[5],
      "High Var" = colors[6]))+
  scale_linetype_manual(
    name = NULL,
    breaks = c("Increasing Var","Low Var","Medium Var","High Var"),
    values = c(
      "Increasing Var" = "dashed",
      "Low Var" = "solid",
      "Medium Var" = "solid",
      "High Var" = "solid")) +
  guides(
    color = guide_legend(override.aes = list(linetype = c("dashed", "solid", "solid", "solid"))),
    linetype = "none")+
  xlim(c(0,2e3))+
  ggtitle("                      32 C")

pD = ggplot(data=NULL)+
  geom_density(aes(x=rllogis(100000, shape=cons.shape.low,scale=scales.cons[temps==40]),lty="Low Var",color="Low Var"),linewidth=1.5,show.legend=FALSE)+
  geom_density(aes(x=rllogis(100000, shape=cons.shape.med,scale=scales.cons[temps==40]),lty="Medium Var",color="Medium Var"),linewidth=1.5,show.legend=FALSE)+
  geom_density(aes(x=rllogis(100000, shape=cons.shape.high,scale=scales.cons[temps==40]),lty="High Var",color="High Var"),linewidth=1.5,show.legend=FALSE)+
  geom_density(aes(x=rllogis(100000, shape=lin.shapes[temps==40],scale=scales.lin[temps==40]),lty="Increasing Var",color="Increasing Var"),linewidth=1.5,show.legend=FALSE)+
  xlab(expression(t[f]~"(minutes)"))+
  ylab("Density")+
    theme_classic()+
  guides(lty="none")+
  scale_color_manual(
    name = NULL,  
    breaks = c("Increasing Var","Low Var","Medium Var","High Var"),
    values = c(
      "Increasing Var" = colors[1],  
      "Low Var" = colors[4],
      "Medium Var" = colors[5],
      "High Var" = colors[6]))+
  scale_linetype_manual(
    name = NULL,
    breaks = c("Increasing Var","Low Var","Medium Var","High Var"),
    values = c(
      "Increasing Var" = "dashed",
      "Low Var" = "solid",
      "Medium Var" = "solid",
      "High Var" = "solid")) +
  guides(
    color = guide_legend(override.aes = list(linetype = c("dashed", "solid", "solid", "solid"))))+
  xlim(c(0,3e1))+
  ggtitle("                      40 C")



# Plot using the same color/linetype mapping
ll.surv.funct = function(x,shape,scale){
  Fx=1-(1/(1+(x/scale)^shape))
  Sx=1-Fx
  return(Sx)}

pE = ggplot(data=NULL)+
  xlim(c(0,2e3))+
  geom_function(color=colors[5],linetype="solid",linewidth=1.5,fun=ll.surv.funct,args=list(shape=cons.shape.med,scale=scales.cons[temps == 32]))+
  geom_function(color=colors[6],linetype="solid",linewidth=1.5,fun=ll.surv.funct,args=list(shape=cons.shape.high,scale=scales.cons[temps == 32]))+
  geom_function(color=colors[4],linetype="solid",linewidth=2,fun=ll.surv.funct,args=list(shape=cons.shape.low,scale=scales.cons[temps == 32]))+
    geom_function(color=colors[1],linetype="dashed",linewidth=1.5,fun=ll.surv.funct,args=list(shape=lin.shapes[temps == 32],scale=scales.lin[temps == 32]))+
    theme_classic()+
  ylab("Survival")+
  xlab("Time (minutes)")+
  ggtitle("                      32 C")

pF = ggplot(data=NULL)+
  xlim(c(0,3e1))+
  geom_function(color=colors[5],linetype="solid",linewidth=1.5,fun=ll.surv.funct,args=list(shape=cons.shape.med,scale=scales.cons[temps == 40]))+
  geom_function(color=colors[4],linetype="solid",linewidth=1.5,fun=ll.surv.funct,args=list(shape=cons.shape.low,scale=scales.cons[temps == 40]))+
  geom_function(color=colors[6],linetype="solid",linewidth=2,fun=ll.surv.funct,args=list(shape=cons.shape.high,scale=scales.cons[temps == 40]))+
    geom_function(color=colors[1],linetype="dashed",linewidth=1.5,fun=ll.surv.funct,args=list(shape=lin.shapes[temps == 40],scale=scales.lin[temps == 40]))+
    theme_classic()+
  ylab("Survival")+
  xlab("Time (minutes)")+
  ggtitle("                      40 C")

wrap_plots(p,p2,pC,pD,pE,pF,ncol=2)+
  plot_layout(guides="collect")&
  theme(legend.position = "bottom")& 
  plot_annotation(tag_levels = "A")
```

```{r}
rm(list = ls())
library(flexsurv)
library(tidyverse)
library(viridis)
library(patchwork)
#create new supplemental figure
#Goal: illustrate effects of log-logistic shape parameter on A) Failure Density, B) Cumulative Survival, C) Hazard function.  Plot these vs Time (min) on linear scale.
#Use one value of scale (e.g. median failure time = 100 minutes), and different lines for value of shape:  0.5, 1, 2, 3, 4 (or something similar, 5-6 lines per plot). Use Mako color scale for (solid) line colors?

man.scale = 100
man.shapes = c(0.5, 1, 2, 3, 4)
colors <- viridis_pal(option = "mako")(5)

ll.dens.funct = function(x,shape,scale){
  a=shape
  b=scale
  fx=((a/b)*(x/b)^(a-1))/((1+(x/b)^a)^2)
  return(fx)}

ll.surv.funct = function(x,shape,scale){
  Fx=1-(1/(1+(x/scale)^shape))
  Sx=1-Fx
  return(Sx)}

ll.haz.funct = function(x,shape,scale){
  a=shape
  b=scale
  hx=((a/b)*(x/b)^(a-1))/(1+(x/b)^a)
  return(hx)}

sup.p1 = ggplot(data=NULL)+
  xlim(0,200)+
  geom_function(fun=ll.dens.funct,
                args=list(shape=man.shapes[1],scale=man.scale),
                aes(color = factor(man.shapes[1])),
                linewidth=1.5)+
  geom_function(fun=ll.dens.funct,
                args=list(shape=man.shapes[2],scale=man.scale),
                aes(color = factor(man.shapes[2])),
                linewidth=1.5)+
  geom_function(fun=ll.dens.funct,
                args=list(shape=man.shapes[3],scale=man.scale),
                aes(color = factor(man.shapes[3])),
                linewidth=1.5)+
  geom_function(fun=ll.dens.funct,
                args=list(shape=man.shapes[4],scale=man.scale),
                aes(color = factor(man.shapes[4])),
                linewidth=1.5)+
  geom_function(fun=ll.dens.funct,
                args=list(shape=man.shapes[5],scale=man.scale),
                aes(color = factor(man.shapes[5])),
                linewidth=1.5)+
  scale_color_manual(
    name = "Shape",
    values = colors,
    labels = man.shapes
  ) +
  theme_classic() +
  ylab("Failure Density") +
  xlab("") +
  theme(
    legend.position = c(0.95, 0.95),
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill="white", color="black")
  )

sup.p2 = ggplot(data=NULL)+
  xlim(0,200)+
  geom_function(fun=ll.surv.funct,args=list(shape=man.shapes[1],scale=man.scale),color=colors[1],linewidth=1.5)+
  geom_function(fun=ll.surv.funct,args=list(shape=man.shapes[2],scale=man.scale),color=colors[2],linewidth=1.5)+
  geom_function(fun=ll.surv.funct,args=list(shape=man.shapes[3],scale=man.scale),color=colors[3],linewidth=1.5)+
  geom_function(fun=ll.surv.funct,args=list(shape=man.shapes[4],scale=man.scale),color=colors[4],linewidth=1.5)+
  geom_function(fun=ll.surv.funct,args=list(shape=man.shapes[5],scale=man.scale),color=colors[5],linewidth=1.5)+
  theme_classic()+
  ylab("Cumulative Survival")+
  xlab("")

sup.p3 = ggplot(data=NULL)+
  xlim(0,200)+
  geom_function(fun=ll.haz.funct,args=list(shape=man.shapes[1],scale=man.scale),color=colors[1],linewidth=1.5)+
  geom_function(fun=ll.haz.funct,args=list(shape=man.shapes[2],scale=man.scale),color=colors[2],linewidth=1.5)+
  geom_function(fun=ll.haz.funct,args=list(shape=man.shapes[3],scale=man.scale),color=colors[3],linewidth=1.5)+
  geom_function(fun=ll.haz.funct,args=list(shape=man.shapes[4],scale=man.scale),color=colors[4],linewidth=1.5)+
  geom_function(fun=ll.haz.funct,args=list(shape=man.shapes[5],scale=man.scale),color=colors[5],linewidth=1.5)+
  theme_classic()+
  ylab("Hazard (1/min)")+
  xlab("Time (minutes)")

wrap_plots(sup.p1,sup.p2,sup.p3,nrow=3)
```

```{r}
rm(list = ls())
#make figure 2
```

```{r}
library(flexsurv)
library(tidyverse)
library(gridExtra)
library(patchwork)
library(RColorBrewer)
library(lme4)
```


```{r}
#Jorgenson et al. data
jorg = read.csv("Jorgensen_Static.csv")
jorg$event= rep(1,nrow(jorg))
colnames(jorg)[colnames(jorg)=="KDTime"] = "time"
colnames(jorg)[colnames(jorg)=="Temp"] = "RearTemp"
#see what species have enough data points to be usable
table(jorg$Species)
#see how many temperatures are covered by each species
spc.temp.tab = table(jorg$RearTemp,jorg$Species)
spc.temp.tab
#most of the species have enough data to be somewhat useful. But, some species/temp combinations do not. So I will remove any trial without at least 10 replicates. For trials with less data, flexsurv models typically can't converge. But if I instead excluded only trials for which the model wouldn't converge, I would be biasing the dataset.
trials.df = as.data.frame(spc.temp.tab>=10)
```

```{r}
sci.names = c("D. buzzatii","D. equinoxialis","D. immigrans","D. melanogaster","D. mercatorum", "D. mojavensis", "D. montana", "D. rufa", "D. subobscura", "D. suzukii", "D. virilis")
```

```{r}
#redoing this analysis for loglogistic
shape.list = list()
shape.var.list = list()
scale.list = list()
scale.var.list = list()
var.log.tkd.list = list()
tkd.list = list()

for (i in 1:length(colnames(trials.df))){
  species = colnames(trials.df)[i]
#narrow down to this particular species
species.data = jorg[jorg$Species==species,]
temp.list = rownames(trials.df)[trials.df[,species]]
#narrow down to only include trials with enough data
species.data = species.data[species.data$RearTemp %in% temp.list,]
#extract loglogistic model parameters
shape.species = c()
shape.species.var = c()
#IMPORTANT: this is log(scale), because flexsurvreg automatically puts scale on a log scale
scale.species = c()
scale.species.var = c()

var.log.tkd.species = c()
tkd.species = c()

#make loglogistic fits for every temperature for this species. Append its shape and scale parameters to the vectors above
for (j in unique(species.data$RearTemp)){
mod.j = flexsurvreg(Surv(time,event) ~ 1, data = species.data[(species.data$RearTemp==j),], dist=
                               "llogis")
shape.species.var = c(shape.species.var, mod.j$cov[[1,1]])
scale.species.var = c(scale.species.var, mod.j$cov[[2,2]])
shape.species = c(shape.species,mod.j$coefficients[[1]])
scale.species = c(scale.species,mod.j$coefficients[[2]])

var.log.tkd.species = c(var.log.tkd.species,var(log10(species.data[(species.data$RearTemp==j),]$time)))
tkd.species = c(tkd.species, median(species.data[(species.data$RearTemp==j),]$time))

}
shape.list[[length(shape.list)+1]] = shape.species
scale.var.list[[length(scale.var.list)+1]] = shape.species.var
scale.list[[length(scale.list)+1]] = scale.species
shape.var.list[[length(shape.var.list)+1]] = scale.species.var
var.log.tkd.list[[length(var.log.tkd.list)+1]] = var.log.tkd.species
tkd.list[[length(tkd.list)+1]] = tkd.species
}
```



```{r}
#make dataframe for species combined line plot
ll.pars.df = data.frame(species=rep(NA,0),temp=rep(NA,0),shape=rep(NA,0),scale=rep(NA,0),var.log.tKD=rep(NA,0))

for (i in 1:length(sci.names)){
  species = colnames(trials.df)[i]
  temp.list = as.numeric(rownames(trials.df)[trials.df[,species]])
  shapes=shape.list[[i]]
  scales=exp(scale.list[[i]])
  n.trials=length(shapes)
  spec.df=data.frame(species=rep(species,n.trials),temp=temp.list,shape=shapes,scale=scales,var.log.tKD=var.log.tkd.list[[i]],median.tKD=tkd.list[[i]])
  ll.pars.df=rbind(ll.pars.df,spec.df)
  
}
```

```{r}
#get modeled value of sCTmax (as defined by Jorgensen et al 2021) for each species
#these will be used to color code and order the species in the figure
sCTmax.df = data.frame(species=rep(NA,length(sci.names)),sCTmax=rep(NA,length(sci.names)))
for (i in 1:length(sci.names)){
  species = unique(ll.pars.df$species)[i]
  temp.list = as.numeric(rownames(trials.df)[trials.df[,i]])
  tdt.mod = lm(log10(median.tKD)~temp,data=ll.pars.df[ll.pars.df$species==species,])
  lm.slope = tdt.mod$coefficients[2]
  lm.int = tdt.mod$coefficients[1]
  sCT_max = (1-lm.int)/lm.slope
  sCTmax.df[i,]=c(species,sCT_max)
  
}

#enter scientific names
sci.names.key=data.frame(sci.names=sci.names,abbrev=c("Buz","Equ","Imm","Mel","Mer","Moj","Mon","Ruf","Sub","Suz","Vir"))
new.names.vec=rep(NA,nrow(ll.pars.df))
for (i in 1:nrow(ll.pars.df)){
new.names.vec[i]=sci.names.key$sci.names[sci.names.key$abbrev==ll.pars.df$species[i]]
  
}
ll.pars.df$species=new.names.vec

ll.pars.df$species = factor(ll.pars.df$species, levels=unique(ll.pars.df$species)[order(sCTmax.df$sCTmax, decreasing=TRUE)])
```



```{r}
#create tdt plot
tdt.plot = ggplot(data=ll.pars.df)+
  geom_line(aes(y=log10(median.tKD),x=temp,color=species),linewidth=1)+
  geom_hline(yintercept=1,linetype="dotted")+
  xlab("")+
  ylab(expression(log[10](t[f])))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species")

shape.plot = ggplot(data=ll.pars.df)+
  geom_line(aes(y=shape,x=temp,color=species),linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab("Shape")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species")

scale.plot = ggplot(data=ll.pars.df)+
  geom_line(aes(y=log10(scale),x=temp,color=species),linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(log[10]("Scale")))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species")

varlogtkd.plot = ggplot(data=ll.pars.df)+
  geom_line(aes(y=var.log.tKD,x=temp,color=species),linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(var(log[10](t[f]))))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.text.x=element_blank(),axis.title.x=element_blank(),legend.position="bottom")+
  scale_color_brewer(palette="RdYlBu")+
  labs(color="Species") 

combined_plot <- wrap_plots(tdt.plot, varlogtkd.plot, scale.plot, shape.plot)+
  plot_layout(guides = 'collect')&
  theme(legend.position="bottom",
        legend.key.size = unit(0.3, "cm"),
    legend.key = element_rect(color = NA, fill = NA),
    legend.text = element_text(face = "italic")) & 
  plot_annotation(tag_levels = "A")
  
  
  combined_plot
```

Create mixed effects model for loglogistic parameters by species
```{r}
null.shape.mod = lmer(shape~1+(1|species), data=ll.pars.df, REML=FALSE)
mixed.shape.mod = lmer(shape~temp+(1|species),data=ll.pars.df, REML=FALSE)
mixed.shape.mod.REML = lmer(shape~temp+(1|species),data=ll.pars.df, REML=TRUE)
# null.log10scale.mod = lmer(scale~1+(1|species), data=ll.pars.df,REML=FALSE)
# mixed.log10scale.mod = lmer(scale~temp+(1|species),data=ll.pars.df, REML=FALSE)
summary(mixed.shape.mod.REML)
anova(null.shape.mod,mixed.shape.mod)
```


supplemental plot recreating fig 2 but with each species separately
Fig S3


```{r}
#iterate through each species
combined.plots=list()

for (i in 1: length(unique(ll.pars.df$species))){
  species_i = unique(ll.pars.df$species)[i]
  color_i = brewer.pal(n=11,name="RdYlBu")[which.max(levels(ll.pars.df$species)==species_i)]
  print(color_i)
  data_i = ll.pars.df[ll.pars.df$species==species_i,]
  
tdt.plot = ggplot(data=data_i)+
  geom_point(aes(y=log10(median.tKD),x=temp),color=color_i)+
  geom_hline(yintercept=1,linetype="dotted")+
  geom_smooth(aes(y=log10(median.tKD),x=temp),color=color_i,linewidth=1,method="lm",se=F)+
  xlab("")+
  ylab(expression(log[10](t[f])))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")


shape.mod_i = lm(shape~temp,data=data_i) 
shape.p.value_i = summary(shape.mod_i)$coefficients[2,4]

shape.plot = ggplot(data=data_i)+
  geom_point(aes(y=shape,x=temp),color=color_i)+
  geom_smooth(aes(y=shape,x=temp),color=color_i,linewidth=1,method="lm",se=F)+
  xlab("Temperature (\u00B0C)")+
  ylab("Shape")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(),legend.position="bottom")+
  annotate("text", 
           x = Inf, y = Inf, 
           label = paste0("p = ", signif(shape.p.value_i, 3)),
           hjust = 1.1, vjust = 1.5,
           size = 4) 

scale.plot = ggplot(data=data_i)+
  geom_point(aes(y=scale,x=temp),color=color_i)+
  geom_smooth(aes(y=scale,x=temp),color=color_i,linewidth=1,method="lm",se=F)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(log[10]("Scale")))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),axis.title.x = element_blank(), axis.text.x = element_blank(),legend.position="bottom")

varlogtkd.plot = ggplot(data=data_i)+
  geom_line(aes(y=var.log.tKD,x=temp),color=color_i,linewidth=1)+
  xlab("Temperature (\u00B0C)")+
  ylab(expression(var(log[10](t[f]))))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"),legend.position="bottom") 

combined_plot <- wrap_plots(tdt.plot, scale.plot, varlogtkd.plot, shape.plot)+
  plot_layout(guides = 'collect')&
  theme(legend.position="none",
        legend.key.size = unit(0.3, "cm"),
    legend.key = element_rect(color = NA, fill = NA),
    legend.text = element_text(face = "italic")) & 
  plot_annotation(tag_levels = "A")&
  ggtitle(species_i)
  
  
  print(combined_plot)
  
  # Ensure the output folder exists
if (!dir.exists("suppfig3")) dir.create("suppfig3")

# Build file name safely
filename <- paste0("suppfig3/", gsub(" ", "_", species_i), ".jpeg")

# Save the plot
ggsave(filename, plot = combined_plot, width = 10, height = 8, units = "in", dpi = 300)
combined.plots[[i]] = combined_plot
}
```

```{r}
fig.s3.a = wrap_plots(combined.plots[[1]],combined.plots[[2]],combined.plots[[3]],combined.plots[[4]],combined.plots[[5]],combined.plots[[6]])
fig.s3.a
fig.s3.b = wrap_plots(combined.plots[[7]],combined.plots[[8]],combined.plots[[9]],combined.plots[[10]],combined.plots[[11]])
fig.s3.b
```

supplemental figure S2 residual plots
```{r}
#create a list to hold the residual plots for each species
plot.list = list()
#iterate through all species, create a residual plot for each, and append to the list
for (i in 1:length(unique(ll.pars.df$species))){
  species_i = unique(ll.pars.df$species)[i]
  data_i = ll.pars.df[ll.pars.df$species==species_i,]
  tdt.mod_i = lm(log10(median.tKD)~temp,data=data_i)
  tdt.slope_i = tdt.mod_i$coefficients[[2]]
  tdt.int_i = tdt.mod_i$coefficients[[1]]
  resids_i = log10(data_i$median.tKD)-(tdt.slope_i*data_i$temp+tdt.int_i)
  resid_df <- data.frame(
  temp = data_i$temp,
  resids = resids_i
)

plot_i <- ggplot(resid_df, aes(x = temp, y = resids)) +
  geom_point() +
  ggtitle(species_i) +
  ylab(expression("Residual "*log[10]*" "*t[f])) +
  xlab("Temperature (\u00B0C)")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(color = "black"))

  plot.list[[i]]=plot_i
}
```

```{r}
combined_resid_plot <- wrap_plots(plot.list[[1]],plot.list[[2]],plot.list[[3]],plot.list[[4]],plot.list[[5]],plot.list[[6]],plot.list[[7]],plot.list[[8]],plot.list[[9]],plot.list[[10]],plot.list[[11]])

combined_resid_plot
```

```{r}
rm(list = ls())
#make figure 3
```

```{r}
library(stats)
library(tidyverse)
library(eha)
library(gridExtra)
library(flexsurv)
library(survival)
library(ggsurvfit)
library(RColorBrewer)
library(survminer)
library(ggplotify)
library(patchwork)
library(viridis)
```
First, I'll run the functions to produce Rezende, LogLogistic models. Rezende code obtained from Rezende et al 2020, slightly modified to account for different sampling rates.

```{r}
#Steps to this model:
#1) First fit log-logistic curves to each tdt curve (each temperature) for the data supplied
#2) Then fit a linear model to both log-logistic shape and scale vs temp
#3) Next, we deal with the fluctuating temp data:
#     start cumulative survival (S(t)) at 1 (S(t_0)=1)
#     for each timestep t_i:
#         Find the log-logistic shape and scale parameters corresponding to the temperature T_i at timestep t_i. These parameters give a cumulative survival function of time (time spent under the constant temperature used to make the LLogis TDT curve, which we will call t*. Not the fluctuating temperature t_i). This distribution has CDF F_w(t*), PDF f_w(t*), cumulative survival S_w(t*)=1-F_w(t*)
#         Find chance of surviving from time t_i-1 to t_i: 
#             i) First, we find: S_w^-1(S(t_i-1))=t*_i
#             ii) Calculate failure in that time interval f.rate_i=S(t_i)/S(t_i-1)
#             iii) Calculate cumulative mortality between t_i-1 and t_i as (t_i-(t_i-1))*f.rate_i
#             iiii)Let S_i=(S_i-1)-[(t_i-(t_i-1))*f_w(t*_i)]
#Note that this loop assumes the temperature is constant at T_i for the time interval from t_i-1 to t_i. You could alternatively take the max or average of the temperatures at these time points. 
#replace with shape="constant" if you want the model to assume constant shape
llogis.mod = function(TDT.dat,temp.dat,shape="linear"){
  #step 1
shape.vec = c()
shape.vec.var = c()
scale.vec = c()
scale.vec.var = c()
for (j in unique(TDT.dat$Tref)){
mod.j = flexsurvreg(Surv(time,event) ~ 1, data = TDT.dat[(TDT.dat$Tref==j),], dist=
                               "llogis")
shape.vec.var = c(shape.vec.var, mod.j$cov[[1,1]])
scale.vec.var = c(scale.vec.var, exp(mod.j$cov[[2,2]]))
shape.vec = c(shape.vec,mod.j$coefficients[[1]])
#exponentiate because the function flexsurvreg automatically puts scale on a log scale
scale.vec = c(scale.vec,exp(mod.j$coefficients[[2]]))
}

#step 2
shape.lm=lm(shape.vec~unique(TDT.dat$Tref))
shape.slope=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[2]]
shape.int=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[1]]
#constant shape lm
cons.shape.lm=lm(shape.vec~1)
cons.shape.int=cons.shape.lm$coefficients[[1]]
#note the log transformation
scale.lm=lm(log10(scale.vec)~unique(TDT.dat$Tref))
scale.slope=scale.lm$coefficients[[2]]
scale.int=scale.lm$coefficients[[1]]

shape.pars=c(shape.slope,shape.int,cons.shape.int)
logscale.pars=c(scale.slope,scale.int)

#predict under fluctuating temperatures with shape as a linear function of temperature
if (shape=="linear"){
#step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  shape_i = shape.slope*mean(T_i,T_prev)+shape.int
  mean.list=c(mean.list,((pi*scale_i)/shape_i)/(sin(pi/shape_i)))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  prob = 1-cum.surv[i-1]
  t_star_i=scale_i*(prob/(1-prob))^(1/shape_i)
  #llogis pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pllogis(t_star_i,shape_i,scale_i))-(1-pllogis(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)]))
}

if (shape=="constant"){
 #step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  #this is the only line that is different
  shape_i = cons.shape.int
  mean.list=c(mean.list,((pi*scale_i)/shape_i)/(sin(pi/shape_i)))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  prob = 1-cum.surv[i-1]
  t_star_i=scale_i*(prob/(1-prob))^(1/shape_i)
  #llogis pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pllogis(t_star_i,shape_i,scale_i))-(1-pllogis(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)])) 
  
  
  
}

output=list(med.tdt,expected.tdt,cum.surv.post,shape.pars,logscale.pars)
names(output)=c("med.tdt","expected.tdt","cum.surv","shape.pars","logscale.pars")
return(output)
}
```



#3 Rezende et al model

```{r}
#First function produces static "tolerance landscape" from a vector of temperatures and a vector of thermal death times


	tolerance.landscape <- function(ta,time){
	
			data <- data.frame(ta,time)
			data <- data[order(data$ta,data$time),]

			# Step 1: Calculate CTmax and z from TDT curve
			ta <- as.numeric(levels(as.factor(data$ta)))			
			model <- lm(log10(data$time) ~ data$ta); summary(model)
			ctmax <- -coef(model)[1]/coef(model)[2]
			z <- -1/coef(model)[2]

			# Step 2: Calculate average log10 time and Ta (mean x and y for interpolation purposes)
			time.mn <- mean(log10(data$time))
			ta.mn <- mean(data$ta)
	
			# Step 3: Interpolating survival probabilities to make them comparable across treatments 
			time.interpol <- matrix(,1001,length(ta))
			for(i in 1:length(ta)){	
			time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(0,100,length.out = length(time))
			time.interpol[,i] <- approx(p,time,n = 1001)$y}			

			# Step 4: Overlap all survival curves into a single one by shifting each curve to mean x and y employing z
			# Step 5: Build expected survival curve with median survival time for each survival probability
			shift <- (10^((ta - ta.mn)/z))
			time.interpol.shift <- t(t(time.interpol)*shift)[-1,]
			surv.pred <- 10^apply(log10(time.interpol.shift),1,median) 	
	
			# Step 6: Expand predicted survival curves to measured Ta (matrix m arranged from lower to higher ta)
			# Step 7: Obtain predicted values comparable to each empirical measurement
			m <- surv.pred*matrix ((10^((ta.mn - rep(ta, each = 1000))/z)), nrow = 1000)
			out <-0
			for(i in 1:length(ta)){
				time <- c(0,data$time[data$ta==ta[i]]); p <- seq(0,100,length.out = length(time))
				out <- c(out,approx(seq(0,100,length.out = 1000),m[,i],xout=p[-1])$y)}
				data$time.pred <- out[-1]
				colnames(m) <- paste("time.at",ta,sep=".")
				m <- cbind(surv.prob=seq(1,0.001,-0.001),m)

			par(mfrow=c(1,2),mar=c(4.5,4,1,1),cex.axis=1.1)
			plot(-10,-10,las=1,xlab="Time (min)",ylab="Survival (%)",col="white",xaxs="i",yaxs="i",xlim=c(0,max(data$time)*1.05),ylim=c(0,105))
				for(i in 1:length(ta)){
				time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(100,0,length.out = length(time))
				points(time,p,pch=21,bg="black",cex=0.5)
				time <- c(0,sort(data$time.pred[data$ta==ta[i]]))
				points(m[,i+1],100*m[,1],type="l",lty=2)}
				segments(max(data$time)*0.7,90,max(data$time)*0.8,90,lty=2)
				text(max(data$time)*0.82,90,"fitted",adj=c(0,0.5))
			plot(log10(data$time.pred),log10(data$time),pch=21,bg="black",cex=0.5,lwd=0.7,las=1,xlab="Fitted Log10 time",ylab="Measured Log10 time")
			abline(0,1,lty=2)
			rsq <- round(summary(lm(log10(data$time) ~ log10(data$time.pred)))$r.square,3)
			text(min(log10(data$time.pred)),max(log10(data$time)),substitute("r"^2*" = "*rsq),adj=c(0,1))
			list(ctmax = as.numeric(ctmax), z = as.numeric(z), ta.mn = ta.mn,  S = data.frame(surv=seq(0.999,0,-0.001),time=surv.pred),
			time.obs.pred=cbind(data$time,data$time.pred), rsq = rsq)}	
```

```{r}
#ta is vector of fluctuating temperature conditions
#times is a vector of the corresponding times
	dynamic.landscape <- function(ta,times,tolerance.landscape){
	  #These are the predicted survival times that correspond to set cumulative survival values for the average temperature
			surv <- tolerance.landscape$S[,2]
			#mean temperature from TDT curves
			ta.mn <- tolerance.landscape$ta.mn
			#thermal tolerance coefficient
			z <- tolerance.landscape$z
			#the change in cumulative survival for a given temperature
			shift <- 10^((ta.mn - ta)/z)	
			time.rel <- 0
			#this will become a vector of cumulative survival percentages
			alive <- 100
			for(i in 1:(length(ta)-1)){	
			  #this is the size of the timestep in question
			  tstep=times[i+1]-times[i]
				if(alive[length(na.omit(alive))] > 0) {		
				  #approx fits a linear regression of relative time 
					alive <- try(c(alive,approx(c(0,shift[i]*surv),seq(100,0,length.out = length(c(0,surv))),xout = time.rel[i] + tstep)$y),silent=TRUE)
					#This essentially uses the inverse CDF of the shifted survival curve to approximate the comparable static temperature time using the next temperature
					time.rel <- try(c(time.rel,approx(seq(100,0,length.out = length(c(0,surv))),c(0,shift[i + 1]*surv),xout = alive[i + 1])$y),silent=TRUE)}
				else{
					alive <- 0}}				
			out <- data.frame(cbind(ta=ta[1:(length(alive)-1)],time=times[1:(length(alive)-1)],alive=alive[1:(length(alive)-1)]))
			par(mar=c(4,4,1,1),mfrow=c(1,2))
			plot(1:length(ta),ta,type="l",xlim=c(0,length(ta)),ylim=c(min(ta),max(ta)),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Temperature (ºC)")			
			plot(out$time,out$alive,type="l",xlim=c(0,length(ta)),ylim=c(0,100),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Survival (%)")
			list(time = out$time,ta = out$ta, alive = out$alive)}
```

Function to plot survival curves

```{r}
#This function takes data from the above models and turns it into a survival curve plot
#exp.data=experimental data
#data1->results from model 1 
plot.surv.curves = function(data1,data2,exp.data,modtitle1,modtitle2,title,xlim){
# Function to convert a vector into a data frame with group label
make_df <- function(vec, group_name) {
  data.frame(time = vec, status = 1, group = group_name)
}

# Combine all data
df <- bind_rows(
  make_df(data1, modtitle1),
  make_df(data2, modtitle2),
  make_df(exp.data, "Data")
)

# Set group order: modtitle1, modtitle2, then "Data"
df$group <- factor(df$group, levels = c(modtitle1, modtitle2, "Data"))

# Fit KM model
km_fit <- survfit(Surv(time, status) ~ group, data = df)
tidy_km <- broom::tidy(km_fit)

# Clean and reorder strata names for matching
tidy_km$strata <- case_match(tidy_km$strata,
                         paste0("group=",modtitle1) ~ modtitle1,
                         paste0("group=",modtitle2)  ~ modtitle2,
                         "group=Data" ~ "Data")

# Set factor order for legend
tidy_km$strata <- factor(tidy_km$strata, levels = c(modtitle1, modtitle2, "Data"))

p = ggplot(tidy_km, aes(x = time, y = estimate, color = strata, linetype = strata)) +
  geom_step(linewidth = 1.6) +
  labs(
    y = "Cumulative Survival",
    x = "Time (minutes)"
  ) +
  theme_minimal(base_size = 14) +
  theme_classic()+
  theme(
    panel.grid = element_blank(),
    legend.title = element_blank()
  )+
  ggtitle(title)+
  xlim(0,xlim)+
  scale_color_manual(values=c("Increasing Variance"="#440154FF","Rezende Model"="darkorange","Data"="gray"))
return(p)}
```

Estimating the MEAN log likelihood of the data given the simulated tf distribution using a kernel density estimate, then testing how likely the data was given the distribution by computing a p-value via bootstrapping. Mean log-liklihood allows for standardization across differing sample sizes
```{r}
log_likelihood_test <- function(sims, data, n_boot = 10000, max_tries = 10, density_floor = 1e-10) {
  # Initial bandwidth
  base_bw <- bw.nrd0(sims)
  bw_factor <- 1
  attempt <- 1
  found_valid <- FALSE
  
  # Try increasing bandwidth until all data points get nonzero density
  while (attempt <= max_tries && !found_valid) {
    bw <- bw_factor * base_bw
    kde <- density(sims, bw = bw, n = 2048)
    kde_fn <- approxfun(kde$x, kde$y, rule = 2)
    
    densities_data <- kde_fn(data)
    
    if (all(densities_data > 0)) {
      found_valid <- TRUE
    } else {
      bw_factor <- bw_factor * 1.5
      attempt <- attempt + 1
    }
  }
  
  if (!found_valid) {
    warning("Log-likelihoods may be unstable: some data points still have zero estimated density after maximum bandwidth increase.")
  }

  # Apply floor to prevent log(0)
  densities_data <- pmax(densities_data, density_floor)
  mean_log_lik_data <- mean(log(densities_data))

  # Bootstrap: sample and compute log-likelihoods
  ll_dens <- replicate(n_boot, {
    sample_data <- sample(sims, size = length(data), replace = TRUE)
    sample_densities <- pmax(kde_fn(sample_data), density_floor)
    mean(log(sample_densities))
  })

  p_val <- mean(ll_dens <= mean_log_lik_data)

  return(list(
    ll.dens = ll_dens,
    ll.data = mean_log_lik_data,
    p_value = p_val,
    bandwidth_used = bw
  ))
}
```

```{r}
#Jorgenson et al. data
jorg = read.csv("Jorgensen_Static.csv")
jorg$event= rep(1,nrow(jorg))
colnames(jorg)[colnames(jorg)=="KDTime"] = "time"
colnames(jorg)[colnames(jorg)=="Temp"] = "RearTemp"
#see what species have enough data points to be usable
table(jorg$Species)
#see how many temperatures are covered by each species
spc.temp.tab = table(jorg$RearTemp,jorg$Species)
spc.temp.tab
#most of the species have enough data to be somewhat useful. But, some species/temp combinations do not. So I will remove any trial without at least 10 replicates. For trials with less data, models typically can't converge. But if I instead excluded only trials for which the model wouldn't converge, I would be biasing the dataset.
trials.df = as.data.frame(spc.temp.tab>=10)
```

```{r}
sci.names = c("D. buzzatii","D. equinoxialis","D. immigrans","D. melanogaster","D. mercatorum", "D. mojavensis", "D. montana", "D. rufa", "D. subobscura", "D. suzukii", "D. virilis")
```


```{r}
surv.fits.list=list()
rz.fits.list=list()
#loglogistic with changing shape
ll.fits.list=list()
#for every species
for (i in 1:length(colnames(trials.df))){
  species = colnames(trials.df)[i]
#narrow down to this particular species
species.data = jorg[jorg$Species==species,]
temp.list = rownames(trials.df)[trials.df[,species]]
#narrow down to only include trials with enough data
species.data = species.data[species.data$RearTemp %in% temp.list,]
#km survival fit to the data
surv.fit = survfit(Surv(time, event) ~ as.numeric(RearTemp), data = species.data)
fit.summary = summary(surv.fit)
times = fit.summary$time
survs=fit.summary$surv
strats=fit.summary$strata
temps=sapply(strats, function(x) sub("^as\\.numeric\\(RearTemp\\)=", "", x))
#make cumulative survival curves out of raw data
surv.fits.list[[i]]=data.frame(times=times,cum.surv=survs,temp=as.numeric(temps))
names(surv.fits.list)[i]=species

#Rezende fits
tol.land.spec = tolerance.landscape(species.data$RearTemp,species.data$time)

ntemp=length(unique(species.data$RearTemp))
             
rz.cum.surv=rep(tol.land.spec$S[,1],ntemp)
rz.surv.times=rep(tol.land.spec$S[,2],ntemp)
rz.temp=c()
rz.times=c()
#run models from Rezende et al 2020
for (j in 1:ntemp){
  n.per.temp=length(tol.land.spec$S[,1])
  rz.temp[c((1+(j-1)*n.per.temp):(j*n.per.temp))] = rep(unique(species.data$RearTemp)[j],n.per.temp)
}

ta.mn = tol.land.spec$ta.mn
z = tol.land.spec$z


rz.times = rz.surv.times*10^((ta.mn - rz.temp)/z)

rz.fits.list[[i]] = data.frame(times=rz.times,cum.surv=rz.cum.surv,temp=as.numeric(rz.temp))
names(rz.fits.list)[i]=species


#failure density by temp
#failure rates by temp
rz.fits.list[[i]]$fail.rates = rep(0,nrow(rz.fits.list[[i]]))
for (j in 1:ntemp){
  temp_j=unique(rz.fits.list[[i]]$temp)[j]
  cum.surv_j = rz.fits.list[[i]]$cum.surv[rz.fits.list[[i]]$temp==temp_j]
  times_j = rz.fits.list[[i]]$times[rz.fits.list[[i]]$temp==temp_j]
density_j = c(-(cum.surv_j[2:length(cum.surv_j)]-cum.surv_j[1:(length(cum.surv_j)-1)]))
delta_t_j= c(times_j[2:length(times_j)]-times_j[1:(length(times_j)-1)])
rz.fits.list[[i]]$fail.rates[rz.fits.list[[i]]$temp==temp_j]=c(NA,density_j/delta_t_j)
}

#Log-Logistic fits
species.data$Tref=species.data$RearTemp
#temp dat doesn't matter: it is fluctuating temp conditions, but we only care about the parameters used to make them. I've included filler here.
ll.spec = llogis.mod(temp.dat=data.frame(Time_min=rep(1,10),Temperature=rep(35,10)),TDT.dat=species.data)
#cut off times as end of last rezende or survival fit, unless manually overwritten, just for graphical clarity
man.t.max = 25000
t.max = max(c(rz.times,times,man.t.max))
ll.cum.surv=rep(tol.land.spec$S[,1],ntemp)
ll.surv.times=rep(tol.land.spec$S[,2],ntemp)
ll.temp=c()
ll.times=c()

for (j in 1:ntemp){
  time.seq = seq(0,t.max,0.1)
  n.times=length(time.seq)
  ll.times[(1+(j-1)*n.times):(j*n.times)] = time.seq
  
  ll.temp[(1+(j-1)*n.times):(j*n.times)] = rep(unique(species.data$RearTemp)[j],n.times)
}

#fit loglogistic distribution parameters as a linear function of temperature
shapes.vec = ll.spec$shape.pars[1]*ll.temp+ll.spec$shape.pars[2]
scales.vec = 10^(ll.spec$logscale.pars[1]*ll.temp+ll.spec$logscale.pars[2])

for (j in 1:nrow(species.data)){
  #failtime for individual j
  x_j = species.data$time[j]
  #temp of individual j
  temp_j = species.data$Tref[j]
  lin.shape_j = ll.spec$shape.pars[1]*temp_j+ll.spec$shape.pars[2]
  scale_j = 10^(ll.spec$logscale.pars[1]*temp_j+ll.spec$logscale.pars[2])
}

#cumulative survival
ll.cum.surv = 1-pllogis(ll.times,shape=shapes.vec,scale=scales.vec)

ll.fits.list[[i]] = data.frame(times=ll.times,cum.surv=ll.cum.surv,temp=as.numeric(ll.temp),shapes=shapes.vec,scales=scales.vec)
names(ll.fits.list)[i]=species


#failure rates by temp
ll.fits.list[[i]]$density = rep(0,nrow(ll.fits.list[[i]]))
for (j in 1:ntemp){
  temp_j=unique(ll.fits.list[[i]]$temp)[j]
  cum.surv_j = ll.fits.list[[i]]$cum.surv[ll.fits.list[[i]]$temp==temp_j]
  times_j = ll.fits.list[[i]]$times[ll.fits.list[[i]]$temp==temp_j]
density_j = c(-(cum.surv_j[2:length(cum.surv_j)]-cum.surv_j[1:(length(cum.surv_j)-1)]))
delta_t_j= c(times_j[2:length(times_j)]-times_j[1:(length(times_j)-1)])
ll.fits.list[[i]]$fail.rates[ll.fits.list[[i]]$temp==temp_j]=c(NA,density_j/delta_t_j)
}


}
```



#density plots
Need to calculate density. But this is difficult given that some of our curves are discrete and some are continuous. To make matters simpler and put all on an even playing field, I'll estimate densities by simulation. I'll generate 10000 unif(0,1) random variables, and treat these as cumulative survival values. Then, I'll map them to associated failure time values, giving me simulated random variables with the density of the model in question.
```{r}
#this function takes a dataframe with one column "times", a second column "cum.surv", and a third column "temp". These should be ordered in increasing time/decreasing cumulative survival. n is the number of random data points generated.
generate.density = function(cdf,n){
  
  temps.list = unique(cdf$temp)
  fail.ests = data.frame(
    tf = rep(NA,0),
    temp = rep(NA,0),
    rv = rep(NA,0))
    rvs = runif(n)
  
for (i in 1:length(temps.list)){
  temp_i = temps.list[i]
  cdf_i = cdf[cdf$temp==temp_i,]
#vector of n unif(0,1) random variables
#vector that will contain all of the simulated failure times
fail.ests.i = rep(NA,n)
for (j in 1:n){
  #the first time at which cumulative survival is less than or equal to our simulated random variable.
  #rounding our time up like this is practical, because if data is collected every dt minutes, and failure occurs just after time t, it won't be detected until time t+dt.
  fail.ests.i[j] = cdf_i$times[cdf_i$cum.surv<=rvs[j]][1]
}
#set an arbitrarily large value to the simulated points that are outside of our graph window. This way they will still be accounted for in the density metric, but won't show up on the graph.
fail.ests.i[is.na(fail.ests.i)==TRUE]=100000
fail.ests=rbind(fail.ests,data.frame(tf=fail.ests.i,temp=rep(temp_i,n),rv=rvs))
}
return(fail.ests)
}
```

```{r}
temp.list=seq(33,41,0.5)
common_color_scale=scale_color_manual(name = "Temperature", values = plasma(length(temp.list)),breaks=as.character(temp.list),limits=as.character(temp.list))

common.linetype.scale=scale_linetype_manual(values=c("Rezende"="dotted","Log-Logistic"="dashed","Data"="solid"))

data.a=surv.fits.list$Mel[surv.fits.list$Mel$temp%in%c("36","41"),]
rz.a = rz.fits.list$Mel[rz.fits.list$Mel$temp%in%c("36","41"),]
ll.a = ll.fits.list$Mel[ll.fits.list$Mel$temp%in%c("36","41"),]

data.dens.a=generate.density(
  cdf = data.a[,c("times","cum.surv","temp")],n=10000)
rz.dens.a = generate.density(
  cdf = rz.a[,c("times","cum.surv","temp")],n=10000)
ll.dens.a = generate.density(
  cdf = ll.a[,c("times","cum.surv","temp")],n=10000)

data.b=surv.fits.list$Sub[surv.fits.list$Sub$temp%in%c("33","40"),]
rz.b = rz.fits.list$Sub[rz.fits.list$Sub$temp%in%c("33","40"),]
ll.b = ll.fits.list$Sub[ll.fits.list$Sub$temp%in%c("33","40"),]
data.dens.b=generate.density(
  cdf = data.b[,c("times","cum.surv","temp")],n=10000)
rz.dens.b = generate.density(
  cdf = rz.b[,c("times","cum.surv","temp")],n=10000)
ll.dens.b = generate.density(
  cdf = ll.b[,c("times","cum.surv","temp")],n=10000)

data.c=surv.fits.list$Moj[surv.fits.list$Moj$temp%in%c("39.5","43.5"),]
rz.c = rz.fits.list$Moj[rz.fits.list$Moj$temp%in%c("39.5","43.5"),]
ll.c = ll.fits.list$Moj[ll.fits.list$Moj$temp%in%c("39.5","43.5"),]
data.dens.c=generate.density(
  cdf = data.c[,c("times","cum.surv","temp")],n=10000)
rz.dens.c = generate.density(
  cdf = rz.c[,c("times","cum.surv","temp")],n=10000)
ll.dens.c = generate.density(
  cdf = ll.c[,c("times","cum.surv","temp")],n=10000)

data.d=surv.fits.list$Vir[surv.fits.list$Vir$temp%in%c("37.5","41.5"),]
rz.d = rz.fits.list$Vir[rz.fits.list$Vir$temp%in%c("37.5","41.5"),]
ll.d = ll.fits.list$Vir[ll.fits.list$Vir$temp%in%c("37.5","41.5"),]

data.dens.d=generate.density(
  cdf = data.d[,c("times","cum.surv","temp")],n=10000)
rz.dens.d = generate.density(
  cdf = rz.d[,c("times","cum.surv","temp")],n=10000)
ll.dens.d = generate.density(
  cdf = ll.d[,c("times","cum.surv","temp")],n=10000)

data.e=surv.fits.list$Buz[surv.fits.list$Buz$temp%in%c("38.5","43"),]
rz.e = rz.fits.list$Buz[rz.fits.list$Buz$temp%in%c("38.5","43"),]
ll.e = ll.fits.list$Buz[ll.fits.list$Buz$temp%in%c("38.5","43"),]

data.dens.e=generate.density(
  cdf = data.e[,c("times","cum.surv","temp")],n=10000)
rz.dens.e = generate.density(
  cdf = rz.e[,c("times","cum.surv","temp")],n=10000)
ll.dens.e = generate.density(
  cdf = ll.e[,c("times","cum.surv","temp")],n=10000)

data.f=surv.fits.list$Equ[surv.fits.list$Equ$temp%in%c("34","40"),]
rz.f = rz.fits.list$Equ[rz.fits.list$Equ$temp%in%c("34","40"),]
ll.f = ll.fits.list$Equ[ll.fits.list$Equ$temp%in%c("34","40"),]

data.dens.f=generate.density(
  cdf = data.f[,c("times","cum.surv","temp")],n=10000)
rz.dens.f = generate.density(
  cdf = rz.f[,c("times","cum.surv","temp")],n=10000)
ll.dens.f = generate.density(
  cdf = ll.f[,c("times","cum.surv","temp")],n=10000)

data.g=surv.fits.list$Imm[surv.fits.list$Imm$temp%in%c("33.5","38.5"),]
rz.g = rz.fits.list$Imm[rz.fits.list$Imm$temp%in%c("33.5","38.5"),]
ll.g = ll.fits.list$Imm[ll.fits.list$Imm$temp%in%c("33.5","38.5"),]

data.dens.g=generate.density(
  cdf = data.g[,c("times","cum.surv","temp")],n=10000)
rz.dens.g = generate.density(
  cdf = rz.g[,c("times","cum.surv","temp")],n=10000)
ll.dens.g = generate.density(
  cdf = ll.g[,c("times","cum.surv","temp")],n=10000)

data.h=surv.fits.list$Mer[surv.fits.list$Mer$temp%in%c("35.5","40.5"),]
rz.h = rz.fits.list$Mer[rz.fits.list$Mer$temp%in%c("35.5","40.5"),]
ll.h = ll.fits.list$Mer[ll.fits.list$Mer$temp%in%c("35.5","40.5"),]

data.dens.h=generate.density(
  cdf = data.h[,c("times","cum.surv","temp")],n=10000)
rz.dens.h = generate.density(
  cdf = rz.h[,c("times","cum.surv","temp")],n=10000)
ll.dens.h = generate.density(
  cdf = ll.h[,c("times","cum.surv","temp")],n=10000)

data.i=surv.fits.list$Mon[surv.fits.list$Mon$temp%in%c("36","39.5"),]
rz.i = rz.fits.list$Mon[rz.fits.list$Mon$temp%in%c("36","39.5"),]
ll.i = ll.fits.list$Mon[ll.fits.list$Mon$temp%in%c("36","39.5"),]

data.dens.i=generate.density(
  cdf = data.i[,c("times","cum.surv","temp")],n=10000)
rz.dens.i = generate.density(
  cdf = rz.i[,c("times","cum.surv","temp")],n=10000)
ll.dens.i = generate.density(
  cdf = ll.i[,c("times","cum.surv","temp")],n=10000)

data.j=surv.fits.list$Ruf[surv.fits.list$Ruf$temp%in%c("34","39.5"),]
rz.j = rz.fits.list$Ruf[rz.fits.list$Ruf$temp%in%c("34","39.5"),]
ll.j = ll.fits.list$Ruf[ll.fits.list$Ruf$temp%in%c("34","39.5"),]

data.dens.j=generate.density(
  cdf = data.j[,c("times","cum.surv","temp")],n=10000)
rz.dens.j = generate.density(
  cdf = rz.j[,c("times","cum.surv","temp")],n=10000)
ll.dens.j = generate.density(
  cdf = ll.j[,c("times","cum.surv","temp")],n=10000)

data.k=surv.fits.list$Suz[surv.fits.list$Suz$temp%in%c("35","39"),]
rz.k = rz.fits.list$Suz[rz.fits.list$Suz$temp%in%c("35","39"),]
ll.k = ll.fits.list$Suz[ll.fits.list$Suz$temp%in%c("35","39"),]

data.dens.k=generate.density(
  cdf = data.k[,c("times","cum.surv","temp")],n=10000)
rz.dens.k = generate.density(
  cdf = rz.k[,c("times","cum.surv","temp")],n=10000)
ll.dens.k = generate.density(
  cdf = ll.k[,c("times","cum.surv","temp")],n=10000)

```

```{r}
#subobscura
surv.plot.33 = plot.surv.curves(data1=ll.dens.b$tf[ll.dens.b$temp==33],data2=rz.dens.b$tf[rz.dens.b$temp==33],exp.data=data.dens.b$tf[data.dens.b$temp==33],modtitle1="Increasing Variance", modtitle2="Rezende Model", title="                    33 C",xlim=600)+ 
  theme(legend.title = element_blank())

surv.plot.40 = plot.surv.curves(data1=ll.dens.b$tf[ll.dens.b$temp==40],data2=rz.dens.b$tf[rz.dens.b$temp==40],exp.data=data.dens.b$tf[data.dens.b$temp==40],modtitle1="Increasing Variance", modtitle2="Rezende Model", title="                    40 C",xlim=10)+
  theme(legend.title = element_blank())

dens.plot.33 = ggplot(data=NULL)+
    geom_density(aes(x=log10(tf),group=as.factor(temp)),data=ll.dens.b[ll.dens.b$temp==33,],size=1,color="#440154FF")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=rz.dens.b[rz.dens.b$temp==33,],size=1,color="darkorange")+
  geom_rug(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.b[data.dens.b$temp==33,],sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.b[data.dens.b$temp==33,], color="gray")+
  xlim(2,3)+
  ylim(0,15)+
  ggtitle("                   33 C")+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))+
  theme(legend.position="none")+
  theme(legend.title = element_blank())+
  ylab("Density")+
  xlab(expression(log[10]*t[f]))

dens.plot.40 = ggplot(data=NULL)+
   geom_density(aes(x=log10(tf),group=as.factor(temp)),data=ll.dens.b[ll.dens.b$temp==40,],size=1,color="#440154FF")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=rz.dens.b[rz.dens.b$temp==40,],size=1,color="darkorange")+
  geom_rug(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.b[data.dens.b$temp==40,],sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.b[data.dens.b$temp==40,], color="gray")+
  ggtitle("                   40 C")+
  xlim(0,1)+
  ylim(0,15)+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))+
  theme(legend.position="none")+ 
  theme(legend.title = element_blank())+  
  ylab("Density")+
  xlab(expression(log[10]*t[f]))

#equinoxialis
surv.plot.34.f = plot.surv.curves(data1=ll.dens.f$tf[ll.dens.f$temp==34],data2=rz.dens.f$tf[rz.dens.f$temp==34],exp.data=data.dens.f$tf[data.dens.f$temp==34],modtitle1="Increasing Variance", modtitle2="Rezende Model", title="                    34 C",xlim=1500)+ 
  theme(legend.title = element_blank())

surv.plot.40.f = plot.surv.curves(data1=ll.dens.f$tf[ll.dens.f$temp==40],data2=rz.dens.f$tf[rz.dens.f$temp==40],exp.data=data.dens.f$tf[data.dens.f$temp==40],modtitle1="Increasing Variance", modtitle2="Rezende Model", title="                    40 C",xlim=15)+
  theme(legend.title = element_blank())

dens.plot.34.f = ggplot(data=NULL)+
    geom_density(aes(x=log10(tf),group=as.factor(temp)),data=ll.dens.f[ll.dens.f$temp==34,],size=1,color="#440154FF")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=rz.dens.f[rz.dens.f$temp==34,],size=1,color="darkorange")+
  geom_rug(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.f[data.dens.f$temp==34,],sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.f[data.dens.f$temp==34,], color="gray")+
  xlim(2.25,3.75)+
  ggtitle("                   34 C")+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))+
  theme(legend.position="none")+
  theme(legend.title = element_blank())+
  ylab("Density")+
  xlab(expression(log[10]*t[f]))

dens.plot.40.f = ggplot(data=NULL)+
   geom_density(aes(x=log10(tf),group=as.factor(temp)),data=ll.dens.f[ll.dens.f$temp==40,],size=1,color="#440154FF")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=rz.dens.f[rz.dens.f$temp==40,],size=1,color="darkorange")+
  geom_rug(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.f[data.dens.f$temp==40,],sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.f[data.dens.f$temp==40,], color="gray")+
  ggtitle("                   40 C")+
  xlim(0,1.5)+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))+
  theme(legend.position="none")+ 
  theme(legend.title = element_blank())+  
  ylab("Density")+
  xlab(expression(log[10]*t[f]))

heading_plot <- function(label) {
  ggplot() +
    annotate(
      "text",
      x = 0.5, y = 0.5,
      label = label,
      fontface = "italic",
      size = 5
    ) +
    theme_void()
}

head_sub <- ggplot() +
  annotate(
    "text", 0.5, 0.5,
    label = expression(italic("D. subobscura")),
    size = 5
  ) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(t = 6, b = 6)
  )

head_mel <- ggplot() +
  annotate(
    "text", 0.5, 0.5,
    label = expression(italic("D. equinoxialis")),
    size = 5
  ) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(t = 6, b = 6)
  )

dens.plot.33  <- dens.plot.33  + labs(tag = "A") + theme_classic()
dens.plot.40  <- dens.plot.40  + labs(tag = "B") + theme_classic()
surv.plot.33  <- surv.plot.33  + labs(tag = "C") + theme_classic()
surv.plot.40  <- surv.plot.40  + labs(tag = "D") + theme_classic()

dens.plot.34.f <- dens.plot.34.f + labs(tag = "E") + theme_classic()
dens.plot.40.f <- dens.plot.40.f + labs(tag = "F") + theme_classic()
surv.plot.34.f <- surv.plot.34.f + labs(tag = "G") + theme_classic()
surv.plot.40.f <- surv.plot.40.f + labs(tag = "H") + theme_classic()

sub_block <- wrap_plots(
  head_sub,        plot_spacer(),
  dens.plot.33,    dens.plot.40,
  surv.plot.33,    surv.plot.40,
  ncol = 2,
  heights = c(0.15, 1, 1)
)


mel_block <- wrap_plots(
  head_mel,        plot_spacer(),
  dens.plot.34.f,  dens.plot.40.f,
  surv.plot.34.f,  surv.plot.40.f,
  ncol = 2,
  heights = c(0.15, 1, 1)
)


final_plot <-
  wrap_plots(sub_block, mel_block, ncol = 1) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )

final_plot 
```

```{r}
#Let's make that generic so we can plot the others for fig S4
plot.stat.temps = function(high.temp,low.temp,ll.dens.df,rz.dens.df,data.dens.df,title,surv.max.x.high,surv.max.x.low,xlim.dens.high,xlim.dens.low,y.max.dens){
surv.plot.low = plot.surv.curves(data1=ll.dens.df$tf[ll.dens.df$temp==low.temp],data2=rz.dens.df$tf[rz.dens.df$temp==low.temp],exp.data=data.dens.df$tf[data.dens.df$temp==low.temp],modtitle1="Increasing Variance", modtitle2="Rezende Model", title=paste0("                          ",low.temp," C"),xlim=surv.max.x.low)+ 
  theme(legend.title = element_blank())

surv.plot.high = plot.surv.curves(data1=ll.dens.df$tf[ll.dens.df$temp==high.temp],data2=rz.dens.df$tf[rz.dens.df$temp==high.temp],exp.data=data.dens.df$tf[data.dens.df$temp==high.temp],modtitle1="Increasing Variance", modtitle2="Rezende Model", title=paste0("                          ",high.temp," C"),xlim=surv.max.x.high)+
  theme(legend.title = element_blank())

dens.plot.low = ggplot(data=NULL)+
    geom_density(aes(x=log10(tf),group=as.factor(temp)),data=ll.dens.df[ll.dens.df$temp==low.temp,],size=1,color="#440154FF")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=rz.dens.df[rz.dens.df$temp==low.temp,],size=1,color="darkorange")+
  geom_rug(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.df[data.dens.df$temp==low.temp,],sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.df[data.dens.df$temp==low.temp,], color="gray")+
  xlim(xlim.dens.low)+
  ggtitle(paste0("                          ",low.temp," C"))+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))+
  theme(legend.position="none")+
  theme(legend.title = element_blank())+
  ylab("Density")+
  xlab(expression(log[10]*t[f]))+
  theme_classic()+
  ylim(0,y.max.dens)

dens.plot.high = ggplot(data=NULL)+
   geom_density(aes(x=log10(tf),group=as.factor(temp)),data=ll.dens.df[ll.dens.df$temp==high.temp,],size=1,color="#440154FF")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=rz.dens.df[rz.dens.df$temp==high.temp,],size=1,color="darkorange")+
  geom_rug(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.df[data.dens.df$temp==high.temp,],sides="b")+
  geom_density(aes(x=log10(tf),group=as.factor(temp)),data=data.dens.df[data.dens.df$temp==high.temp,], color="gray")+
  ggtitle(paste0("                          ",high.temp," C"))+
  xlim(xlim.dens.high)+
  common.linetype.scale+
  theme(axis.title = element_blank(),plot.title = element_text(face = "italic"))+
  theme(legend.position="none")+ 
  theme(legend.title = element_blank())+  
  ylab("Density")+
  xlab(expression(log[10]*t[f]))+
  theme_classic()+
  ylim(0,y.max.dens)

head_spc <- ggplot() +
  annotate(
    "text", 0.5, 0.5,
    label = title,
    size = 5
  ) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(t = 6, b = 6)
  )

block <- wrap_plots(
  head_spc,        plot_spacer(),
  dens.plot.low,  dens.plot.high,
  surv.plot.low,  surv.plot.high,
  ncol = 2,
  heights = c(0.15, 1, 1)
)+
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.title = element_blank())

print(block)}

```

```{r}
#Moj 39.5, 43.5
plot.stat.temps(high.temp="43.5",low.temp="39.5",ll.dens.df=ll.dens.c,rz.dens.df=rz.dens.c,data.dens.df=data.dens.c,title=expression(italic("D. mojavensis")),
                surv.max.x.high=10,surv.max.x.low=600,xlim.dens.high=c(0,1),xlim.dens.low=c(2,3),y.max.dens=15)
#Vir "37.5","41.5"
plot.stat.temps(high.temp="41.5",low.temp="37.5",ll.dens.df=ll.dens.d,rz.dens.df=rz.dens.d,data.dens.df=data.dens.d,title=expression(italic("D. virilis")),
                surv.max.x.high=10,surv.max.x.low=600,xlim.dens.high=c(0,1),xlim.dens.low=c(2,3),y.max.dens=15)
#Buz "38.5","43"
plot.stat.temps(high.temp="43",low.temp="38.5",ll.dens.df=ll.dens.e,rz.dens.df=rz.dens.e,data.dens.df=data.dens.e,title=expression(italic("D. buzzatii")),
                surv.max.x.high=10,surv.max.x.low=600,xlim.dens.high=c(0,1),xlim.dens.low=c(2,3),y.max.dens=15)
#Mel
plot.stat.temps(high.temp="41",low.temp="36",ll.dens.df=ll.dens.a,rz.dens.df=rz.dens.a,data.dens.df=data.dens.a,title=expression(italic("D. melanogaster")),
                surv.max.x.high=30,surv.max.x.low=600,xlim.dens.high=c(0,2),xlim.dens.low=c(1.5,3.5),y.max.dens=15)
#Imm
plot.stat.temps(high.temp="38.5",low.temp="33.5",ll.dens.df=ll.dens.g,rz.dens.df=rz.dens.g,data.dens.df=data.dens.g,title=expression(italic("D. immigrans")),
                surv.max.x.high=10,surv.max.x.low=600,xlim.dens.high=c(0,1),xlim.dens.low=c(2,3),y.max.dens=15)
#Mer
plot.stat.temps(high.temp="40.5",low.temp="35.5",ll.dens.df=ll.dens.h,rz.dens.df=rz.dens.h,data.dens.df=data.dens.h,title=expression(italic("D. mercatorum")),
                surv.max.x.high=15,surv.max.x.low=400,xlim.dens.high=c(0,1.75),xlim.dens.low=c(2,2.75),y.max.dens=15)
#Mon
plot.stat.temps(high.temp="39.5",low.temp="36",ll.dens.df=ll.dens.i,rz.dens.df=rz.dens.i,data.dens.df=data.dens.i,title=expression(italic("D. montana")),
                surv.max.x.high=12.5,surv.max.x.low=600,xlim.dens.high=c(0,1.5),xlim.dens.low=c(1.75,3.25),y.max.dens=15)
#Ruf
plot.stat.temps(high.temp="39.5",low.temp="34",ll.dens.df=ll.dens.j,rz.dens.df=rz.dens.j,data.dens.df=data.dens.j,title=expression(italic("D. rufa")),
                surv.max.x.high=15,surv.max.x.low=600,xlim.dens.high=c(0,1.5),xlim.dens.low=c(1.75,3.25),y.max.dens=15)
#Suz
plot.stat.temps(high.temp="39",low.temp="35",ll.dens.df=ll.dens.k,rz.dens.df=rz.dens.k,data.dens.df=data.dens.k,title=expression(italic("D. suzukii")),
                surv.max.x.high=10,surv.max.x.low=600,xlim.dens.high=c(0,1.5),xlim.dens.low=c(1.75,3.25),y.max.dens=15)
```

Now we want to compute the mean log-likelihoods for increasing variance and Rezende models for each temperature and each species


```{r}
#We next want to extract log-likelihoods
#The problem is that the dynamic.landscape function returns a CDF (or really 1-CDF=survival curve), but the log likelihood function estimates a PDF through samples failure times. So we have to sample failure times through simulation and obtain a PDF, since it is not an analytical distribution.
Surv.to.tf.sims = function(Surv.df,n=10000){
  sims.vec=rep(NA,n)
  unifs = runif(n)*100
  for (i in 1:n){
    sims.vec[i] = Surv.df$time[which.max(Surv.df$alive<=unifs[i])]
  }
  return(sims.vec)
}
```
-Make an empty list that will contain dataframes for each species with columns for temperature, rz mean log-likelihood and ll mean log likelihood
For each species
for each unique temperature in that species

Step 0 refit survival models to data excluding the temperature of interest
Step 1 Surv.to.tf.sims for the static temperature log-logistic and rezende data
Step 2 log likelihood test function for the result of step 1
Step 3 Extract log likelihood and add to a dataframe for that species inside the list

```{r}
static.logl.list = list()
for (spec in colnames(trials.df)){
  
  temps = rownames(trials.df)[trials.df[,spec]]
  
  #this is the output df with loglikelihoods
  species.df = data.frame(species = rep(NA,length(temps)), temp=rep(NA,length(temps)), logl.rz = rep(NA,length(temps)), logl.ll = rep(NA,length(temps)))
  
  #this has failure data
  species.data = jorg[jorg$Species==spec,]
  temp.list = rownames(trials.df)[trials.df[,spec]]
#narrow down to only include trials with enough data
species.data = species.data[species.data$RearTemp %in% temp.list,]
  
  ll.fit.spec.df = as.data.frame(ll.fits.list[spec])[,c(1,2,3)]
  colnames(ll.fit.spec.df) = c("time","cum.surv","temp")
  ll.fit.spec.df$alive=ll.fit.spec.df$cum.surv*100
  
  data.spec.df = jorg[jorg$Species==spec,]
  
  counter = 1
  
  for (temp in temps){
    #we need to refit the models without the temp in question. Leave one out cross validation, basically.
    temp.data = species.data[species.data$RearTemp!=temp,]
    
       #Rezende fits
    tol.land.spec = tolerance.landscape(temp.data$RearTemp,temp.data$time)
                 
    rz.cum.surv=tol.land.spec$S[,1]
    rz.surv.times=tol.land.spec$S[,2]
    rz.temp=c()
    rz.times=c()
      n.per.temp=length(tol.land.spec$S[,1])
      #use the temp that was excluded from the fits
      rz.temp = rep(as.numeric(temp),n.per.temp)
    
    ta.mn = tol.land.spec$ta.mn
    z = tol.land.spec$z
    #use surv fit to create estimates for the temperature that wasn't included
    rz.times = rz.surv.times*10^((ta.mn - rz.temp)/z)
    
    rz.fit.temp = data.frame(times=rz.times,cum.surv=rz.cum.surv,temp=as.numeric(rz.temp))
    colnames(rz.fit.temp) = c("time","cum.surv","temp")
    rz.fit.temp$alive=rz.fit.temp$cum.surv*100
    
    #Log-Logistic fits
temp.data$Tref=temp.data$RearTemp
#temperature data doesn't matter: it is fluctuating temp conditions, but we only care about the parameters used to make them. I've included filler here.
#Once again fit using the data that doesn't include our temperature of interest
ll.spec = llogis.mod(temp.dat=data.frame(Time_min=rep(1,10),Temperature=rep(35,10)),TDT.dat=temp.data)
#cut off times
man.t.max = 25000
t.max = man.t.max
ll.cum.surv=tol.land.spec$S[,1]
ll.surv.times=tol.land.spec$S[,2]
ll.temp=c()
ll.times=c()

  time.seq = seq(0,t.max,0.1)
  n.times=length(time.seq)
  ll.times[(1+(j-1)*n.times):(j*n.times)] = time.seq
  
  ll.temp[(1+(j-1)*n.times):(j*n.times)] = rep(temp,n.times)

#cumulative survival
ll.cum.surv = 1-pllogis(ll.times,shape=shapes.vec,scale=scales.vec)

ll.fit.temp = data.frame(times=ll.times,cum.surv=ll.cum.surv,temp=as.numeric(ll.temp))
    colnames(ll.fit.temp) = c("time","cum.surv","temp")
  ll.fit.temp$alive=ll.fit.temp$cum.surv*100

#now use the fits to get simulated tf and calculate log likelihood
    data.temp = data.spec.df[data.spec.df$RearTemp==temp,]$time
    
    sims.temp.rz = Surv.to.tf.sims(rz.fit.temp)
    sims.temp.ll = Surv.to.tf.sims(ll.fit.temp)
    
    species.df$temp[counter] = temp
    species.df$logl.rz[counter] = log_likelihood_test(sims=sims.temp.rz,data=data.temp)$ll.data
    species.df$logl.ll[counter] = log_likelihood_test(sims=sims.temp.ll,data=data.temp)$ll.data
    species.df$species[counter] = spec
    counter=counter+1
    print(counter)
  }
  
  species.df$logl.diff = species.df$logl.ll-species.df$logl.rz
  static.logl.list[[length(static.logl.list)+1]] = species.df
}

names(static.logl.list) = colnames(trials.df)
```

```{r}
#reformat df
static.logl.long = data.frame(temp=NULL, logl.rz = NULL, logl.ll = NULL)
for (i in 1:length(static.logl.list)){
  static.logl.long = rbind(static.logl.long,static.logl.list[[i]])
}
```

```{r}
ggplot(data=static.logl.long)+
  geom_point(aes(y=logl.diff,x=temp,color=species))

ggplot(data=static.logl.long)+
  geom_histogram(aes(x=logl.diff))+
  geom_vline(xintercept=median(static.logl.long$logl.diff))
```

```{r}
#test to see if there is a relationship with temperature
library(lme4)
null.ll.mod = lmer(logl.diff~(1|species),data=static.logl.long,REML=FALSE)
ll.mod = lmer(logl.diff~temp+(1|species),data=static.logl.long,REML=FALSE)
anova(null.ll.mod,ll.mod)
#no non-linear trends in residuals. Variance is reasonable.
plot(ll.mod)
#no effect of temperature!
#check to see if one test is significantly better across all species/temps
t.test(static.logl.long$logl.diff, mu = 0)
```

```{r}
#Make Figure 4
rm(list = ls())
```

```{r}
library(tidyverse)
library(patchwork)
library(viridis)
library(flexsurv)
```

```{r}
#load functions for making LogLogistic and Rezende models
```

```{r}
#Steps to this model:
#1) First fit log-logistic curves to each tdt curve (each temperature) for the data supplied
#2) Then fit a linear model to both log-logistic shape and scale vs temp
#3) Next, we deal with the fluctuating temp data:
#     start cumulative survival (S(t)) at 1 (S(t_0)=1)
#     for each timestep t_i:
#         Find the log-logistic shape and scale parameters corresponding to the temperature T_i at timestep t_i. These parameters give a cumulative survival function of time (time spent under the constant temperature used to make the LLogis TDT curve, which we will call t*. Not the fluctuating temperature t_i). This distribution has CDF F_w(t*), PDF f_w(t*), cumulative survival S_w(t*)=1-F_w(t*)
#         Find chance of surviving from time t_i-1 to t_i: 
#             i) First, we find: S_w^-1(S(t_i-1))=t*_i
#             ii) Calculate failure in that time interval f.rate_i=S(t_i)/S(t_i-1)
#             iii) Calculate cumulative mortality between t_i-1 and t_i as (t_i-(t_i-1))*f.rate_i
#             iiii)Let S_i=(S_i-1)-[(t_i-(t_i-1))*f_w(t*_i)]
#Note that this loop assumes the temperature is constant at T_i for the time interval from t_i-1 to t_i. You could alternatively take the max or average of the temperatures at these time points. 
#replace with shape="constant" if you want the model to assume constant shape
llogis.mod = function(TDT.dat,temp.dat,shape="linear"){
  #step 1
shape.vec = c()
shape.vec.var = c()
scale.vec = c()
scale.vec.var = c()
for (j in unique(TDT.dat$Tref)){
mod.j = flexsurvreg(Surv(time,event) ~ 1, data = TDT.dat[(TDT.dat$Tref==j),], dist=
                               "llogis")
shape.vec.var = c(shape.vec.var, mod.j$cov[[1,1]])
scale.vec.var = c(scale.vec.var, exp(mod.j$cov[[2,2]]))
shape.vec = c(shape.vec,mod.j$coefficients[[1]])
#exponentiate because the function flexsurvreg automatically puts scale on a log scale
scale.vec = c(scale.vec,exp(mod.j$coefficients[[2]]))
}

#step 2
shape.lm=lm(shape.vec~unique(TDT.dat$Tref))
shape.slope=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[2]]
shape.int=lm(shape.vec~unique(TDT.dat$Tref))$coefficients[[1]]
#constant shape lm
cons.shape.lm=lm(shape.vec~1)
cons.shape.int=cons.shape.lm$coefficients[[1]]
#note the log transformation
scale.lm=lm(log10(scale.vec)~unique(TDT.dat$Tref))
scale.slope=scale.lm$coefficients[[2]]
scale.int=scale.lm$coefficients[[1]]

shape.pars=c(shape.slope,shape.int,cons.shape.int)
logscale.pars=c(scale.slope,scale.int)

#predict under fluctuating temperatures with shape as a linear function of temperature
if (shape=="linear"){
#step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  shape_i = shape.slope*mean(T_i,T_prev)+shape.int
  mean.list=c(mean.list,((pi*scale_i)/shape_i)/(sin(pi/shape_i)))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  prob = 1-cum.surv[i-1]
  t_star_i=scale_i*(prob/(1-prob))^(1/shape_i)
  #llogis pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pllogis(t_star_i,shape_i,scale_i))-(1-pllogis(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)]))
}

if (shape=="constant"){
 #step 3
cum.surv = rep(NA,nrow(temp.dat))
cum.surv[1]=1
fail.rates = rep(0,nrow(temp.dat))
mean.list=c()
shape.list=c()
log.list=c()

for (i in 2:nrow(temp.dat)){
  T_i=temp.dat$Temperature[i]
  T_prev=temp.dat$Temperature[i-1]
  t_i=temp.dat$Time_min[i]
  t_prev=temp.dat$Time_min[i-1]
  t_diff=t_i-t_prev
  #untransform from log scale
  scale_i = 10^(scale.slope*mean(T_i,T_prev)+scale.int)
  #this is the only line that is different
  shape_i = cons.shape.int
  mean.list=c(mean.list,((pi*scale_i)/shape_i)/(sin(pi/shape_i)))
  log.list=c(log.list,cum.surv[i-1])
  #inverse CDF
  prob = 1-cum.surv[i-1]
  t_star_i=scale_i*(prob/(1-prob))^(1/shape_i)
  #llogis pdf/failure rate
  #Get the proportion that made it th timestep i/prop that made it to i-1 to get the prop that died in the meantime
  fail.rates[i]=(1-pllogis(t_star_i,shape_i,scale_i))-(1-pllogis(t_star_i+t_diff,shape_i,scale_i))
  cum.surv[i]=cum.surv[i-1]-fail.rates[i]
}
med.tdt=temp.dat$Time_min[which.max(cum.surv<0.5)]
cum.surv.post=na.omit(cum.surv)
times.post=temp.dat$Time_min[1:length(cum.surv.post)]
expected.tdt=mean(-sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[2:length(cum.surv.post)]), -sum((cum.surv.post[2:length(cum.surv.post)]-cum.surv.post[1:(length(cum.surv.post)-1)])*times.post[1:(length(cum.surv.post)-1)])) 
  
  
  
}

output=list(med.tdt,expected.tdt,cum.surv.post,shape.pars,logscale.pars)
names(output)=c("med.tdt","expected.tdt","cum.surv","shape.pars","logscale.pars")
return(output)
}
```

Rezende et al model

```{r}
#First function produces static "tolerance landscape" from a vector of temperatures and a vector of thermal death times


	tolerance.landscape <- function(ta,time){
	
			data <- data.frame(ta,time)
			data <- data[order(data$ta,data$time),]

			# Step 1: Calculate CTmax and z from TDT curve
			ta <- as.numeric(levels(as.factor(data$ta)))			
			model <- lm(log10(data$time) ~ data$ta); summary(model)
			ctmax <- -coef(model)[1]/coef(model)[2]
			z <- -1/coef(model)[2]

			# Step 2: Calculate average log10 time and Ta (mean x and y for interpolation purposes)
			time.mn <- mean(log10(data$time))
			ta.mn <- mean(data$ta)
	
			# Step 3: Interpolating survival probabilities to make them comparable across treatments 
			time.interpol <- matrix(,1001,length(ta))
			for(i in 1:length(ta)){	
			time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(0,100,length.out = length(time))
			time.interpol[,i] <- approx(p,time,n = 1001)$y}			

			# Step 4: Overlap all survival curves into a single one by shifting each curve to mean x and y employing z
			# Step 5: Build expected survival curve with median survival time for each survival probability
			shift <- (10^((ta - ta.mn)/z))
			time.interpol.shift <- t(t(time.interpol)*shift)[-1,]
			surv.pred <- 10^apply(log10(time.interpol.shift),1,median) 	
	
			# Step 6: Expand predicted survival curves to measured Ta (matrix m arranged from lower to higher ta)
			# Step 7: Obtain predicted values comparable to each empirical measurement
			m <- surv.pred*matrix ((10^((ta.mn - rep(ta, each = 1000))/z)), nrow = 1000)
			out <-0
			for(i in 1:length(ta)){
				time <- c(0,data$time[data$ta==ta[i]]); p <- seq(0,100,length.out = length(time))
				out <- c(out,approx(seq(0,100,length.out = 1000),m[,i],xout=p[-1])$y)}
				data$time.pred <- out[-1]
				colnames(m) <- paste("time.at",ta,sep=".")
				m <- cbind(surv.prob=seq(1,0.001,-0.001),m)

			par(mfrow=c(1,2),mar=c(4.5,4,1,1),cex.axis=1.1)
			plot(-10,-10,las=1,xlab="Time (min)",ylab="Survival (%)",col="white",xaxs="i",yaxs="i",xlim=c(0,max(data$time)*1.05),ylim=c(0,105))
				for(i in 1:length(ta)){
				time <- c(0,sort(data$time[data$ta==ta[i]])); p <- seq(100,0,length.out = length(time))
				points(time,p,pch=21,bg="black",cex=0.5)
				time <- c(0,sort(data$time.pred[data$ta==ta[i]]))
				points(m[,i+1],100*m[,1],type="l",lty=2)}
				segments(max(data$time)*0.7,90,max(data$time)*0.8,90,lty=2)
				text(max(data$time)*0.82,90,"fitted",adj=c(0,0.5))
			plot(log10(data$time.pred),log10(data$time),pch=21,bg="black",cex=0.5,lwd=0.7,las=1,xlab="Fitted Log10 time",ylab="Measured Log10 time")
			abline(0,1,lty=2)
			rsq <- round(summary(lm(log10(data$time) ~ log10(data$time.pred)))$r.square,3)
			text(min(log10(data$time.pred)),max(log10(data$time)),substitute("r"^2*" = "*rsq),adj=c(0,1))
			list(ctmax = as.numeric(ctmax), z = as.numeric(z), ta.mn = ta.mn,  S = data.frame(surv=seq(0.999,0,-0.001),time=surv.pred),
			time.obs.pred=cbind(data$time,data$time.pred), rsq = rsq)}	
```

```{r}
#ta is vector of fluctuating temperature conditions
#times is a vector of the corresponding times
	dynamic.landscape <- function(ta,times,tolerance.landscape){
	  #These are the predicted survival times that correspond to set cumulative survival values for the average temperature
			surv <- tolerance.landscape$S[,2]
			#mean temperature from TDT curves
			ta.mn <- tolerance.landscape$ta.mn
			#thermal tolerance coefficient
			z <- tolerance.landscape$z
			#the change in cumulative survival for a given temperature
			shift <- 10^((ta.mn - ta)/z)	
			time.rel <- 0
			#this will become a vector of cumulative survival percentages
			alive <- 100
			for(i in 1:(length(ta)-1)){	
			  #this is the size of the timestep in question
			  tstep=times[i+1]-times[i]
				if(alive[length(na.omit(alive))] > 0) {		
				  #approx fits a linear regression of relative time 
					alive <- try(c(alive,approx(c(0,shift[i]*surv),seq(100,0,length.out = length(c(0,surv))),xout = time.rel[i] + tstep)$y),silent=TRUE)
					#This essentially uses the inverse CDF of the shifted survival curve to approximate the comparable static temperature time using the next temperature
					time.rel <- try(c(time.rel,approx(seq(100,0,length.out = length(c(0,surv))),c(0,shift[i + 1]*surv),xout = alive[i + 1])$y),silent=TRUE)}
				else{
					alive <- 0}}				
			out <- data.frame(cbind(ta=ta[1:(length(alive)-1)],time=times[1:(length(alive)-1)],alive=alive[1:(length(alive)-1)]))
			par(mar=c(4,4,1,1),mfrow=c(1,2))
			plot(1:length(ta),ta,type="l",xlim=c(0,length(ta)),ylim=c(min(ta),max(ta)),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Temperature (ºC)")			
			plot(out$time,out$alive,type="l",xlim=c(0,length(ta)),ylim=c(0,100),col="black",lwd=1.5,las=1,
				xlab = "Time (min)", ylab = "Survival (%)")
			list(time = out$time,ta = out$ta, alive = out$alive)}
```

```{r}
#for every time interval (t_i-1,t_i):
#1) Find max temperature over that time interval max(T_i,T_i-1)
#2) Find size of time interval (t_i-(t_i-1))
#3) Find damage accumulation rate: dd/dt=1/tKD(T)=1/(10^(Beta*T+alpha))
#4) Multiply rate by time interval to calculate accumulated damage for that time interval
#5) Add to total damage, record total damage at time t_i
#we start i-1 at the start time for the given trial group, and assume d=0 prior to this point. So, for trial group 1, i=2, but for trial group 3, which starts at about 117 mins, we start indexing at i=70 (calculations above)

jorg.mod = function(TDT.dat,temp.dat,KD.dat){
  #extract tdt coefficients
  tdt.lm = lm(log10.tcoma.~as.numeric(Tref),data=TDT.dat)
  tdt.slope=tdt.lm$coefficients[2]
  tdt.int=tdt.lm$coefficients[1]
  
  
  #injury ACC rate vector
acc.rates=rep(0,nrow(temp.dat))
#trial group 1
#initialize cumulative damage vector
cd.vec = rep(0,nrow(temp.dat))
#initialize timestep damage vector
td.vec = rep(0,nrow(temp.dat))
#initialize times vector
t.vec = temp.dat$Time_min
for (i in 2:nrow(temp.dat)){
  t_i = temp.dat$Time_min[i]
  t_prev = temp.dat$Time_min[i-1]
  T_i = temp.dat$Temperature[i]
  T_prev = temp.dat$Temperature[i-1]
  #1
  max.temp=max(T_i,T_prev)
  #2
  size.int=t_i-t_prev
  #3
  est.tKD = 10^(tdt.slope*max.temp+tdt.int)
  dd_dt = 1/est.tKD
  acc.rates[i]=dd_dt
  #4
  d.int = dd_dt*size.int
  td.vec[i-1]=d.int
  #5
  cd.vec[i]=cd.vec[i-1]+d.int
}
# 
# #now generate estimated damage at knockdown
# #initialize vector
dKD = rep(NA,nrow(KD.dat))

for (i in 1:nrow(KD.dat)){
   start.time = KD.dat$start[i]
   tKD = KD.dat$t_coma[i]
   dKD[i]=100*sum(td.vec[t.vec>=start.time & t.vec<=tKD])
}

est.tKD = rep(NA,nrow(KD.dat))
for (i in 1:nrow(KD.dat)){
  start.index.group = which.max(t.vec>=KD.dat$start[i])
  cd.group = cd.vec-cd.vec[start.index.group]
  est.tKD.index = which.max(cd.group>=1)
  est.tKD[i]=t.vec[est.tKD.index]
}
  
damage.df = data.frame(minutes=t.vec,cum.damage=cd.vec,interval.damage=td.vec)  
KD.dat$est.dKD = dKD  
KD.dat$est.tKD = est.tKD
#estimated and actual kdtimes starting at start time
KD.dat$est.tKD.adj = est.tKD-KD.dat$start
KD.dat$tcoma.adj=KD.dat$t_coma-KD.dat$start
ret.list=list(damage.df, KD.dat)
names(ret.list)=c("damage.df","KD.dat")


return(ret.list)
}
```

Estimating the MEAN log likelihood of the data given the simulated tf distribution using a kernel density estimate, then testing how likely the data was given the distribution by compluting a p-value via bootstrapping. Mean log-liklihood allows for standardization across differing sample sizes
```{r}
log_likelihood_test <- function(sims, data, n_boot = 10000, max_tries = 10, density_floor = 1e-10) {
  # Initial bandwidth
  base_bw <- bw.nrd0(sims)
  bw_factor <- 1
  attempt <- 1
  found_valid <- FALSE
  
  # Try increasing bandwidth until all data points get nonzero density
  while (attempt <= max_tries && !found_valid) {
    bw <- bw_factor * base_bw
    kde <- density(sims, bw = bw, n = 2048)
    kde_fn <- approxfun(kde$x, kde$y, rule = 2)
    
    densities_data <- kde_fn(data)
    
    if (all(densities_data > 0)) {
      found_valid <- TRUE
    } else {
      bw_factor <- bw_factor * 1.5
      attempt <- attempt + 1
    }
  }
  
  if (!found_valid) {
    warning("Log-likelihoods may be unstable: some data points still have zero estimated density after maximum bandwidth increase.")
  }

  # Apply floor to prevent log(0)
  densities_data <- pmax(densities_data, density_floor)
  mean_log_lik_data <- mean(log(densities_data))

  # Bootstrap: sample and compute log-likelihoods
  ll_dens <- replicate(n_boot, {
    sample_data <- sample(sims, size = length(data), replace = TRUE)
    sample_densities <- pmax(kde_fn(sample_data), density_floor)
    mean(log(sample_densities))
  })

  p_val <- mean(ll_dens <= mean_log_lik_data)

  return(list(
    ll.dens = ll_dens,
    ll.data = mean_log_lik_data,
    p_value = p_val,
    bandwidth_used = bw
  ))
}
```

```{r}
#constant temperature knockdown data
#used to train models
niko.tdt = read.csv("Niko_temps_csvs/Nikolaj_TDT_raw.csv")
#remove rows without Tref
niko.tdt=niko.tdt[is.na(niko.tdt$Tref)==F,]
#add columns for flexsurv formatting
niko.tdt$time=niko.tdt$tcoma
niko.tdt$event=rep(1,nrow(niko.tdt))
head(niko.tdt)
females.tdt = niko.tdt[niko.tdt$sex=="f",]
males.tdt = niko.tdt[niko.tdt$sex=="m",]
#remove 2 points with no tcoma (not censored, just missing)
males.tdt=males.tdt[is.na(males.tdt$tcoma)==FALSE,]
```

```{r}
#import trial temp condition data
niko.temps = list(
  read.csv("Niko_temps_csvs/Niko_temps_exp1.csv"),
  read.csv("Niko_temps_csvs/Niko_temps_exp2.csv"),
  read.csv("Niko_temps_csvs/Niko_temps_exp3.csv"))
names(niko.temps)=c("exp1","exp2","exp3")
```

```{r}
#knockdown data associated with these trial conditions
#used to test models
niko.fluc.tKD = read.csv("Niko_temps_csvs/Nikolaj_fluc_tKD_formatted.csv")
head(niko.fluc.tKD)
```

```{r}
#fluctuating knockdown data
#females exp 1
exp1.f.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==1),]
exp2.f.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2),]
exp3.f.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==3),]
```

Jorgensen analysis
```{r}
#females
jorg.exp1.f = jorg.mod(TDT.dat=females.tdt,temp.dat=niko.temps[[1]],KD.dat=exp1.f.tKD.dat)
jorg.exp2.f = jorg.mod(TDT.dat=females.tdt,temp.dat=niko.temps[[2]],KD.dat=exp2.f.tKD.dat)
jorg.exp3.f = jorg.mod(TDT.dat=females.tdt,temp.dat=niko.temps[[3]],KD.dat=exp3.f.tKD.dat)
jorg.f.comb=rbind(jorg.exp1.f$KD.dat,jorg.exp2.f$KD.dat,jorg.exp3.f$KD.dat)
```

```{r}
#fluctuating knockdown data
#males exp 1
exp1.m.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==1),]
exp2.m.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2),]
exp3.m.tKD.dat=niko.fluc.tKD[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==3),]
```

```{r}
#males
jorg.exp1.m = jorg.mod(TDT.dat=males.tdt,temp.dat=niko.temps[[1]],KD.dat=exp1.m.tKD.dat)
jorg.exp2.m = jorg.mod(TDT.dat=males.tdt,temp.dat=niko.temps[[2]],KD.dat=exp2.m.tKD.dat)
jorg.exp3.m = jorg.mod(TDT.dat=males.tdt,temp.dat=niko.temps[[3]],KD.dat=exp3.m.tKD.dat)
jorg.m.comb=rbind(jorg.exp1.m$KD.dat,jorg.exp2.m$KD.dat,jorg.exp3.m$KD.dat)
plot(est.dKD~tcoma.adj,data=jorg.m.comb)
```

The Weibull, LogLogistic and Rezende model functions just make predictions based on temp conditions: they don't adjust for group start times. So, we need to define the temperature conditions for each group.
```{r}
strt.times = exp1.f.tKD.dat %>% group_by(grp) %>% summarise(mean=mean(start))

#fluctuating temperature dataframe starting at each groups start time
temps.e1.g1 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[1,2]):nrow(niko.temps[[1]]),]
#subtract start time to rescale
temps.e1.g1$Time_min=temps.e1.g1$Time_min - temps.e1.g1$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[1,2])]

temps.e1.g2 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[[2,2]]):nrow(niko.temps[[1]]),]

temps.e1.g2$Time_min=temps.e1.g2$Time_min - temps.e1.g2$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[2,2])]

temps.e1.g3 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[[3,2]]):nrow(niko.temps[[1]]),]

temps.e1.g3$Time_min=temps.e1.g3$Time_min - temps.e1.g3$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[3,2])]

temps.e1.g4 = niko.temps[[1]][which.max(niko.temps[[1]]$Time_min>=strt.times[[4,2]]):nrow(niko.temps[[1]]),]

temps.e1.g4$Time_min=temps.e1.g4$Time_min - temps.e1.g4$Time_min[which.max(niko.temps[[1]]$Time_min>=strt.times[4,2])]
```

#exp 2
```{r}
strt.times = exp2.f.tKD.dat %>% group_by(grp) %>% summarise(mean=mean(start))

#fluctuating temperature dataframe starting at each groups start time
temps.e2.g1 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[1,2]):nrow(niko.temps[[2]]),]
#subtract start time to rescale
temps.e2.g1$Time_min=temps.e2.g1$Time_min - temps.e2.g1$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[1,2])]

temps.e2.g2 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[2,2]]):nrow(niko.temps[[2]]),]

temps.e2.g2$Time_min=temps.e2.g2$Time_min - temps.e2.g2$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[2,2])]

temps.e2.g3 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[3,2]]):nrow(niko.temps[[2]]),]

temps.e2.g3$Time_min=temps.e2.g3$Time_min - temps.e2.g3$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[3,2])]

temps.e2.g4 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[4,2]]):nrow(niko.temps[[2]]),]

temps.e2.g4$Time_min=temps.e2.g4$Time_min - temps.e2.g4$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[4,2])]

temps.e2.g5 = niko.temps[[2]][which.max(niko.temps[[2]]$Time_min>=strt.times[[5,2]]):nrow(niko.temps[[2]]),]

temps.e2.g5$Time_min=temps.e2.g5$Time_min - temps.e2.g5$Time_min[which.max(niko.temps[[2]]$Time_min>=strt.times[5,2])]
```

```{r}
#exp 3
strt.times = exp3.f.tKD.dat %>% group_by(grp) %>% summarise(mean=mean(start))

#fluctuating temperature dataframe starting at each groups start time
temps.e3.g1 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[1,2]):nrow(niko.temps[[3]]),]
#subtract start time to rescale
temps.e3.g1$Time_min=temps.e3.g1$Time_min - temps.e3.g1$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[1,2])]

temps.e3.g2 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[[2,2]]):nrow(niko.temps[[3]]),]

temps.e3.g2$Time_min=temps.e3.g2$Time_min - temps.e3.g2$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[2,2])]

temps.e3.g3 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[[3,2]]):nrow(niko.temps[[3]]),]

temps.e3.g3$Time_min=temps.e3.g3$Time_min - temps.e3.g3$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[3,2])]

temps.e3.g4 = niko.temps[[3]][which.max(niko.temps[[3]]$Time_min>=strt.times[[4,2]]):nrow(niko.temps[[3]]),]

temps.e3.g4$Time_min=temps.e3.g4$Time_min - temps.e3.g4$Time_min[which.max(niko.temps[[3]]$Time_min>=strt.times[4,2])]

```

```{r}
#actual failure times
#reset failure time to begin from start time
niko.fluc.tKD$t_coma=niko.fluc.tKD$t_coma-niko.fluc.tKD$start
#exp 1
tf.f.e1.g1 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==1)]
tf.m.e1.g1 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==1)]
tf.f.e1.g2 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==2)]
tf.m.e1.g2 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==2)]
tf.f.e1.g3 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==3)]
tf.m.e1.g3 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==3)]
tf.f.e1.g4 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==4)]
tf.m.e1.g4 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==1 & niko.fluc.tKD$grp==4)]
#exp 2
tf.f.e2.g1 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==1)]
tf.m.e2.g1 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==1)]
tf.f.e2.g2 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==2)]
tf.m.e2.g2 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==2)]
tf.f.e2.g3 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==3)]
tf.m.e2.g3 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==3)]
tf.f.e2.g4 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==4)]
tf.m.e2.g4 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==4)]
tf.f.e2.g5 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==5)]
tf.m.e2.g5 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==2 & niko.fluc.tKD$grp==5)]
#exp 3
tf.f.e3.g1 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==1)]
tf.m.e3.g1 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==1)]
tf.f.e3.g2 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==2)]
tf.m.e3.g2 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==2)]
tf.f.e3.g3 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==3)]
tf.m.e3.g3 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==3)]
tf.f.e3.g4 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="f" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==4)]
tf.m.e3.g4 = niko.fluc.tKD$t_coma[(niko.fluc.tKD$sex=="m" & niko.fluc.tKD$expno==3 & niko.fluc.tKD$grp==4)]

```



llogis Analysis (linear shape)
```{r}
#exp 1
#females
ll.e1.g1.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g1)
ll.e1.g2.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g2)
ll.e1.g3.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g3)
ll.e1.g4.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g4)
c(ll.e1.g1.f$med.tdt,ll.e1.g2.f$med.tdt,ll.e1.g3.f$med.tdt,ll.e1.g4.f$med.tdt)
c(ll.e1.g1.f$expected.tdt,ll.e1.g2.f$expected.tdt,ll.e1.g3.f$expected.tdt,ll.e1.g4.f$expected.tdt)
#males
ll.e1.g1.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g1)
ll.e1.g2.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g2)
ll.e1.g3.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g3)
ll.e1.g4.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g4)
```

```{r}
#exp 2
ll.e2.g1.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g1)
ll.e2.g2.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g2)
ll.e2.g3.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g3)
ll.e2.g4.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g4)
ll.e2.g5.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g5)
c(ll.e2.g1.f$med.tdt,ll.e2.g2.f$med.tdt,ll.e2.g3.f$med.tdt,ll.e2.g4.f$med.tdt,ll.e2.g5.f$med.tdt)
c(ll.e2.g1.f$expected.tdt,ll.e2.g2.f$expected.tdt,ll.e2.g3.f$expected.tdt,ll.e2.g4.f$expected.tdt,ll.e2.g5.f$med.tdt)
#males
ll.e2.g1.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g1)
ll.e2.g2.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g2)
ll.e2.g3.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g3)
ll.e2.g4.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g4)
ll.e2.g5.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g5)
```

```{r}
#exp 3
ll.e3.g1.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g1)
ll.e3.g2.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g2)
ll.e3.g3.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g3)
ll.e3.g4.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g4)
c(ll.e3.g1.f$med.tdt,ll.e3.g2.f$med.tdt,ll.e3.g3.f$med.tdt,ll.e3.g4.f$med.tdt)
c(ll.e3.g1.f$expected.tdt,ll.e3.g2.f$expected.tdt,ll.e3.g3.f$expected.tdt,ll.e3.g4.f$expected.tdt)
#males
ll.e3.g1.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g1)
ll.e3.g2.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g2)
ll.e3.g3.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g3)
ll.e3.g4.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g4)
```

llogis Analysis (constant shape)

```{r}
#exp 1
#females
ll.cons.e1.g1.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g1,shape="constant")
ll.cons.e1.g2.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g2,shape="constant")
ll.cons.e1.g3.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g3,shape="constant")
ll.cons.e1.g4.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e1.g4,shape="constant")
c(ll.cons.e1.g1.f$med.tdt,ll.cons.e1.g2.f$med.tdt,ll.cons.e1.g3.f$med.tdt,ll.cons.e1.g4.f$med.tdt)
c(ll.cons.e1.g1.f$expected.tdt,ll.cons.e1.g2.f$expected.tdt,ll.cons.e1.g3.f$expected.tdt,ll.cons.e1.g4.f$expected.tdt)
#males
ll.cons.e1.g1.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g1,shape="constant")
ll.cons.e1.g2.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g2,shape="constant")
ll.cons.e1.g3.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g3,shape="constant")
ll.cons.e1.g4.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e1.g4,shape="constant")
```

```{r}
#exp 2
ll.cons.e2.g1.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g1,shape="constant")
ll.cons.e2.g2.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g2,shape="constant")
ll.cons.e2.g3.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g3,shape="constant")
ll.cons.e2.g4.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g4,shape="constant")
ll.cons.e2.g5.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e2.g5,shape="constant")
#males
ll.cons.e2.g1.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g1,shape="constant")
ll.cons.e2.g2.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g2,shape="constant")
ll.cons.e2.g3.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g3,shape="constant")
ll.cons.e2.g4.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g4,shape="constant")
ll.cons.e2.g5.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e2.g5,shape="constant")
```

```{r}
#exp 3
ll.cons.e3.g1.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g1,shape="constant")
ll.cons.e3.g2.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g2,shape="constant")
ll.cons.e3.g3.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g3,shape="constant")
ll.cons.e3.g4.f = llogis.mod(TDT.dat=females.tdt,temp.dat=temps.e3.g4,shape="constant")
#males
ll.cons.e3.g1.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g1,shape="constant")
ll.cons.e3.g2.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g2,shape="constant")
ll.cons.e3.g3.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g3,shape="constant")
ll.cons.e3.g4.m = llogis.mod(TDT.dat=males.tdt,temp.dat=temps.e3.g4,shape="constant")
```

Rezende model output
Note: these functions are slightly modified to account for temperatures not collected once per minute
```{r}
#females
tol.land.females=tolerance.landscape(females.tdt$Tref,females.tdt$tcoma)
#exp 1
dy.land.e1.g1.f = dynamic.landscape(temps.e1.g1$Temperature,temps.e1.g1$Time_min,tol.land.females)
dy.land.e1.g2.f = dynamic.landscape(temps.e1.g2$Temperature,temps.e1.g2$Time_min,tol.land.females)
dy.land.e1.g3.f = dynamic.landscape(temps.e1.g3$Temperature,temps.e1.g3$Time_min,tol.land.females)
dy.land.e1.g4.f = dynamic.landscape(temps.e1.g4$Temperature,temps.e1.g4$Time_min,tol.land.females)
#median fail time
rz.med.e1.g1.f = dy.land.e1.g1.f$time[which.max(dy.land.e1.g1.f$alive<50)]
rz.med.e1.g2.f = dy.land.e1.g2.f$time[which.max(dy.land.e1.g2.f$alive<50)]
rz.med.e1.g3.f = dy.land.e1.g3.f$time[which.max(dy.land.e1.g3.f$alive<50)]
rz.med.e1.g4.f = dy.land.e1.g4.f$time[which.max(dy.land.e1.g4.f$alive<50)]

#exp 2
dy.land.e2.g1.f = dynamic.landscape(temps.e2.g1$Temperature,temps.e2.g1$Time_min,tol.land.females)
dy.land.e2.g2.f = dynamic.landscape(temps.e2.g2$Temperature,temps.e2.g2$Time_min,tol.land.females)
dy.land.e2.g3.f = dynamic.landscape(temps.e2.g3$Temperature,temps.e2.g3$Time_min,tol.land.females)
dy.land.e2.g4.f = dynamic.landscape(temps.e2.g4$Temperature,temps.e2.g4$Time_min,tol.land.females)
dy.land.e2.g5.f = dynamic.landscape(temps.e2.g5$Temperature,temps.e2.g5$Time_min,tol.land.females)
#median fail time
rz.med.e2.g1.f = dy.land.e2.g1.f$time[which.max(dy.land.e2.g1.f$alive<50)]
rz.med.e2.g2.f = dy.land.e2.g2.f$time[which.max(dy.land.e2.g2.f$alive<50)]
rz.med.e2.g3.f = dy.land.e2.g3.f$time[which.max(dy.land.e2.g3.f$alive<50)]
rz.med.e2.g4.f = dy.land.e2.g4.f$time[which.max(dy.land.e2.g4.f$alive<50)]
rz.med.e2.g5.f = dy.land.e2.g5.f$time[which.max(dy.land.e2.g5.f$alive<50)]

#exp 3
dy.land.e3.g1.f = dynamic.landscape(temps.e3.g1$Temperature,temps.e3.g1$Time_min,tol.land.females)
dy.land.e3.g2.f = dynamic.landscape(temps.e3.g2$Temperature,temps.e3.g2$Time_min,tol.land.females)
dy.land.e3.g3.f = dynamic.landscape(temps.e3.g3$Temperature,temps.e3.g3$Time_min,tol.land.females)
dy.land.e3.g4.f = dynamic.landscape(temps.e3.g4$Temperature,temps.e3.g4$Time_min,tol.land.females)
#median fail time
rz.med.e3.g1.f = dy.land.e3.g1.f$time[which.max(dy.land.e3.g1.f$alive<50)]
rz.med.e3.g2.f = dy.land.e3.g2.f$time[which.max(dy.land.e3.g2.f$alive<50)]
rz.med.e3.g3.f = dy.land.e3.g3.f$time[which.max(dy.land.e3.g3.f$alive<50)]
rz.med.e3.g4.f = dy.land.e3.g4.f$time[which.max(dy.land.e3.g4.f$alive<50)]

#males
tol.land.males=tolerance.landscape(males.tdt$Tref,males.tdt$tcoma)
#exp 1
dy.land.e1.g1.m = dynamic.landscape(temps.e1.g1$Temperature,temps.e1.g1$Time_min,tol.land.males)
dy.land.e1.g2.m = dynamic.landscape(temps.e1.g2$Temperature,temps.e1.g2$Time_min,tol.land.males)
dy.land.e1.g3.m = dynamic.landscape(temps.e1.g3$Temperature,temps.e1.g3$Time_min,tol.land.males)
dy.land.e1.g4.m = dynamic.landscape(temps.e1.g4$Temperature,temps.e1.g4$Time_min,tol.land.males)
#median fail time
rz.med.e1.g1.m = dy.land.e1.g1.m$time[which.max(dy.land.e1.g1.m$alive<50)]
rz.med.e1.g2.m = dy.land.e1.g2.m$time[which.max(dy.land.e1.g2.m$alive<50)]
rz.med.e1.g3.m = dy.land.e1.g3.m$time[which.max(dy.land.e1.g3.m$alive<50)]
rz.med.e1.g4.m = dy.land.e1.g4.m$time[which.max(dy.land.e1.g4.m$alive<50)]

#exp 2
dy.land.e2.g1.m = dynamic.landscape(temps.e2.g1$Temperature,temps.e2.g1$Time_min,tol.land.males)
dy.land.e2.g2.m = dynamic.landscape(temps.e2.g2$Temperature,temps.e2.g2$Time_min,tol.land.males)
dy.land.e2.g3.m = dynamic.landscape(temps.e2.g3$Temperature,temps.e2.g3$Time_min,tol.land.males)
dy.land.e2.g4.m = dynamic.landscape(temps.e2.g4$Temperature,temps.e2.g4$Time_min,tol.land.males)
dy.land.e2.g5.m = dynamic.landscape(temps.e2.g5$Temperature,temps.e2.g5$Time_min,tol.land.males)
#median fail time
rz.med.e2.g1.m = dy.land.e2.g1.m$time[which.max(dy.land.e2.g1.m$alive<50)]
rz.med.e2.g2.m = dy.land.e2.g2.m$time[which.max(dy.land.e2.g2.m$alive<50)]
rz.med.e2.g3.m = dy.land.e2.g3.m$time[which.max(dy.land.e2.g3.m$alive<50)]
rz.med.e2.g4.m = dy.land.e2.g4.m$time[which.max(dy.land.e2.g4.m$alive<50)]
rz.med.e2.g5.m = dy.land.e2.g5.m$time[which.max(dy.land.e2.g5.m$alive<50)]

#exp 3
dy.land.e3.g1.m = dynamic.landscape(temps.e3.g1$Temperature,temps.e3.g1$Time_min,tol.land.males)
dy.land.e3.g2.m = dynamic.landscape(temps.e3.g2$Temperature,temps.e3.g2$Time_min,tol.land.males)
dy.land.e3.g3.m = dynamic.landscape(temps.e3.g3$Temperature,temps.e3.g3$Time_min,tol.land.males)
dy.land.e3.g4.m = dynamic.landscape(temps.e3.g4$Temperature,temps.e3.g4$Time_min,tol.land.males)
#median fail time
rz.med.e3.g1.m = dy.land.e3.g1.m$time[which.max(dy.land.e3.g1.m$alive<50)]
rz.med.e3.g2.m = dy.land.e3.g2.m$time[which.max(dy.land.e3.g2.m$alive<50)]
rz.med.e3.g3.m = dy.land.e3.g3.m$time[which.max(dy.land.e3.g3.m$alive<50)]
rz.med.e3.g4.m = dy.land.e3.g4.m$time[which.max(dy.land.e3.g4.m$alive<50)]

```


```{r}
#let's compare model outputs from the four models
#females
#exp 1
exp1.f.summ = jorg.exp1.f$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
exp1.f.summ$rz.med.tkd=c(rz.med.e1.g1.f,rz.med.e1.g2.f,rz.med.e1.g3.f,rz.med.e1.g4.f)
exp1.f.summ$exp=1

#exp 2
exp2.f.summ = jorg.exp2.f$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
exp2.f.summ$rz.med.tkd=c(rz.med.e2.g1.f,rz.med.e2.g2.f,rz.med.e2.g3.f,rz.med.e2.g4.f,rz.med.e2.g5.f)
exp2.f.summ$exp=2

#exp 3
exp3.f.summ = jorg.exp3.f$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
exp3.f.summ$rz.med.tkd=c(rz.med.e3.g1.f,rz.med.e3.g2.f,rz.med.e3.g3.f,rz.med.e3.g4.f)
exp3.f.summ$exp=3

#ADD LLOGIT DATA 
exp1.f.summ$ll.med.tkd=c(ll.e1.g1.f$med.tdt,ll.e1.g2.f$med.tdt,ll.e1.g3.f$med.tdt,ll.e1.g4.f$med.tdt)
exp1.f.summ$ll.cons.med.tkd=c(ll.cons.e1.g1.f$med.tdt,ll.cons.e1.g2.f$med.tdt,ll.cons.e1.g3.f$med.tdt,ll.cons.e1.g4.f$med.tdt)
#linear shape
exp1.f.summ$ll.mean.tkd=c(ll.e1.g1.f$expected.tdt,ll.e1.g2.f$expected.tdt,ll.e1.g3.f$expected.tdt,ll.e1.g4.f$expected.tdt)
#constant shape
exp1.f.summ$ll.cons.mean.tkd=c(ll.cons.e1.g1.f$expected.tdt,ll.cons.e1.g2.f$expected.tdt,ll.cons.e1.g3.f$expected.tdt,ll.cons.e1.g4.f$expected.tdt)

#exp 2
#linear shape
exp2.f.summ$ll.med.tkd=c(ll.e2.g1.f$med.tdt,ll.e2.g2.f$med.tdt,ll.e2.g3.f$med.tdt,ll.e2.g4.f$med.tdt,ll.e2.g5.f$med.tdt)
exp2.f.summ$ll.mean.tkd=c(ll.e2.g1.f$expected.tdt,ll.e2.g2.f$expected.tdt,ll.e2.g3.f$expected.tdt,ll.e2.g4.f$expected.tdt,ll.e2.g5.f$med.tdt)
#constant shape
exp2.f.summ$ll.cons.med.tkd=c(ll.cons.e2.g1.f$med.tdt,ll.cons.e2.g2.f$med.tdt,ll.cons.e2.g3.f$med.tdt,ll.cons.e2.g4.f$med.tdt,ll.cons.e2.g5.f$med.tdt)
exp2.f.summ$ll.cons.mean.tkd=c(ll.cons.e2.g1.f$expected.tdt,ll.cons.e2.g2.f$expected.tdt,ll.cons.e2.g3.f$expected.tdt,ll.cons.e2.g4.f$expected.tdt,ll.cons.e2.g5.f$med.tdt)

#exp 3
#linear shape
exp3.f.summ$ll.med.tkd=c(ll.e3.g1.f$med.tdt,ll.e3.g2.f$med.tdt,ll.e3.g3.f$med.tdt,ll.e3.g4.f$med.tdt)
exp3.f.summ$ll.mean.tkd=c(ll.e3.g1.f$expected.tdt,ll.e3.g2.f$expected.tdt,ll.e3.g3.f$expected.tdt,ll.e3.g4.f$expected.tdt)
#constant shape
exp3.f.summ$ll.cons.med.tkd=c(ll.cons.e3.g1.f$med.tdt,ll.cons.e3.g2.f$med.tdt,ll.cons.e3.g3.f$med.tdt,ll.cons.e3.g4.f$med.tdt)
exp3.f.summ$ll.cons.mean.tkd=c(ll.cons.e3.g1.f$expected.tdt,ll.cons.e3.g2.f$expected.tdt,ll.cons.e3.g3.f$expected.tdt,ll.cons.e3.g4.f$expected.tdt)


four.mod.results.f = rbind(exp1.f.summ,exp2.f.summ,exp3.f.summ)
four.mod.results.f$sex="f"

four.mod.results.f$jorg.error=four.mod.results.f$est.tKD-four.mod.results.f$true.tKD.med
four.mod.results.f$rz.error=four.mod.results.f$rz.med.tkd-four.mod.results.f$true.tKD.med
four.mod.results.f$ll.error=four.mod.results.f$ll.med.tkd-four.mod.results.f$true.tKD.med
four.mod.results.f$ll.cons.error=four.mod.results.f$ll.cons.med.tkd-four.mod.results.f$true.tKD.med

#Note: we don't need to use a comparison metric like AIC because our fluctuating temp data was not used to create our predictions, so overfitting is not a worry.

sum(abs(four.mod.results.f$jorg.error))
sum(abs(four.mod.results.f$rz.error))
sum(abs(four.mod.results.f$ll.error))
sum(abs(four.mod.results.f$ll.cons.error))
sum((four.mod.results.f$jorg.error)^2)
sum((four.mod.results.f$rz.error)^2)
sum((four.mod.results.f$ll.error)^2)
sum((four.mod.results.f$ll.cons.error)^2)
```

```{r}
#males
#exp 1
exp1.m.summ = jorg.exp1.m$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))

exp1.m.summ$rz.med.tkd=c(rz.med.e1.g1.m,rz.med.e1.g2.m,rz.med.e1.g3.m,rz.med.e1.g4.m)
exp1.m.summ$exp=1

#exp 2
exp2.m.summ = jorg.exp2.m$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
exp2.m.summ$rz.med.tkd=c(rz.med.e2.g1.m,rz.med.e2.g2.m,rz.med.e2.g3.m,rz.med.e2.g4.m,rz.med.e2.g5.m)
exp2.m.summ$exp=2

#exp 3
exp3.m.summ = jorg.exp3.m$KD.dat %>% group_by(grp) %>% summarise(true.tKD.mean=mean(tcoma.adj),true.tKD.med=median(tcoma.adj),est.tKD=mean(est.tKD.adj))
exp3.m.summ$rz.med.tkd=c(rz.med.e3.g1.m,rz.med.e3.g2.m,rz.med.e3.g3.m,rz.med.e3.g4.m)
exp3.m.summ$exp=3

#ADD LLOGIT DATA 
exp1.m.summ$ll.med.tkd=c(ll.e1.g1.m$med.tdt,ll.e1.g2.m$med.tdt,ll.e1.g3.m$med.tdt,ll.e1.g4.m$med.tdt)
exp1.m.summ$ll.cons.med.tkd=c(ll.cons.e1.g1.m$med.tdt,ll.cons.e1.g2.m$med.tdt,ll.cons.e1.g3.m$med.tdt,ll.cons.e1.g4.m$med.tdt)
#linear shape
exp1.m.summ$ll.mean.tkd=c(ll.e1.g1.m$expected.tdt,ll.e1.g2.m$expected.tdt,ll.e1.g3.m$expected.tdt,ll.e1.g4.m$expected.tdt)
#constant shape
exp1.m.summ$ll.cons.mean.tkd=c(ll.cons.e1.g1.m$expected.tdt,ll.cons.e1.g2.m$expected.tdt,ll.cons.e1.g3.m$expected.tdt,ll.cons.e1.g4.m$expected.tdt)

#exp 2
#linear shape
exp2.m.summ$ll.med.tkd=c(ll.e2.g1.m$med.tdt,ll.e2.g2.m$med.tdt,ll.e2.g3.m$med.tdt,ll.e2.g4.m$med.tdt,ll.e2.g5.m$med.tdt)
exp2.m.summ$ll.mean.tkd=c(ll.e2.g1.m$expected.tdt,ll.e2.g2.m$expected.tdt,ll.e2.g3.m$expected.tdt,ll.e2.g4.m$expected.tdt,ll.e2.g5.m$med.tdt)
#constant shape
exp2.m.summ$ll.cons.med.tkd=c(ll.cons.e2.g1.m$med.tdt,ll.cons.e2.g2.m$med.tdt,ll.cons.e2.g3.m$med.tdt,ll.cons.e2.g4.m$med.tdt,ll.cons.e2.g5.m$med.tdt)
exp2.m.summ$ll.cons.mean.tkd=c(ll.cons.e2.g1.m$expected.tdt,ll.cons.e2.g2.m$expected.tdt,ll.cons.e2.g3.m$expected.tdt,ll.cons.e2.g4.m$expected.tdt,ll.cons.e2.g5.m$med.tdt)

#exp 3
#linear shape
exp3.m.summ$ll.med.tkd=c(ll.e3.g1.m$med.tdt,ll.e3.g2.m$med.tdt,ll.e3.g3.m$med.tdt,ll.e3.g4.m$med.tdt)
exp3.m.summ$ll.mean.tkd=c(ll.e3.g1.m$expected.tdt,ll.e3.g2.m$expected.tdt,ll.e3.g3.m$expected.tdt,ll.e3.g4.m$expected.tdt)
#constant shape
exp3.m.summ$ll.cons.med.tkd=c(ll.cons.e3.g1.m$med.tdt,ll.cons.e3.g2.m$med.tdt,ll.cons.e3.g3.m$med.tdt,ll.cons.e3.g4.m$med.tdt)
exp3.m.summ$ll.cons.mean.tkd=c(ll.cons.e3.g1.m$expected.tdt,ll.cons.e3.g2.m$expected.tdt,ll.cons.e3.g3.m$expected.tdt,ll.cons.e3.g4.m$expected.tdt)


four.mod.results.m = rbind(exp1.m.summ,exp2.m.summ,exp3.m.summ)
four.mod.results.m$sex="m"
#note that later I will calculate error based on individual points, not medians. That will be better, but this will give us an idea for now.
four.mod.results.m$jorg.error=four.mod.results.m$est.tKD-four.mod.results.m$true.tKD.med
four.mod.results.m$rz.error=four.mod.results.m$rz.med.tkd-four.mod.results.m$true.tKD.med
four.mod.results.m$ll.error=four.mod.results.m$ll.med.tkd-four.mod.results.m$true.tKD.med
four.mod.results.m$ll.cons.error=four.mod.results.m$ll.cons.med.tkd-four.mod.results.m$true.tKD.med
sum(abs(four.mod.results.m$jorg.error))
sum(abs(four.mod.results.m$rz.error))
sum(abs(four.mod.results.m$ll.error))
sum(abs(four.mod.results.m$ll.cons.error))
sum((four.mod.results.m$jorg.error)^2)
sum((four.mod.results.m$rz.error)^2)
sum((four.mod.results.m$ll.error)^2)
sum((four.mod.results.m$ll.cons.error)^2)
```

```{r}
#combine data
four.mod.results = rbind(four.mod.results.f[,colnames(four.mod.results.f)%in%colnames(four.mod.results.m)],four.mod.results.m)
```

make plot (Fig 4 panels a and b)
```{r}
fig4.males = ggplot(data = four.mod.results.m) +
  geom_point(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Increasing Variance"), size = 3) +
  geom_point(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), size = 3) +
  geom_point(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), size = 3) +
  geom_abline(slope = 1, intercept = 0,linetype="dashed") +
  geom_smooth(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Increasing Variance"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), method = "lm", se = FALSE) +
  labs(
    title = "Males",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("#440154FF", "#22A884FF", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
  plot.tag = element_text(size = 16)
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )+
  theme(
  legend.position = c(0.98, 0.05),   
  legend.justification = c(1, 0),
  legend.key.size = unit(10, "pt"),     
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12)
)
  
```

```{r}
fig4.females = ggplot(data = four.mod.results.f) +
  geom_point(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Increasing Variance"), size = 3) +
  geom_point(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), size = 3) +
  geom_point(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), size = 3) +
  geom_abline(slope = 1, intercept = 0,linetype="dashed") +
  geom_smooth(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Increasing Variance"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(est.tKD), x = log10(true.tKD.med), color = "Jorgensen"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = log10(rz.med.tkd), x = log10(true.tKD.med), color = "Rezende"), method = "lm", se = FALSE) +
  labs(
    title = "Females",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("#440154FF", "#22A884FF", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
    legend.position="none",
  plot.tag = element_text(size = 16)
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
  
```

Make supplemental figure (S5) comparing Constant and Varying Shape Log_logistic Models
```{r}
ggplot(data = four.mod.results.f) +
  geom_point(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Log-Logistic"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(y = log10(ll.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Log-Logistic"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Log-Logistic"), method = "lm", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(ll.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Log-Logistic"), method = "lm", se = FALSE) +
  labs(
    title = "Females",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("gray", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
```

```{r}
ggplot(data = four.mod.results.m) +
  geom_point(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Log-Logistic"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(y = log10(ll.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Log-Logistic"), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(ll.med.tkd), x = log10(true.tKD.med), color = "Linear Shape Log-Logistic"), method = "lm", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  geom_smooth(aes(y = log10(ll.cons.med.tkd), x = log10(true.tKD.med), color = "Constant Shape Log-Logistic"), method = "lm", se = FALSE) +
  labs(
    title = "Males",
    x = expression("Observed Median t"[f] * " (minutes)"),
    y = expression("Predicted Median t"[f] * " (minutes)"),
    color = "Model"
  ) +
  scale_color_manual(values = c("gray", "darkorange")) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  )
```

Next, let's compare the estimated distributions of failure from the Rezende and log-logistic models with the real failure data. To do with we will use the mean log-likelihood function run above.

```{r}
log_like_df = data.frame(ll.rz=rep(NA,26),ll.ll.lin=rep(NA,26),ll.ll.cons=rep(NA,26))
```


```{r}
#First up: Rezende
#We next want to extract log-likelihoods
#The problem is that the dynamic.landscape function returns a CDF (or really 1-CDF=survival curve), but the log likelihood function estimates a PDF through samples failure times. So we have to sample failure times through simulation and obtain a PDF, since it is not an analytical distribution.
Surv.to.tf.sims = function(Surv.df,n=10000){
  sims.vec=rep(NA,n)
  unifs = runif(n)*100
  for (i in 1:n){
    sims.vec[i] = Surv.df$time[which.max(Surv.df$alive<=unifs[i])]
  }
  return(sims.vec)
}

rz.sims.e1.g1.f =Surv.to.tf.sims(dy.land.e1.g1.f)
rz.sims.e1.g2.f =Surv.to.tf.sims(dy.land.e1.g2.f)
rz.sims.e1.g3.f =Surv.to.tf.sims(dy.land.e1.g3.f)
rz.sims.e1.g4.f =Surv.to.tf.sims(dy.land.e1.g4.f)
rz.sims.e2.g1.f =Surv.to.tf.sims(dy.land.e2.g1.f)
rz.sims.e2.g2.f =Surv.to.tf.sims(dy.land.e2.g2.f)
rz.sims.e2.g3.f =Surv.to.tf.sims(dy.land.e2.g3.f)
rz.sims.e2.g4.f =Surv.to.tf.sims(dy.land.e2.g4.f)
rz.sims.e2.g5.f =Surv.to.tf.sims(dy.land.e2.g5.f)
rz.sims.e3.g1.f =Surv.to.tf.sims(dy.land.e3.g1.f)
rz.sims.e3.g2.f =Surv.to.tf.sims(dy.land.e3.g2.f)
rz.sims.e3.g3.f =Surv.to.tf.sims(dy.land.e3.g3.f)
rz.sims.e3.g4.f =Surv.to.tf.sims(dy.land.e3.g4.f)
rz.sims.e1.g1.m =Surv.to.tf.sims(dy.land.e1.g1.m)
rz.sims.e1.g2.m =Surv.to.tf.sims(dy.land.e1.g2.m)
rz.sims.e1.g3.m =Surv.to.tf.sims(dy.land.e1.g3.m)
rz.sims.e1.g4.m =Surv.to.tf.sims(dy.land.e1.g4.m)
rz.sims.e2.g1.m =Surv.to.tf.sims(dy.land.e2.g1.m)
rz.sims.e2.g2.m =Surv.to.tf.sims(dy.land.e2.g2.m)
rz.sims.e2.g3.m =Surv.to.tf.sims(dy.land.e2.g3.m)
rz.sims.e2.g4.m =Surv.to.tf.sims(dy.land.e2.g4.m)
rz.sims.e2.g5.m =Surv.to.tf.sims(dy.land.e2.g5.m)
rz.sims.e3.g1.m =Surv.to.tf.sims(dy.land.e3.g1.m)
rz.sims.e3.g2.m =Surv.to.tf.sims(dy.land.e3.g2.m)
rz.sims.e3.g3.m =Surv.to.tf.sims(dy.land.e3.g3.m)
rz.sims.e3.g4.m =Surv.to.tf.sims(dy.land.e3.g4.m)
log_like_df$ll.rz[1] = log_likelihood_test(sims=rz.sims.e1.g1.f,data=tf.f.e1.g1)$ll.data
log_like_df$ll.rz[2] = log_likelihood_test(sims=rz.sims.e1.g2.f,data=tf.f.e1.g2)$ll.data
log_like_df$ll.rz[3] = log_likelihood_test(sims=rz.sims.e1.g3.f,data=tf.f.e1.g3)$ll.data
log_like_df$ll.rz[4] = log_likelihood_test(sims=rz.sims.e1.g4.f,data=tf.f.e1.g4)$ll.data
log_like_df$ll.rz[5] = log_likelihood_test(sims=rz.sims.e2.g1.f,data=tf.f.e2.g1)$ll.data
log_like_df$ll.rz[6] = log_likelihood_test(sims=rz.sims.e2.g2.f,data=tf.f.e2.g2)$ll.data
log_like_df$ll.rz[7] = log_likelihood_test(sims=rz.sims.e2.g3.f,data=tf.f.e2.g3)$ll.data
log_like_df$ll.rz[8] = log_likelihood_test(sims=rz.sims.e2.g4.f,data=tf.f.e2.g4)$ll.data
log_like_df$ll.rz[9] = log_likelihood_test(sims=rz.sims.e2.g5.f,data=tf.f.e2.g5)$ll.data
log_like_df$ll.rz[10] = log_likelihood_test(sims=rz.sims.e3.g1.f,data=tf.f.e3.g1)$ll.data
log_like_df$ll.rz[11] = log_likelihood_test(sims=rz.sims.e3.g2.f,data=tf.f.e3.g2)$ll.data
log_like_df$ll.rz[12] = log_likelihood_test(sims=rz.sims.e3.g3.f,data=tf.f.e3.g3)$ll.data
log_like_df$ll.rz[13] = log_likelihood_test(sims=rz.sims.e3.g4.f,data=tf.f.e3.g4)$ll.data
log_like_df$ll.rz[14] = log_likelihood_test(sims=rz.sims.e1.g1.m,data=tf.m.e1.g1)$ll.data
log_like_df$ll.rz[15] = log_likelihood_test(sims=rz.sims.e1.g2.m,data=tf.m.e1.g2)$ll.data
log_like_df$ll.rz[16] = log_likelihood_test(sims=rz.sims.e1.g3.m,data=tf.m.e1.g3)$ll.data
log_like_df$ll.rz[17] = log_likelihood_test(sims=rz.sims.e1.g4.m,data=tf.m.e1.g4)$ll.data
log_like_df$ll.rz[18] = log_likelihood_test(sims=rz.sims.e2.g1.m,data=tf.m.e2.g1)$ll.data
log_like_df$ll.rz[19] = log_likelihood_test(sims=rz.sims.e2.g2.m,data=tf.m.e2.g2)$ll.data
log_like_df$ll.rz[20] = log_likelihood_test(sims=rz.sims.e2.g3.m,data=tf.m.e2.g3)$ll.data
log_like_df$ll.rz[21] = log_likelihood_test(sims=rz.sims.e2.g4.m,data=tf.m.e2.g4)$ll.data
log_like_df$ll.rz[22] = log_likelihood_test(sims=rz.sims.e2.g5.m,data=tf.m.e2.g5)$ll.data
log_like_df$ll.rz[23] = log_likelihood_test(sims=rz.sims.e3.g1.m,data=tf.m.e3.g1)$ll.data
log_like_df$ll.rz[24] = log_likelihood_test(sims=rz.sims.e3.g2.m,data=tf.m.e3.g2)$ll.data
log_like_df$ll.rz[25] = log_likelihood_test(sims=rz.sims.e3.g3.m,data=tf.m.e3.g3)$ll.data
log_like_df$ll.rz[26] = log_likelihood_test(sims=rz.sims.e3.g4.m,data=tf.m.e3.g4)$ll.data
```

```{r}
#now for log-logistic linear
#need to reformat dataframe
ll.temps.e1.g1.f = data.frame(time=temps.e1.g1$Time_min[1:length(c(ll.e1.g1.f$cum.surv))], alive = c(ll.e1.g1.f$cum.surv)*100)
ll.temps.e1.g2.f = data.frame(time=temps.e1.g2$Time_min[1:length(c(ll.e1.g2.f$cum.surv))], alive = c(ll.e1.g2.f$cum.surv)*100)
ll.temps.e1.g3.f = data.frame(time=temps.e1.g3$Time_min[1:length(c(ll.e1.g3.f$cum.surv))], alive = c(ll.e1.g3.f$cum.surv)*100)
ll.temps.e1.g4.f = data.frame(time=temps.e1.g4$Time_min[1:length(c(ll.e1.g4.f$cum.surv))], alive = c(ll.e1.g4.f$cum.surv)*100)
ll.temps.e2.g1.f = data.frame(time=temps.e2.g1$Time_min[1:length(c(ll.e2.g1.f$cum.surv))], alive = c(ll.e2.g1.f$cum.surv)*100)
ll.temps.e2.g2.f = data.frame(time=temps.e2.g2$Time_min[1:length(c(ll.e2.g2.f$cum.surv))], alive = c(ll.e2.g2.f$cum.surv)*100)
ll.temps.e2.g3.f = data.frame(time=temps.e2.g3$Time_min[1:length(c(ll.e2.g3.f$cum.surv))], alive = c(ll.e2.g3.f$cum.surv)*100)
ll.temps.e2.g4.f = data.frame(time=temps.e2.g4$Time_min[1:length(c(ll.e2.g4.f$cum.surv))], alive = c(ll.e2.g4.f$cum.surv)*100)
ll.temps.e2.g5.f = data.frame(time=temps.e2.g5$Time_min[1:length(c(ll.e2.g5.f$cum.surv))], alive = c(ll.e2.g5.f$cum.surv)*100)
ll.temps.e3.g1.f = data.frame(time=temps.e3.g1$Time_min[1:length(c(ll.e3.g1.f$cum.surv))], alive = c(ll.e3.g1.f$cum.surv)*100)
ll.temps.e3.g2.f = data.frame(time=temps.e3.g2$Time_min[1:length(c(ll.e3.g2.f$cum.surv))], alive = c(ll.e3.g2.f$cum.surv)*100)
ll.temps.e3.g3.f = data.frame(time=temps.e3.g3$Time_min[1:length(c(ll.e3.g3.f$cum.surv))], alive = c(ll.e3.g3.f$cum.surv)*100)
ll.temps.e3.g4.f = data.frame(time=temps.e3.g4$Time_min[1:length(c(ll.e3.g4.f$cum.surv))], alive = c(ll.e3.g4.f$cum.surv)*100)
ll.temps.e1.g1.m = data.frame(time=temps.e1.g1$Time_min[1:length(c(ll.e1.g1.m$cum.surv))], alive = c(ll.e1.g1.m$cum.surv)*100)
ll.temps.e1.g2.m = data.frame(time=temps.e1.g2$Time_min[1:length(c(ll.e1.g2.m$cum.surv))], alive = c(ll.e1.g2.m$cum.surv)*100)
ll.temps.e1.g3.m = data.frame(time=temps.e1.g3$Time_min[1:length(c(ll.e1.g3.m$cum.surv))], alive = c(ll.e1.g3.m$cum.surv)*100)
ll.temps.e1.g4.m = data.frame(time=temps.e1.g4$Time_min[1:length(c(ll.e1.g4.m$cum.surv))], alive = c(ll.e1.g4.m$cum.surv)*100)
ll.temps.e2.g1.m = data.frame(time=temps.e2.g1$Time_min[1:length(c(ll.e2.g1.m$cum.surv))], alive = c(ll.e2.g1.m$cum.surv)*100)
ll.temps.e2.g2.m = data.frame(time=temps.e2.g2$Time_min[1:length(c(ll.e2.g2.m$cum.surv))], alive = c(ll.e2.g2.m$cum.surv)*100)
ll.temps.e2.g3.m = data.frame(time=temps.e2.g3$Time_min[1:length(c(ll.e2.g3.m$cum.surv))], alive = c(ll.e2.g3.m$cum.surv)*100)
ll.temps.e2.g4.m = data.frame(time=temps.e2.g4$Time_min[1:length(c(ll.e2.g4.m$cum.surv))], alive = c(ll.e2.g4.m$cum.surv)*100)
ll.temps.e2.g5.m = data.frame(time=temps.e2.g5$Time_min[1:length(c(ll.e2.g5.m$cum.surv))], alive = c(ll.e2.g5.m$cum.surv)*100)
ll.temps.e3.g1.m = data.frame(time=temps.e3.g1$Time_min[1:length(c(ll.e3.g1.m$cum.surv))], alive = c(ll.e3.g1.m$cum.surv)*100)
ll.temps.e3.g2.m = data.frame(time=temps.e3.g2$Time_min[1:length(c(ll.e3.g2.m$cum.surv))], alive = c(ll.e3.g2.m$cum.surv)*100)
ll.temps.e3.g3.m = data.frame(time=temps.e3.g3$Time_min[1:length(c(ll.e3.g3.m$cum.surv))], alive = c(ll.e3.g3.m$cum.surv)*100)
ll.temps.e3.g4.m = data.frame(time=temps.e3.g4$Time_min[1:length(c(ll.e3.g4.m$cum.surv))], alive = c(ll.e3.g4.m$cum.surv)*100)

ll.lin.sims.e1.g1.f = Surv.to.tf.sims(ll.temps.e1.g1.f)
ll.lin.sims.e1.g2.f =Surv.to.tf.sims(ll.temps.e1.g2.f)
ll.lin.sims.e1.g3.f =Surv.to.tf.sims(ll.temps.e1.g3.f)
ll.lin.sims.e1.g4.f =Surv.to.tf.sims(ll.temps.e1.g4.f)
ll.lin.sims.e2.g1.f =Surv.to.tf.sims(ll.temps.e2.g1.f)
ll.lin.sims.e2.g2.f =Surv.to.tf.sims(ll.temps.e2.g2.f)
ll.lin.sims.e2.g3.f =Surv.to.tf.sims(ll.temps.e2.g3.f)
ll.lin.sims.e2.g4.f =Surv.to.tf.sims(ll.temps.e2.g4.f)
ll.lin.sims.e2.g5.f =Surv.to.tf.sims(ll.temps.e2.g5.f)
ll.lin.sims.e3.g1.f =Surv.to.tf.sims(ll.temps.e3.g1.f)
ll.lin.sims.e3.g2.f =Surv.to.tf.sims(ll.temps.e3.g2.f)
ll.lin.sims.e3.g3.f =Surv.to.tf.sims(ll.temps.e3.g3.f)
ll.lin.sims.e3.g4.f =Surv.to.tf.sims(ll.temps.e3.g4.f)
ll.lin.sims.e1.g1.m =Surv.to.tf.sims(ll.temps.e1.g1.m)
ll.lin.sims.e1.g2.m =Surv.to.tf.sims(ll.temps.e1.g2.m)
ll.lin.sims.e1.g3.m =Surv.to.tf.sims(ll.temps.e1.g3.m)
ll.lin.sims.e1.g4.m =Surv.to.tf.sims(ll.temps.e1.g4.m)
ll.lin.sims.e2.g1.m =Surv.to.tf.sims(ll.temps.e2.g1.m)
ll.lin.sims.e2.g2.m =Surv.to.tf.sims(ll.temps.e2.g2.m)
ll.lin.sims.e2.g3.m =Surv.to.tf.sims(ll.temps.e2.g3.m)
ll.lin.sims.e2.g4.m =Surv.to.tf.sims(ll.temps.e2.g4.m)
ll.lin.sims.e2.g5.m =Surv.to.tf.sims(ll.temps.e2.g5.m)
ll.lin.sims.e3.g1.m =Surv.to.tf.sims(ll.temps.e3.g1.m)
ll.lin.sims.e3.g2.m =Surv.to.tf.sims(ll.temps.e3.g2.m)
ll.lin.sims.e3.g3.m =Surv.to.tf.sims(ll.temps.e3.g3.m)
ll.lin.sims.e3.g4.m =Surv.to.tf.sims(ll.temps.e3.g4.m)
log_like_df$ll.ll.lin[1] = log_likelihood_test(sims=ll.lin.sims.e1.g1.f,data=tf.f.e1.g1)$ll.data
log_like_df$ll.ll.lin[2] = log_likelihood_test(sims=ll.lin.sims.e1.g2.f,data=tf.f.e1.g2)$ll.data
log_like_df$ll.ll.lin[3] = log_likelihood_test(sims=ll.lin.sims.e1.g3.f,data=tf.f.e1.g3)$ll.data
log_like_df$ll.ll.lin[4] = log_likelihood_test(sims=ll.lin.sims.e1.g4.f,data=tf.f.e1.g4)$ll.data
log_like_df$ll.ll.lin[5] = log_likelihood_test(sims=ll.lin.sims.e2.g1.f,data=tf.f.e2.g1)$ll.data
log_like_df$ll.ll.lin[6] = log_likelihood_test(sims=ll.lin.sims.e2.g2.f,data=tf.f.e2.g2)$ll.data
log_like_df$ll.ll.lin[7] = log_likelihood_test(sims=ll.lin.sims.e2.g3.f,data=tf.f.e2.g3)$ll.data
log_like_df$ll.ll.lin[8] = log_likelihood_test(sims=ll.lin.sims.e2.g4.f,data=tf.f.e2.g4)$ll.data
log_like_df$ll.ll.lin[9] = log_likelihood_test(sims=ll.lin.sims.e2.g5.f,data=tf.f.e2.g5)$ll.data
log_like_df$ll.ll.lin[10] = log_likelihood_test(sims=ll.lin.sims.e3.g1.f,data=tf.f.e3.g1)$ll.data
log_like_df$ll.ll.lin[11] = log_likelihood_test(sims=ll.lin.sims.e3.g2.f,data=tf.f.e3.g2)$ll.data
log_like_df$ll.ll.lin[12] = log_likelihood_test(sims=ll.lin.sims.e3.g3.f,data=tf.f.e3.g3)$ll.data
log_like_df$ll.ll.lin[13] = log_likelihood_test(sims=ll.lin.sims.e3.g4.f,data=tf.f.e3.g4)$ll.data
log_like_df$ll.ll.lin[14] = log_likelihood_test(sims=ll.lin.sims.e1.g1.m,data=tf.m.e1.g1)$ll.data
log_like_df$ll.ll.lin[15] = log_likelihood_test(sims=ll.lin.sims.e1.g2.m,data=tf.m.e1.g2)$ll.data
log_like_df$ll.ll.lin[16] = log_likelihood_test(sims=ll.lin.sims.e1.g3.m,data=tf.m.e1.g3)$ll.data
log_like_df$ll.ll.lin[17] = log_likelihood_test(sims=ll.lin.sims.e1.g4.m,data=tf.m.e1.g4)$ll.data
log_like_df$ll.ll.lin[18] = log_likelihood_test(sims=ll.lin.sims.e2.g1.m,data=tf.m.e2.g1)$ll.data
log_like_df$ll.ll.lin[19] = log_likelihood_test(sims=ll.lin.sims.e2.g2.m,data=tf.m.e2.g2)$ll.data
log_like_df$ll.ll.lin[20] = log_likelihood_test(sims=ll.lin.sims.e2.g3.m,data=tf.m.e2.g3)$ll.data
log_like_df$ll.ll.lin[21] = log_likelihood_test(sims=ll.lin.sims.e2.g4.m,data=tf.m.e2.g4)$ll.data
log_like_df$ll.ll.lin[22] = log_likelihood_test(sims=ll.lin.sims.e2.g5.m,data=tf.m.e2.g5)$ll.data
log_like_df$ll.ll.lin[23] = log_likelihood_test(sims=ll.lin.sims.e3.g1.m,data=tf.m.e3.g1)$ll.data
log_like_df$ll.ll.lin[24] = log_likelihood_test(sims=ll.lin.sims.e3.g2.m,data=tf.m.e3.g2)$ll.data
log_like_df$ll.ll.lin[25] = log_likelihood_test(sims=ll.lin.sims.e3.g3.m,data=tf.m.e3.g3)$ll.data
log_like_df$ll.ll.lin[26] = log_likelihood_test(sims=ll.lin.sims.e3.g4.m,data=tf.m.e3.g4)$ll.data

```

```{r}
#now for log-logistic constant
#need to reformat dataframe
ll.cons.temps.e1.g1.f = data.frame(time=temps.e1.g1$Time_min[1:length(c(ll.cons.e1.g1.f$cum.surv))], alive = c(ll.cons.e1.g1.f$cum.surv)*100)
ll.cons.temps.e1.g2.f = data.frame(time=temps.e1.g2$Time_min[1:length(c(ll.cons.e1.g2.f$cum.surv))], alive = c(ll.cons.e1.g2.f$cum.surv)*100)
ll.cons.temps.e1.g3.f = data.frame(time=temps.e1.g3$Time_min[1:length(c(ll.cons.e1.g3.f$cum.surv))], alive = c(ll.cons.e1.g3.f$cum.surv)*100)
ll.cons.temps.e1.g4.f = data.frame(time=temps.e1.g4$Time_min[1:length(c(ll.cons.e1.g4.f$cum.surv))], alive = c(ll.cons.e1.g4.f$cum.surv)*100)
ll.cons.temps.e2.g1.f = data.frame(time=temps.e2.g1$Time_min[1:length(c(ll.cons.e2.g1.f$cum.surv))], alive = c(ll.cons.e2.g1.f$cum.surv)*100)
ll.cons.temps.e2.g2.f = data.frame(time=temps.e2.g2$Time_min[1:length(c(ll.cons.e2.g2.f$cum.surv))], alive = c(ll.cons.e2.g2.f$cum.surv)*100)
ll.cons.temps.e2.g3.f = data.frame(time=temps.e2.g3$Time_min[1:length(c(ll.cons.e2.g3.f$cum.surv))], alive = c(ll.cons.e2.g3.f$cum.surv)*100)
ll.cons.temps.e2.g4.f = data.frame(time=temps.e2.g4$Time_min[1:length(c(ll.cons.e2.g4.f$cum.surv))], alive = c(ll.cons.e2.g4.f$cum.surv)*100)
ll.cons.temps.e2.g5.f = data.frame(time=temps.e2.g5$Time_min[1:length(c(ll.cons.e2.g5.f$cum.surv))], alive = c(ll.cons.e2.g5.f$cum.surv)*100)
ll.cons.temps.e3.g1.f = data.frame(time=temps.e3.g1$Time_min[1:length(c(ll.cons.e3.g1.f$cum.surv))], alive = c(ll.cons.e3.g1.f$cum.surv)*100)
ll.cons.temps.e3.g2.f = data.frame(time=temps.e3.g2$Time_min[1:length(c(ll.cons.e3.g2.f$cum.surv))], alive = c(ll.cons.e3.g2.f$cum.surv)*100)
ll.cons.temps.e3.g3.f = data.frame(time=temps.e3.g3$Time_min[1:length(c(ll.cons.e3.g3.f$cum.surv))], alive = c(ll.cons.e3.g3.f$cum.surv)*100)
ll.cons.temps.e3.g4.f = data.frame(time=temps.e3.g4$Time_min[1:length(c(ll.cons.e3.g4.f$cum.surv))], alive = c(ll.cons.e3.g4.f$cum.surv)*100)
ll.cons.temps.e1.g1.m = data.frame(time=temps.e1.g1$Time_min[1:length(c(ll.cons.e1.g1.m$cum.surv))], alive = c(ll.cons.e1.g1.m$cum.surv)*100)
ll.cons.temps.e1.g2.m = data.frame(time=temps.e1.g2$Time_min[1:length(c(ll.cons.e1.g2.m$cum.surv))], alive = c(ll.cons.e1.g2.m$cum.surv)*100)
ll.cons.temps.e1.g3.m = data.frame(time=temps.e1.g3$Time_min[1:length(c(ll.cons.e1.g3.m$cum.surv))], alive = c(ll.cons.e1.g3.m$cum.surv)*100)
ll.cons.temps.e1.g4.m = data.frame(time=temps.e1.g4$Time_min[1:length(c(ll.cons.e1.g4.m$cum.surv))], alive = c(ll.cons.e1.g4.m$cum.surv)*100)
ll.cons.temps.e2.g1.m = data.frame(time=temps.e2.g1$Time_min[1:length(c(ll.cons.e2.g1.m$cum.surv))], alive = c(ll.cons.e2.g1.m$cum.surv)*100)
ll.cons.temps.e2.g2.m = data.frame(time=temps.e2.g2$Time_min[1:length(c(ll.cons.e2.g2.m$cum.surv))], alive = c(ll.cons.e2.g2.m$cum.surv)*100)
ll.cons.temps.e2.g3.m = data.frame(time=temps.e2.g3$Time_min[1:length(c(ll.cons.e2.g3.m$cum.surv))], alive = c(ll.cons.e2.g3.m$cum.surv)*100)
ll.cons.temps.e2.g4.m = data.frame(time=temps.e2.g4$Time_min[1:length(c(ll.cons.e2.g4.m$cum.surv))], alive = c(ll.cons.e2.g4.m$cum.surv)*100)
ll.cons.temps.e2.g5.m = data.frame(time=temps.e2.g5$Time_min[1:length(c(ll.cons.e2.g5.m$cum.surv))], alive = c(ll.cons.e2.g5.m$cum.surv)*100)
ll.cons.temps.e3.g1.m = data.frame(time=temps.e3.g1$Time_min[1:length(c(ll.cons.e3.g1.m$cum.surv))], alive = c(ll.cons.e3.g1.m$cum.surv)*100)
ll.cons.temps.e3.g2.m = data.frame(time=temps.e3.g2$Time_min[1:length(c(ll.cons.e3.g2.m$cum.surv))], alive = c(ll.cons.e3.g2.m$cum.surv)*100)
ll.cons.temps.e3.g3.m = data.frame(time=temps.e3.g3$Time_min[1:length(c(ll.cons.e3.g3.m$cum.surv))], alive = c(ll.cons.e3.g3.m$cum.surv)*100)
ll.cons.temps.e3.g4.m = data.frame(time=temps.e3.g4$Time_min[1:length(c(ll.cons.e3.g4.m$cum.surv))], alive = c(ll.cons.e3.g4.m$cum.surv)*100)

ll.cons.sims.e1.g1.f = Surv.to.tf.sims(ll.cons.temps.e1.g1.f)
ll.cons.sims.e1.g2.f =Surv.to.tf.sims(ll.cons.temps.e1.g2.f)
ll.cons.sims.e1.g3.f =Surv.to.tf.sims(ll.cons.temps.e1.g3.f)
ll.cons.sims.e1.g4.f =Surv.to.tf.sims(ll.cons.temps.e1.g4.f)
ll.cons.sims.e2.g1.f =Surv.to.tf.sims(ll.cons.temps.e2.g1.f)
ll.cons.sims.e2.g2.f =Surv.to.tf.sims(ll.cons.temps.e2.g2.f)
ll.cons.sims.e2.g3.f =Surv.to.tf.sims(ll.cons.temps.e2.g3.f)
ll.cons.sims.e2.g4.f =Surv.to.tf.sims(ll.cons.temps.e2.g4.f)
ll.cons.sims.e2.g5.f =Surv.to.tf.sims(ll.cons.temps.e2.g5.f)
ll.cons.sims.e3.g1.f =Surv.to.tf.sims(ll.cons.temps.e3.g1.f)
ll.cons.sims.e3.g2.f =Surv.to.tf.sims(ll.cons.temps.e3.g2.f)
ll.cons.sims.e3.g3.f =Surv.to.tf.sims(ll.cons.temps.e3.g3.f)
ll.cons.sims.e3.g4.f =Surv.to.tf.sims(ll.cons.temps.e3.g4.f)
ll.cons.sims.e1.g1.m =Surv.to.tf.sims(ll.cons.temps.e1.g1.m)
ll.cons.sims.e1.g2.m =Surv.to.tf.sims(ll.cons.temps.e1.g2.m)
ll.cons.sims.e1.g3.m =Surv.to.tf.sims(ll.cons.temps.e1.g3.m)
ll.cons.sims.e1.g4.m =Surv.to.tf.sims(ll.cons.temps.e1.g4.m)
ll.cons.sims.e2.g1.m =Surv.to.tf.sims(ll.cons.temps.e2.g1.m)
ll.cons.sims.e2.g2.m =Surv.to.tf.sims(ll.cons.temps.e2.g2.m)
ll.cons.sims.e2.g3.m =Surv.to.tf.sims(ll.cons.temps.e2.g3.m)
ll.cons.sims.e2.g4.m =Surv.to.tf.sims(ll.cons.temps.e2.g4.m)
ll.cons.sims.e2.g5.m =Surv.to.tf.sims(ll.cons.temps.e2.g5.m)
ll.cons.sims.e3.g1.m =Surv.to.tf.sims(ll.cons.temps.e3.g1.m)
ll.cons.sims.e3.g2.m =Surv.to.tf.sims(ll.cons.temps.e3.g2.m)
ll.cons.sims.e3.g3.m =Surv.to.tf.sims(ll.cons.temps.e3.g3.m)
ll.cons.sims.e3.g4.m =Surv.to.tf.sims(ll.cons.temps.e3.g4.m)
log_like_df$ll.ll.cons[1] = log_likelihood_test(sims=ll.cons.sims.e1.g1.f,data=tf.f.e1.g1)$ll.data
log_like_df$ll.ll.cons[2] = log_likelihood_test(sims=ll.cons.sims.e1.g2.f,data=tf.f.e1.g2)$ll.data
log_like_df$ll.ll.cons[3] = log_likelihood_test(sims=ll.cons.sims.e1.g3.f,data=tf.f.e1.g3)$ll.data
log_like_df$ll.ll.cons[4] = log_likelihood_test(sims=ll.cons.sims.e1.g4.f,data=tf.f.e1.g4)$ll.data
log_like_df$ll.ll.cons[5] = log_likelihood_test(sims=ll.cons.sims.e2.g1.f,data=tf.f.e2.g1)$ll.data
log_like_df$ll.ll.cons[6] = log_likelihood_test(sims=ll.cons.sims.e2.g2.f,data=tf.f.e2.g2)$ll.data
log_like_df$ll.ll.cons[7] = log_likelihood_test(sims=ll.cons.sims.e2.g3.f,data=tf.f.e2.g3)$ll.data
log_like_df$ll.ll.cons[8] = log_likelihood_test(sims=ll.cons.sims.e2.g4.f,data=tf.f.e2.g4)$ll.data
log_like_df$ll.ll.cons[9] = log_likelihood_test(sims=ll.cons.sims.e2.g5.f,data=tf.f.e2.g5)$ll.data
log_like_df$ll.ll.cons[10] = log_likelihood_test(sims=ll.cons.sims.e3.g1.f,data=tf.f.e3.g1)$ll.data
log_like_df$ll.ll.cons[11] = log_likelihood_test(sims=ll.cons.sims.e3.g2.f,data=tf.f.e3.g2)$ll.data
log_like_df$ll.ll.cons[12] = log_likelihood_test(sims=ll.cons.sims.e3.g3.f,data=tf.f.e3.g3)$ll.data
log_like_df$ll.ll.cons[13] = log_likelihood_test(sims=ll.cons.sims.e3.g4.f,data=tf.f.e3.g4)$ll.data
log_like_df$ll.ll.cons[14] = log_likelihood_test(sims=ll.cons.sims.e1.g1.m,data=tf.m.e1.g1)$ll.data
log_like_df$ll.ll.cons[15] = log_likelihood_test(sims=ll.cons.sims.e1.g2.m,data=tf.m.e1.g2)$ll.data
log_like_df$ll.ll.cons[16] = log_likelihood_test(sims=ll.cons.sims.e1.g3.m,data=tf.m.e1.g3)$ll.data
log_like_df$ll.ll.cons[17] = log_likelihood_test(sims=ll.cons.sims.e1.g4.m,data=tf.m.e1.g4)$ll.data
log_like_df$ll.ll.cons[18] = log_likelihood_test(sims=ll.cons.sims.e2.g1.m,data=tf.m.e2.g1)$ll.data
log_like_df$ll.ll.cons[19] = log_likelihood_test(sims=ll.cons.sims.e2.g2.m,data=tf.m.e2.g2)$ll.data
log_like_df$ll.ll.cons[20] = log_likelihood_test(sims=ll.cons.sims.e2.g3.m,data=tf.m.e2.g3)$ll.data
log_like_df$ll.ll.cons[21] = log_likelihood_test(sims=ll.cons.sims.e2.g4.m,data=tf.m.e2.g4)$ll.data
log_like_df$ll.ll.cons[22] = log_likelihood_test(sims=ll.cons.sims.e2.g5.m,data=tf.m.e2.g5)$ll.data
log_like_df$ll.ll.cons[23] = log_likelihood_test(sims=ll.cons.sims.e3.g1.m,data=tf.m.e3.g1)$ll.data
log_like_df$ll.ll.cons[24] = log_likelihood_test(sims=ll.cons.sims.e3.g2.m,data=tf.m.e3.g2)$ll.data
log_like_df$ll.ll.cons[25] = log_likelihood_test(sims=ll.cons.sims.e3.g3.m,data=tf.m.e3.g3)$ll.data
log_like_df$ll.ll.cons[26] = log_likelihood_test(sims=ll.cons.sims.e3.g4.m,data=tf.m.e3.g4)$ll.data
```


```{r}
#Make fig 4 panel C
four.mod.results$rz.lin.ll.diff=log_like_df$ll.ll.lin-log_like_df$ll.rz
fig4.pC=ggplot(data=four.mod.results)+
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1.4, 2.45),
    breaks = c(1.47, 1.78, 2.01, 2.38),
    labels = c("30", "60", "120", "240")
  ) +
  geom_point(aes(x=log10(true.tKD.mean),y=rz.lin.ll.diff,color=sex),size=2)+
  ylab(expression(Delta*" log likelihood"))+
  xlab(expression("Observed Median"~t[f]~'(minutes)'))+
  geom_smooth(aes(x=log10(true.tKD.mean),y=rz.lin.ll.diff,color=sex),method="lm",se=FALSE)+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
    axis.title.y=element_blank()
  )+
  labs(tag = "C   ",)+
  theme(
  legend.position = c(0.16, 0.5),   
  legend.justification = c(1, 0),
  legend.key.size = unit(10, "pt"),    
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
  plot.tag = element_text(size = 16)
)+
  geom_abline(slope=0,intercept=0,linetype="dashed")+
  scale_color_grey(start = 0.2, end = 0.6)

```

```{r}
# Remove individual axis titles and legends from each plot
fig4.females2 <- fig4.females + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.title.y=element_blank(), legend.position = "none")

fig4.males2 <- fig4.males + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.title.y=element_blank())

y_lab_plot <- ggplot() + 
  theme_void() +
  theme(
    plot.margin = margin(t = 0, r = 5, b = 0, l = 5)  
  ) +
  geom_text(
    aes(x = 0.5, y = 0.5, label = "Predicted~Median~t[f]~'(minutes)'"),
    parse = TRUE, size = 6, angle = 90
  )

y_lab_plot_2 = ggplot() + 
  theme_void() +
  theme(
    plot.margin = margin(t = 0, r = 5, b = 0, l = 5)  
  ) +
  geom_text(
    aes(x = 0.5, y = 0.5, label = expression(Delta*" log likelihood")),
    parse = TRUE, size = 6, angle = 90
  )

fig4.females2 <- fig4.females2 + labs(tag = "A")
fig4.males2   <- fig4.males2  + labs(tag = "B")

#Assemble everything into one patchwork layout with shared axis labels
top_two <- fig4.females2 / fig4.males2

# top row with shared y-axis label
top_row <- 
  (y_lab_plot | top_two | plot_spacer()) +
  plot_layout(widths = c(0.08, 0.85, 0.07))

bottom_row =(y_lab_plot_2 | fig4.pC | plot_spacer()) +
  plot_layout(widths = c(0.07, 0.86, 0.07))
  
final_plot <- top_row / bottom_row +
  plot_layout(heights = c(0.667, 0.333))
final_plot
```

```{r}
#Quick lm analysis to see if the trends in pC are significant
ll.diff.mod.f = lm(rz.lin.ll.diff~log10(true.tKD.mean),data=four.mod.results[four.mod.results$sex=="f",])
ll.diff.mod.m = lm(rz.lin.ll.diff~log10(true.tKD.mean),data=four.mod.results[four.mod.results$sex=="m",])
plot(ll.diff.mod.f)
summary(ll.diff.mod.f)
plot(ll.diff.mod.m)
summary(ll.diff.mod.m)
#difference in log likelihoods between the two models does not change significantly with temperature
#Let's check if they are significantly >0
#Paired t-test to see if the log logistic model is significantly better than the Rezende model
# t.test(log_like_df$ll.ll.lin[1:13],log_like_df$ll.rz[1:13],paired=TRUE)
# t.test(log_like_df$ll.ll.lin[14:26],log_like_df$ll.rz[14:26],paired=TRUE)
t.test(log_like_df$ll.ll.lin,log_like_df$ll.rz,paired=TRUE)
```


```{r}
#Which is better at predicting failure under fluctuating temperature conditions? Constant or linear shape?
#compare mean log likelihoods between linear and constant log-likelihood models
rank_magnitude <- function(x) {
  rank(-x, ties.method = "first")
}

# Apply to each row
ranked_df <- t(apply(log_like_df[,c("ll.ll.lin","ll.ll.cons")], 1, rank_magnitude))

# Convert result to data frame and add names if needed
ranked_df <- as.data.frame(ranked_df)
colnames(ranked_df) <- colnames(log_like_df[,c("ll.ll.lin","ll.ll.cons")])

print(ranked_df)

#average ranking
apply(ranked_df,2,sum)/nrow(ranked_df)
```

```{r}
#compare mean log likelihoods between linear and Rezende

# Apply to each row
ranked_df <- t(apply(log_like_df[,c("ll.ll.lin","ll.rz")], 1, rank_magnitude))

# Convert result to data frame and add names if needed
ranked_df <- as.data.frame(ranked_df)
colnames(ranked_df) <- colnames(log_like_df[,c("ll.ll.lin","ll.rz")])

print(ranked_df)

#average ranking
apply(ranked_df,2,sum)/nrow(ranked_df)
```


```{r}
#This function takes data from the above models and turns it into a survival curve plot
#exp.data=experimental data
#data1->results from model 1 
plot.surv.curves = function(data1,data2,exp.data,modtitle1,modtitle2,title){
# Function to convert a vector into a data frame with group label
make_df <- function(vec, group_name) {
  data.frame(time = vec, status = 1, group = group_name)
}

# Combine all data
df <- bind_rows(
  make_df(data1, modtitle1),
  make_df(data2, modtitle2),
  make_df(exp.data, "Data")
)

# Set group order: modtitle1, modtitle2, then "Data"
df$group <- factor(df$group, levels = c(modtitle1, modtitle2, "Data"))

# Fit KM model
km_fit <- survfit(Surv(time, status) ~ group, data = df)
tidy_km <- broom::tidy(km_fit)

# Clean and reorder strata names for matching
tidy_km$strata <- case_match(tidy_km$strata,
                         paste0("group=",modtitle1) ~ modtitle1,
                         paste0("group=",modtitle2)  ~ modtitle2,
                         "group=Data" ~ "Data")

# Set factor order for legend
tidy_km$strata <- factor(tidy_km$strata, levels = c(modtitle1, modtitle2, "Data"))

p = ggplot(tidy_km, aes(x = time, y = estimate, color = strata, linetype = strata)) +
  geom_step(linewidth = 1.6) +
  labs(
    y = "Cumulative Survival",
    x = "Time (minutes)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    legend.title = element_blank()
  )+
  ggtitle(title)
return(p)}
```

```{r}
p1 = plot.surv.curves(data1=ll.lin.sims.e1.g1.f,data2=rz.sims.e1.g1.f,exp.data = tf.f.e1.g1,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 1, Females")
p2 = plot.surv.curves(data1=ll.lin.sims.e1.g2.f,data2=rz.sims.e1.g2.f,exp.data = tf.f.e1.g2,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 2, Females")
p3 = plot.surv.curves(data1=ll.lin.sims.e1.g3.f,data2=rz.sims.e1.g3.f,exp.data = tf.f.e1.g3,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 3, Females")
p4 = plot.surv.curves(data1=ll.lin.sims.e1.g4.f,data2=rz.sims.e1.g4.f,exp.data = tf.f.e1.g4,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 4, Females")
p5 = plot.surv.curves(data1=ll.lin.sims.e2.g1.f,data2=rz.sims.e2.g1.f,exp.data = tf.f.e2.g1,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 1, Females")
p6 = plot.surv.curves(data1=ll.lin.sims.e2.g2.f,data2=rz.sims.e2.g2.f,exp.data = tf.f.e2.g2,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 2, Females")
p7 = plot.surv.curves(data1=ll.lin.sims.e2.g3.f,data2=rz.sims.e2.g3.f,exp.data = tf.f.e2.g3,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 3, Females")
p8 = plot.surv.curves(data1=ll.lin.sims.e2.g4.f,data2=rz.sims.e2.g4.f,exp.data = tf.f.e2.g4,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 4, Females")
p9 = plot.surv.curves(data1=ll.lin.sims.e2.g5.f,data2=rz.sims.e2.g5.f,exp.data = tf.f.e2.g5,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 5, Females")
p10 = plot.surv.curves(data1=ll.lin.sims.e3.g1.f,data2=rz.sims.e3.g1.f,exp.data = tf.f.e3.g1,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 1, Females")
p11 = plot.surv.curves(data1=ll.lin.sims.e3.g2.f,data2=rz.sims.e3.g2.f,exp.data = tf.f.e3.g2,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 2, Females")
p12 = plot.surv.curves(data1=ll.lin.sims.e3.g3.f,data2=rz.sims.e3.g3.f,exp.data = tf.f.e3.g3,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 3, Females")
p13 = plot.surv.curves(data1=ll.lin.sims.e3.g4.f,data2=rz.sims.e3.g4.f,exp.data = tf.f.e3.g4,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 4, Females")
p14 = plot.surv.curves(data1=ll.lin.sims.e1.g1.m,data2=rz.sims.e1.g1.m,exp.data = tf.m.e1.g1,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 1, Males")
p15 = plot.surv.curves(data1=ll.lin.sims.e1.g2.m,data2=rz.sims.e1.g2.m,exp.data = tf.m.e1.g2,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 2, Males")
p16 = plot.surv.curves(data1=ll.lin.sims.e1.g3.m,data2=rz.sims.e1.g3.m,exp.data = tf.m.e1.g3,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 3, Males")
p17 = plot.surv.curves(data1=ll.lin.sims.e1.g4.m,data2=rz.sims.e1.g4.m,exp.data = tf.m.e1.g4,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 1, Group 4, Males")
p18 = plot.surv.curves(data1=ll.lin.sims.e2.g1.m,data2=rz.sims.e2.g1.m,exp.data = tf.m.e2.g1,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 1, Males")
p19 = plot.surv.curves(data1=ll.lin.sims.e2.g2.m,data2=rz.sims.e2.g2.m,exp.data = tf.m.e2.g2,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 2, Males")
p20 = plot.surv.curves(data1=ll.lin.sims.e2.g3.m,data2=rz.sims.e2.g3.m,exp.data = tf.m.e2.g3,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 3, Males")
p21 = plot.surv.curves(data1=ll.lin.sims.e2.g4.m,data2=rz.sims.e2.g4.m,exp.data = tf.m.e2.g4,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 4, Males")
p22 = plot.surv.curves(data1=ll.lin.sims.e2.g5.m,data2=rz.sims.e2.g5.m,exp.data = tf.m.e2.g5,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 2, Group 5, Males")
p23 = plot.surv.curves(data1=ll.lin.sims.e3.g1.m,data2=rz.sims.e3.g1.m,exp.data = tf.m.e3.g1,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 1, Males")
p24 = plot.surv.curves(data1=ll.lin.sims.e3.g2.m,data2=rz.sims.e3.g2.m,exp.data = tf.m.e3.g2,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 2, Males")
p25 = plot.surv.curves(data1=ll.lin.sims.e3.g3.m,data2=rz.sims.e3.g3.m,exp.data = tf.m.e3.g3,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 3, Males")
p26 = plot.surv.curves(data1=ll.lin.sims.e3.g4.m,data2=rz.sims.e3.g4.m,exp.data = tf.m.e3.g4,modtitle1="Increasing Variance", modtitle2="Rezende Model", title="Exp. 3, Group 4, Males")
```

```{r}
wrap_plots(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13)+
plot_layout(guides = 'collect')&
  theme(legend.position="bottom",
        legend.key.size = unit(0.3, "cm"),
    legend.key = element_rect(color = NA, fill = NA),
    legend.text = element_text(face = "italic"))
wrap_plots(p14,p15,p16,p17,p18,p19,p20,p21,p22,p23,p24,p25,p26)+
plot_layout(guides = 'collect')&
  theme(legend.position="bottom",
        legend.key.size = unit(0.3, "cm"),
    legend.key = element_rect(color = NA, fill = NA),
    legend.text = element_text(face = "italic"))
```

